### 2024-09-30 manuscript/main.tex
main template

\documentclass[conference]{IEEEtran}

\usepackage{amsfonts,amsmath,amssymb}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{natbib}

%notations
\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{defn}[thm]{Definition}
\newtheorem{exm}[thm]{Example}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{rem}[thm]{Remark}

\newcommand{\corref}[1]{Corollary~{\rm \ref{#1}}}
\newcommand{\defref}[1]{Definition~{\rm \ref{#1}}}
\newcommand{\exmref}[1]{Example~{\rm \ref{#1}}}
\newcommand{\lemref}[1]{Lemma~{\rm \ref{#1}}}
\newcommand{\propref}[1]{Proposition~{\rm \ref{#1}}}
\newcommand{\prpropref}[1]{Proposition~{\bf \ref{#1}}}
\newcommand{\remref}[1]{Remark~{\rm \ref{#1}}}
\newcommand{\thmref}[1]{Theorem~{\rm \ref{#1}}}
\newcommand{\ththmref}[1]{Theorem~{\bf \ref{#1}}}

\renewcommand{\thesection}{\arabic{section}}
\renewcommand{\theequation}{\thesection.\arabic{equation}}
\newcommand{\beq}[1]{\begin{equation} \label{#1}}
	\newcommand{\eeq}{\end{equation}}
\newcommand{\bea}{\bed\begin{array}{rl}}
	\newcommand{\eea}{\end{array}\eed}
\newcommand{\bed}{\begin{displaymath}}
	\newcommand{\eed}{\end{displaymath}}
\newcommand{\barray}{\begin{array}{ll}}
	\newcommand{\earray}{\end{array}}

\newcommand{\aad}{&\disp}
\newcommand{\ad}{&\!\!\!\disp}
\newcommand{\al}{\alpha}
\newcommand{\be}{\beta}
\newcommand{\disp}{\displaystyle}
\newcommand{\dl}{\delta}
\newcommand{\E}{{\mathbb E}}
\newcommand{\e}{\varepsilon}
\newcommand{\ee}{\Delta}
\newcommand{\la}{\lambda}
\newcommand{\PP}{{\mathbb P}}
\newcommand{\nd}{\noindent}
\newcommand{\sg}{\sigma}

\newcommand{\NM}{{\rm NM}}
\newcommand{\AM}{{\rm AM}}

\newcommand{\A}{{\cal A}}
\newcommand{\F}{{\cal F}}
\newcommand{\I}{{\cal I}}
\newcommand{\M}{{\cal M}}
\newcommand{\U}{{\cal U}}
\newcommand{\X}{{\cal X}}
\newcommand{\II}{{\cal I}}
\newcommand{\HH}{{\cal H}}

\newcommand{\bdd}{\hspace*{-0.08in}{\bf.}\hspace*{0.05in}}

\newcommand{\lbar}{\overline}
\newcommand{\wdt}{\widetilde}
\newcommand{\wdh}{\widehat}
\newcommand{\diag}{{\rm diag}}
\def\th{\theta}
\def\para#1{\vskip 0.4\baselineskip\noindent{\bf #1}}
\def\paraa#1{\vskip 0.4\baselineskip\noindent{\sf #1}}

\def\rr{{\Bbb R}}
\def\({\left(}
\def\){\right)}
\def\one{{\hbox{1{\kern -0.35em}1}}}
\def\op{{\cal L}}

\def\gg{{\cal G}}
\def\sth{\sum^2_{\th=1}}

\begin{document}

\title{CSE 5830 Course Project Proposal}

\author{\IEEEauthorblockN{Xiaohang Ma}
\IEEEauthorblockA{\textit{Department of Mathematics} \\
\textit{University of Connecticut}\\
Storrs, CT, USA \\
xiaohang.ma@uconn.edu}
\and
\IEEEauthorblockN{Shiying Xiao}
\IEEEauthorblockA{\textit{Department of Statistics} \\
\textit{University of Connecticut}\\
Storrs, CT, USA \\
shiying.xiao@uconn.edu}
\and
\IEEEauthorblockN{Xiaohui Yin}
\IEEEauthorblockA{\textit{Department of Statistics} \\
\textit{University of Connecticut}\\
Storrs, CT, USA \\
xiaohui.yin@uconn.edu}
}

\maketitle

\begin{abstract}
In this project, we will integrate powerful deep neural networks and
statistical methods with a probabilistic graphical model (PGM) to effectively
tackle sequence labeling tasks on a heterogeneous medical care dataset.
\end{abstract}

\begin{IEEEkeywords}
probabilistic graphical model
\end{IEEEkeywords}

\section{Introduction}

\citep{huang2015bidirectional}

\bibliographystyle{plain}
\bibliography{refs}

\end{document}

### 2024-10-04 slides/proposal.tex
proposal slides template

\documentclass[leqno,aspectratio=169]{beamer}
\usetheme{Madrid}
\usecolortheme{seahorse}

\usepackage[linesnumbered,algoruled,boxed,lined]{algorithm2e}
\usepackage{amsfonts,amsmath,amsthm}
\usepackage{bm}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{color}
\usepackage{graphicx}
\graphicspath{{./}{../image/}} % graphic path
\usepackage{hyperref}
\hypersetup{colorlinks,citecolor=blue,linkcolor=[RGB]{50,50,172}}
\usepackage{multicol}
\usepackage{multirow}
\usepackage[authoryear]{natbib}
\usepackage{setspace}
\usepackage{soul}
\usepackage{textpos}
\usepackage{tikz}

%% notations
\newcommand{\blue}[1]{\textcolor{blue}{#1}}
\newcommand{\red}[1]{\textcolor{red}{#1}}
\newcommand{\bmI}{\bm{I}}
\newcommand{\bmOmega}{\bm{\Omega}}
\newcommand{\bmSigma}{\bm{\Sigma}}
\newcommand{\bmS}{\bm{S}}
\newcommand{\bmW}{\bm{W}}
\newcommand{\bmX}{\bm{X}}
\newcommand{\bmY}{\bm{Y}}
\newcommand{\bmZ}{\bm{Z}}
\newcommand{\bmbeta}{\bm{\beta}}
\newcommand{\hatbmOmega}{\hat{\bm{\Omega}}}
\newcommand{\tr}{\text{tr}}
\DeclareMathOperator*{\argmin}{arg \, min}

% consistent with R manual
\newcommand\pkg[1]{\texttt{#1}}
\let\proglang=\textsf \let\code=\texttt

\setbeamercovered{transparent}
\setbeamertemplate{caption}[numbered]
\setbeamertemplate{enumerate item}[default]
\setbeamertemplate{itemize item}[circle]
\setbeamertemplate{section in toc}[default]
\setbeamertemplate{subsection in toc}[default]

\AtBeginSection[]{
\begin{frame}<beamer>{Overview}
\tableofcontents[currentsection]
\end{frame}}

\title[\textcolor{black}{FCABN}]{
{\bf \normalsize Title}}
%\subtitle[]{}
\author[Xiaohang Ma, Shiying Xiao, Xiaohui Yin]{\bf
	Xiaohang Ma\textsuperscript{1},
	Shiying Xiao\textsuperscript{2},
	Xiaohui Yin\textsuperscript{2}}
\institute[UConn]{
\textsuperscript{1}{\small Department of Mathematics, University of Connecticut}
\\
\textsuperscript{2}{\small Department of Statistics, University of Connecticut}
}
\date[October 7, 2024]{
{\small Storrs, CT} \\
{\small October 7, 2024}}

\begin{document}

\begin{frame}[plain]
\titlepage
\end{frame}

\addtobeamertemplate{frametitle}{}{
\begin{textblock*}{70mm}(.9\textwidth,-0.65cm)
\includegraphics[width=.2\textwidth]{uconn-wordmark-single-blue.png}
\end{textblock*}}

\begin{frame}
\frametitle{Overview}
\tableofcontents
\end{frame}

\section[Introduction]{Introduction}

\begin{frame}
\frametitle{Motivations}
\begin{itemize}
\item Functional magnetic resonance imaging (fMRI), especially resting-state 
fMRI, is crucial for understanding brain interactions and neurological diseases 
like Alzheimer's disease (AD).
\bigskip
\item Recent studies combine fMRI and tau-PET imaging to link brain 
architecture with tau protein accumulation, a key feature of AD.
\bigskip
\item Network-based functional connectivity analysis methods have emerged as a 
powerful tool for exploring interactions among brain regions.
\bigskip
\item A comprehensive examination of the methods has been lacking.
\end{itemize}
\end{frame}

\begin{frame}[allowframebreaks]
\frametitle{References}
\bibliographystyle{chicago}
\bibliography{../manuscript/refs}
\end{frame}

\end{document}

### 2024-10-06 slides/Slides-Xiaohui/intro.tex
slides of eills

\section{Introduction}
\begin{frame}
  \frametitle{Motivating Example: Risks of Using Spurious Variables}
   A task to classify cows and camels based on extracted hierarchical features: 
  \begin{itemize}
  \item There is a dataset $\mathcal D$ containing 10k images of cows and camels from the Internet. When Using 70\% to train the classifier, top two features are the back shape $x_1$, and the background color $x_2$, and it performs well on the remaining 30\% testing set. It is contemplated that cows often appear on the grass while most camels appear on the sand from this dataset.
  \end{itemize}
  \vspace{6pt}
  \centering
  \begin{minipage}{.3\textwidth}
    \begin{figure}[H]
      \centering \includegraphics[width=0.8\textwidth]{figs/camel_sand.png}
      \caption{Camel in brown}
    \end{figure}
  \end{minipage}% This must go next to `\end{minipage}`
  \begin{minipage}{.3\textwidth}
    \begin{figure}[H]
      \centering \includegraphics[width=0.8\textwidth]{figs/cow_grass.jpg}
      \caption{Cow in green}
    \end{figure}
  \end{minipage}
\end{frame}

\begin{frame}
  \frametitle{Motivating Example (Cond)}
  \begin{itemize}
  \item However, introducing $x_2$ is not what we expected: the classifier can not work well in a place farming camels and cows, in which the background color is fixed.
  \item If we have another dataset $\widetilde {\mathcal D}$, in which the association between the background color and object label still exists yet slightly perturbs. Intuitively, we may infer that $x_2$ may be a ``spurious'' variable for prediction or causation.
  \end{itemize}
  \vspace{6pt}
  \centering
  \begin{minipage}{.3\textwidth}
    \begin{figure}[H]
      \centering \includegraphics[width=0.8\textwidth]{figs/camel_grass}
      \caption{Camel in green}
    \end{figure}
  \end{minipage}% This must go next to `\end{minipage}`
  \begin{minipage}{.3\textwidth}
    \begin{figure}[H]
      \centering \includegraphics[width=0.8\textwidth]{figs/cow_sand.jpeg}
      \caption{Cow in brown}
    \end{figure}
  \end{minipage}
\end{frame}

\begin{frame}
  \frametitle{Some Definition}
  This paper propose the multiple-environment version of linear least squares. It utilizes the heterogeneity across datasets.
  It combines the linear least squares solutions across different datasets and uses their differences to determine the true parameter and important variables in a completely data-driven manner.

  \vspace{8pt}

  Here are some definitions needed: 
    \begin{itemize}
    \item response variable $y \in \mathbb R$
    \item explanatory variable $\bx \in \mathbb R^p$
    \item set of environments $\mathcal E$
    \item observations from environment $e\in \mathcal E$:
      $(\bx_1^{(e)}, y_1^{(e)}), \dots,  (\bx_n^{(e)}, y_n^{(e)}) \sim \mu^{(e)}$
    \end{itemize}

\end{frame}

\begin{frame}
  \frametitle{Model Formulation}
  The true linear model is assumed as
  \begin{align}
    \label{main_eq}
    y^{(e)} = (\bm\beta^*_{S^*})^\top \bx_{S^*}^{(e)} + \varepsilon^{(e)} ~~~~~ \text{with} ~~~~ \mathbb{E}[\varepsilon^{(e)}|\bx_{S^*}^{(e)}] \equiv 0,
  \end{align} where the unknown set of important variables $S^* = \{j: \beta_j^* \neq 0\}$ and the model parameters $\bm \beta^*$ are the same across different environments, while $\mu^{(e)}$, the distribution of $(\bx^{(e)},y^{(e)})$, may vary.

  \vspace{8pt}

  The aim is to estimate $\bm\beta^*$ and $S^*$ using the $n\cdot |\mathcal{E}|$ data $\{(\bx_i^{(e)}, y_i^{(e)})\}_{e\in \mathcal{E}, i\in \{1,\ldots, n\}}$.
\end{frame}

\begin{frame}
  \frametitle{Challenges}
  \begin{itemize}
  \item $\bm \beta^*$ is mot the best linear predictor for each single environment.
  \item A smaller mean squared error by incorporating the linear spurious variables $\bm x_{S^{*\perp}}$ defined outside the set of important variables for all environments.
    $\bm x_{S^{*\perp}}^{(e)}$ can capture information in $\varepsilon^{(e)}$.
  \item Such spurious variables are not stable when generalized to other environments, because the association between these variables and $y$ are not stable. (background color in the motivating example)
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Loss Function}
  The population-level EILLS objective is 
  \begin{scriptsize}
  \begin{align}
    \mathsf{Q}_\gamma(\bm\beta) &= \underbrace{\sum_{e\in \mathcal{E}} \mathbb{E} \left[|y^{(e)} - \bm\beta^\top \bx^{(e)}|^2\right]}_{\mathsf{R}(\bm\beta)} \nonumber\\
                                &\qquad + \gamma \underbrace{\sum_{j=1}^p \mathbf{1}\{\beta_j\neq 0\} \times  \sum_{e\in \mathcal{E}}  \left|\mathbb{E} [(y^{(e)} - \bm\beta^\top \bx^{(e)}) x_j^{(e)}]\right|^2}_{\mathsf{J}(\bm\beta)},
  \end{align}
  \end{scriptsize}
  where $\gamma$ is a tuning parameter.
  \begin{itemize}
  \item $\mathsf R(\bm\beta)$ requires a good overall solution in each single environment.
  \item $\mathsf J(\bm\beta)$ discourages selecting variables that have strong correlation with the fitted residuals in some environments.
  \item From computational perspective, we have
    \begin{align}
    \mathsf{J}(\bm \beta) &= \sum_{j=1}^p \mathbf{1}\{\beta_j \neq 0\} \sum_{e\in \mathcal{E}} \frac{1}{4} |\nabla_j \mathsf{R}^{(e)}(\bm\beta)|^2,
  \end{align}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Empirical Loss Function}
  \begin{align*}
    \hat{\mathsf{J}}(\bm\beta) &= \sum_{j=1}^p \mathbf{1}\{\beta_j \neq 0\} \sum_{e\in \mathcal{E}}  \left| \hat{\mathbb{E}}[x_j^{(e)} (y^{(e)} - \bm\beta^\top \bx^{(e)})]\right|^2\\
    \hat{\mathsf{R}}(\bm\beta)& = \sum_{e\in \mathcal{E}}  \frac{1}{n^{(e)}} \sum_{i=1}^{n^{(e)}} \{ y^{(e)}_i - \bm\beta^\top \bx^{(e)}_i \}^2\\
    \hat{\mathsf{Q}}(\bm\beta,\gamma) &= \hat{\mathsf{R}}(\bm\beta) + \lambda \hat{\mathsf{J}}(\bm\beta)
  \end{align*}
  Some linear exogenous variables which do not contribute to explaining $y$ can
  be eliminated further by introducing an $l_0$ penalty as
    \begin{align*}
   \hat{\mathsf{L}}(\bm\beta,\gamma,\lambda) =  \hat{\mathsf{R}}(\bm\beta) + \lambda \hat{\mathsf{J}}(\bm\beta) + \lambda \|\bm\beta\|_0.
  \end{align*}
\end{frame}

\begin{frame}
  \frametitle{An Illustration}
  \begin{figure}[H]
      \centering \includegraphics[width=0.8\textwidth]{figs/example}
      \caption{An Example for Illustration, figure from \cite{fan2023environmentinvariantlinearsquares}}
    \end{figure}
\end{frame}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "template"
%%% End:

### 2024-10-06 slides/literature-xiaohui.tex
revise proposal slides

--- 2024-10-06 slides/Slides-Xiaohui/intro.tex
@@
     \item response variable $y \in \mathbb R$
-    \item explanatory variable $\bx \in \mathbb R^p$
+    \item explanatory variable $\bm x \in \mathbb R^p$
     \item set of environments $\mathcal E$
     \item observations from environment $e\in \mathcal E$:
-      $(\bx_1^{(e)}, y_1^{(e)}), \dots,  (\bx_n^{(e)}, y_n^{(e)}) \sim \mu^{(e)}$
+      $(\bm x_1^{(e)}, y_1^{(e)}), \dots,  (\bm x_n^{(e)}, y_n^{(e)}) \sim \mu^{(e)}$
     \end{itemize}
@@
     \label{main_eq}
-    y^{(e)} = (\bm\beta^*_{S^*})^\top \bx_{S^*}^{(e)} + \varepsilon^{(e)} ~~~~~ \text{with} ~~~~ \mathbb{E}[\varepsilon^{(e)}|\bx_{S^*}^{(e)}] \equiv 0,
-  \end{align} where the unknown set of important variables $S^* = \{j: \beta_j^* \neq 0\}$ and the model parameters $\bm \beta^*$ are the same across different environments, while $\mu^{(e)}$, the distribution of $(\bx^{(e)},y^{(e)})$, may vary.
+    y^{(e)} = (\bm\beta^*_{S^*})^\top \bm x_{S^*}^{(e)} + \varepsilon^{(e)} ~~~~~ \text{with} ~~~~ \mathbb{E}[\varepsilon^{(e)}|\bm x_{S^*}^{(e)}] \equiv 0,
+  \end{align} where the unknown set of important variables $S^* = \{j: \beta_j^* \neq 0\}$ and the model parameters $\bm \beta^*$ are the same across different environments, while $\mu^{(e)}$, the distribution of $(\bm x^{(e)},y^{(e)})$, may vary.
 
@@
 
-  The aim is to estimate $\bm\beta^*$ and $S^*$ using the $n\cdot |\mathcal{E}|$ data $\{(\bx_i^{(e)}, y_i^{(e)})\}_{e\in \mathcal{E}, i\in \{1,\ldots, n\}}$.
+  The aim is to estimate $\bm\beta^*$ and $S^*$ using the $n\cdot |\mathcal{E}|$ data $\{(\bm x_i^{(e)}, y_i^{(e)})\}_{e\in \mathcal{E}, i\in \{1,\ldots, n\}}$.
 \end{frame}
@@
   \begin{scriptsize}
-  \begin{align}
-    \mathsf{Q}_\gamma(\bm\beta) &= \underbrace{\sum_{e\in \mathcal{E}} \mathbb{E} \left[|y^{(e)} - \bm\beta^\top \bx^{(e)}|^2\right]}_{\mathsf{R}(\bm\beta)} \nonumber\\
-                                &\qquad + \gamma \underbrace{\sum_{j=1}^p \mathbf{1}\{\beta_j\neq 0\} \times  \sum_{e\in \mathcal{E}}  \left|\mathbb{E} [(y^{(e)} - \bm\beta^\top \bx^{(e)}) x_j^{(e)}]\right|^2}_{\mathsf{J}(\bm\beta)},
-  \end{align}
+  \begin{align*}
+    \mathsf{Q}_\gamma(\bm\beta) &= \underbrace{\sum_{e\in \mathcal{E}} \mathbb{E} \left[|y^{(e)} - \bm\beta^\top \bm x^{(e)}|^2\right]}_{\mathsf{R}(\bm\beta)} \nonumber\\
+                                &\qquad + \gamma \underbrace{\sum_{j=1}^p \mathbf{1}\{\beta_j\neq 0\} \times  \sum_{e\in \mathcal{E}}  \left|\mathbb{E} [(y^{(e)} - \bm\beta^\top \bm x^{(e)}) x_j^{(e)}]\right|^2}_{\mathsf{J}(\bm\beta)},
+  \end{align*}
   \end{scriptsize}
@@
   \item From computational perspective, we have
-    \begin{align}
+    \begin{align*}
     \mathsf{J}(\bm \beta) &= \sum_{j=1}^p \mathbf{1}\{\beta_j \neq 0\} \sum_{e\in \mathcal{E}} \frac{1}{4} |\nabla_j \mathsf{R}^{(e)}(\bm\beta)|^2,
-  \end{align}
+  \end{align*}
   \end{itemize}
@@
   \begin{align*}
-    \hat{\mathsf{J}}(\bm\beta) &= \sum_{j=1}^p \mathbf{1}\{\beta_j \neq 0\} \sum_{e\in \mathcal{E}}  \left| \hat{\mathbb{E}}[x_j^{(e)} (y^{(e)} - \bm\beta^\top \bx^{(e)})]\right|^2\\
-    \hat{\mathsf{R}}(\bm\beta)& = \sum_{e\in \mathcal{E}}  \frac{1}{n^{(e)}} \sum_{i=1}^{n^{(e)}} \{ y^{(e)}_i - \bm\beta^\top \bx^{(e)}_i \}^2\\
+    \hat{\mathsf{J}}(\bm\beta) &= \sum_{j=1}^p \mathbf{1}\{\beta_j \neq 0\} \sum_{e\in \mathcal{E}}  \left| \hat{\mathbb{E}}[x_j^{(e)} (y^{(e)} - \bm\beta^\top \bm x^{(e)})]\right|^2\\
+    \hat{\mathsf{R}}(\bm\beta)& = \sum_{e\in \mathcal{E}}  \frac{1}{n^{(e)}} \sum_{i=1}^{n^{(e)}} \{ y^{(e)}_i - \bm\beta^\top \bm x^{(e)}_i \}^2\\
     \hat{\mathsf{Q}}(\bm\beta,\gamma) &= \hat{\mathsf{R}}(\bm\beta) + \lambda \hat{\mathsf{J}}(\bm\beta)
@@
 %%% mode: latex
-%%% TeX-master: "template"
+%%% TeX-master: "proposal"
 %%% End:

### 2024-10-09 proposal/CRF_formula.tex
proposal slides

\documentclass{article}

\usepackage{mathrsfs}
\usepackage{calligra}
% \usepackage{algorithm}
% \usepackage{algpseudocode}
% \usepackage{algorithm2e}
% \usepackage{algorithmicx}
\usepackage{multicol}
\usepackage{amsmath}  % Ensure amsmath is loaded
\usepackage{amssymb}  % Load amssymb for additional math symbols
\def\X{\mathbb{X}}
\newcommand{\C}{C([-\tau, 0]: \rr^r)}
\newcommand{\B}{{\mathscr B}}
\newcommand{\D}{{\mathscr D}}
\newcommand{\Var}{{{\mathbb V}\rm ar}}
\newcommand{\Cov}{{{\mathbb C}\rm ov}}
\newcommand{\jj}{\jmath}
\newcommand{\DD}{\mathbb{D}}
\newcommand{\sg}{\sigma}
\newcommand{\sgn}{{\operatorname{sgn}}}
\newcommand{\Sg}{{\operatorname{Sgn}}}
\def\clG{{\mathcal G}}
\newcommand{\wh}{{\wdh W}}
\newcommand{\Ga}{\Gamma}
%\newcommand{\sg}{\sigma}
\newcommand{\one}{{\rm 1}\hspace*{-0.035in}{\rm l}}
\newcommand{\diag}{{\rm diag}}
\newcommand{\lbar}{\overline}
\newcommand{\wdh}{\widehat}
\newcommand{\lf}{\lfloor}
\newcommand{\rf}{\rfloor}
\newcommand{\dl}{\delta}
\newcommand{\ga}{\gamma}
\newcommand{\ps}{\psi}
\newcommand{\Dl}{\Delta}
\newcommand{\la}{\lambda}
\def\op{{\cal L}}
\def\l{\left|}
\def\r{\right|}
\newcommand{\mz}{{m}}
\newcommand{\F}{{\cal F}}
\newcommand{\U}{{\cal U}}
\newcommand{\wdt}{\widetilde}
\newcommand{\e}{\varepsilon}
\newcommand{\rr}{{\mathbb R}}
\newcommand{\M}{{\cal M}}
\newcommand{\cd}{(\cdot)}
\newcommand{\nd}{\noindent}
\newcommand{\tr}{{\rm tr}}
\def\para#1{\vskip .25\baselineskip\noindent{\bf #1}}
\def\qed{\strut\hfill $\Box$}
\def\eqdef {\stackrel {\rm def}{=}}
\newtheorem{thm}{Theorem}[section]
\newtheorem{prop}[thm]{Proposition}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{rem}[thm]{Remark}
\newtheorem{defn}[thm]{Definition}
\newtheorem{exm}[thm]{Example}
\newtheorem{asm}{Assumption}
% \newtheorem{algorithm}{Algorithm}
\newcommand{\thmref}[1]{Theorem~{\rm \ref{#1}}}
\newcommand{\lemref}[1]{Lemma~{\rm \ref{#1}}}
\newcommand{\corref}[1]{Corollary~{\rm \ref{#1}}}
\newcommand{\propref}[1]{Proposition~{\rm \ref{#1}}}
\newcommand{\defref}[1]{Definition~{\rm \ref{#1}}}
\newcommand{\remref}[1]{Remark~{\rm \ref{#1}}}
\newcommand{\exmref}[1]{Example~{\rm \ref{#1}}}
\newcommand{\figref}[1]{Figure~{\rm \ref{#1}}}
\newcommand{\lelemref}[1]{Lemma~{\bf \ref{#1}}}
\newcommand{\ththmref}[1]{Theorem~{\bf \ref{#1}}}

\def\al{\alpha}
\def\th{\theta}
\def\dl{\delta}

\numberwithin{equation}{section}
\newcommand{\bdd}{\hspace*{-0.08in}{\bf.}\hspace*{0.05in}}
\newcommand{\beq}[1]{\begin{equation} \label{#1}}
	\newcommand{\eeq}{\end{equation}}
\newcommand{\bed}{\begin{displaymath}}
	\newcommand{\eed}{\end{displaymath}}
\newcommand{\bea}{\bed\begin{array}{rl}}
	\newcommand{\eea}{\end{array}\eed}
\newcommand{\ad}{&\!\!\!\disp}
\newcommand{\aad}{&\disp}
\newcommand{\barray}{\begin{array}{ll}}
	\newcommand{\earray}{\end{array}}
\def\({\left(}
\def\){\right)}
\def\disp{\displaystyle}
\newcommand{\LL}{{\mathfrak L}}
\def\dbL{{\rm I\hskip-2.2pt L}}
\def\clM{{\cal M}}

\def\clD{{\cal D}}
\def\clG{{\cal G}}
\def\clV{{\cal V}}
\def\clE{{\cal E}}
\newcommand{\CO}{\mathcal{O}}
%\newcommand{\Nnvspc}{\vspace*{-3mm}}
\newcommand{\Nnvspc}{\vspace*{-0.5mm}}
%\newcommand{\Nvspc}{\vspace*{-2mm}}
\newcommand{\Nvspc}{\vspace*{-0.5mm}}
\newcommand{\nvspc}{\vspace*{-0.5mm}}

\newcommand{\R}{\mathbb{R}}
\newcommand{\PP}{\mathbb{P}}
\newcommand{\p}{\mathfrak{p}}
\newcommand{\EE}{\mathbb{E}}
\newcommand{\N}{{\mathbb{Z}}_+}
\newcommand{\m}{\mathfrak{m}}
\newcommand{\Z}{{\cal Z}}
\newcommand{\1}{\boldsymbol{1}}
\def\indi{{\bf 1}}
\newcommand{\BF}{\mathbb{F}}
\newcommand{\trace}{\hbox{\rm tr}}
\def\clN{{\cal N}}
\newcommand{\CT}{{\cal T}}

\newcommand{\tth}{\theta}
\newcommand{\Th}{\Theta}
\newcommand{\ka}{\kappa}
\newcommand{\ph}{\varphi}
\def\VY{{V^{h}(y,j)}}
\def\VX{{V^{h}(x, i)}}

\def\para#1{\vskip .5\baselineskip\noindent{\bf #1}}

\newcommand{\blue}[1]{{\color{blue}#1}}
\newcommand{\red}[1]{{\color{red}#1}}
\newcommand{\gray}[1]{{\color{gray} #1}}

\begin{document}

%\usepackage[notcite,notref]{showkeys}
\title[Title of the presentation]{CSE-$5830$: Project Proposal}
% \subtitle{Project Structure \& Literature Review about CRF as RNN}
% \author{\textbf{Xiaohang Ma} }
% \institute{University of Connecticut \\ Department of Mathematics}
% \date{\today}

\section{Efficient Inference of fully connected CRF}
\subsection{Definition of Conditional Random Fields}
A conditional random field $(\mathbf{I}, \mathbf{X})$ is characterized by a Gibbs distribution

\[
P(\mathbf{X}|\mathbf{I}) = \frac{1}{Z(\mathbf{I})} \exp\left( - \sum_{c \in C_{\mathcal{G}}} \phi_c(\mathbf{X}_c|\mathbf{I}) \right),
\]

where $\mathcal{G} = (\mathcal{V}, \mathcal{E})$ is a graph on $\mathbf{X}$ and each clique $c$ in a set of cliques $C_{\mathcal{G}}$ in $\mathcal{G}$ induces a potential $\phi_c$. The Gibbs energy of a labeling $\mathbf{x} \in \mathcal{L}^N$ is

\[
E(\mathbf{x}|\mathbf{I}) = \sum_{c \in C_{\mathcal{G}}} \phi_c(\mathbf{x}_c|\mathbf{I}).
\]

% The maximum a posteriori (MAP) labeling of the random field is $\mathbf{x}^* = \arg \max_{\mathbf{x} \in \mathcal{L}^N} P(\mathbf{x}|\mathbf{I})$. For notational convenience, we will omit the conditioning in the rest of the paper and use $\psi_c(\mathbf{x}_c)$ to denote $\phi_c(\mathbf{x}_c|\mathbf{I})$.

In the fully connected pairwise CRF model, $\mathcal{G}$ is the complete graph on $\mathbf{X}$ and $C_{\mathcal{G}}$ is the set of all unary and pairwise cliques. The corresponding Gibbs energy i : 
\[
E(\mathbf{x}) = \sum_i \psi_u(x_i) + \sum_{i < j} \psi_p(x_i, x_j),
\]

\subsection{Mean Field Approximation of Inference of CRF}
	 Instead of computing the exact distribution $P(\mathbf{X})$, the mean field approximation would transfer the computations of CRF into a distribution $Q(\mathbf{X})$ with simpler structure.
	 The ideal choice of the approximation measure should satisfy \begin{itemize}
	 \item $Q$ can be expressed product of independent marginals, $Q(\mathbf{X}) = \prod_i Q_i(X_i)$.
	 \item $Q$ will minimize the KL-divergence $\mathcal{D}(Q\|P)$.
	\end{itemize}
	 By applying Lagrainge multiplier method, we could obtain the necessary condition of the optimal measure is a system of nonlinear equations:
	\[
	    Q_i(x_i = l) = \frac{1}{Z_i} \exp \left\{ -\psi_u(x_i) - \sum_{l' \in \mathcal{L}} \mu(l, l') \sum_{m=1}^K w^{(m)} \sum_{j \neq i} k^{(m)}(f_i, f_j) Q_j(l') \right\}.
	\]
	Recall that we have $N$ different data samples and for each sample we have $L-1$ parameters to be determinated, $i.e. \{Q_{i}(x_i = l)\}_{l = 1}^{L-1}$. 
	Thus analytically solving this system of equations may still be hard. Practically, a fixed point iteration approximation could be an efficient way to obtain the numerical solution.
	The fixed -point iteration is given by the following procedure.
	\begin{itemize}
		\item Initialize Q by \[
	  Q_i(x_i) \leftarrow \frac{1}{Z_i} \exp\left\{-\phi_u(x_i)\right\}.
		\]
		\item $\hat{Q}_i^{(m)}(l) \gets \sum_{j \neq i} k^{(m)}(f_i, f_j) Q_j(l)$\;
		\item $\hat{Q}_i(x_i) \gets \sum_{l' \in \mathcal{L}} \mu^{(m)}(x_i, l) \sum_m w^{(m)} \hat{Q}_i^{(m)}(l)$\;
		\item $Q_i(x_i) \gets \exp \left\{-\psi_u(x_i) - \hat{Q}_i(x_i) \right\}$\;
		\item 
		    normalize $Q_i(x_i)$\;
		\item repeat until convergence.
	      \end{itemize}
\section{End-to-end Learning and Inference of CRF}
For the mean field approximation inference of a CRF, we can embed all the calculation into a CNN layer,
and the iteration can be regarded as another RNN structure. Thus the inference step can be solely based on 
Neural Networks structure.
% \centering
For the training of this RNN-CRF layer, we use maximal log-likelihood strategy for the parameter updating. 
However, in this case the loss function of the Neural Networks will admite an intractable gradient.
\begin{align}
	&\mathcal{L}(\theta) = \sum_{n=1}^{N} \log P(\mathbf{x}^{(n)}|\mathbf{I}^{(n)}; \theta) = - \sum_{n=1}^{N} [E(\mathbf{x}|\mathbf{I}^{(n)};\theta) + \log Z(\mathbf{I}^{(n)}; \theta)] \\
	& \Rightarrow \nabla_{\theta} \mathcal{L}(\theta) = - \sum_{n=1}^{N} \nabla_{\theta}  [E(\mathbf{x}|\mathbf{I}^{(n)};\theta) + \log Z(\mathbf{I}^{(n)}; \theta)].
      \end{align}

      and a straightforward calculation leads to 
      \[\nabla_{\theta} \mathcal{L}(\theta) \log Z(\mathbf{I}^{(n)}; \theta) = - \mathbb{E}_{\mathbf{x}^{(n)}\thicksim P(\mathbf{x}^{(n)}|\mathbf{I}^{(n)};\theta)} \nabla_{\theta}E(\mathbf{x}|\mathbf{I}^{(n)};\theta). \]
In oreder to address this issue, we use mean field approximation again for the calculation of the marginal expectation term in the gradient. Thus, we build an end-to-end CRF model.
\end{document}

### 2024-10-09 proposal/proposal.tex
edit sec1

\documentclass[12pt]{article}

\usepackage[margin = 1in]{geometry}
\usepackage{authblk}
\usepackage[colorlinks=true,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{natbib}
\usepackage{setspace}
\onehalfspacing

\newcommand{\sx}[1]{\textcolor{red}{(SX: #1)}}

\begin{document}

\title{Proposal:
BERT-CRF Integration for Sequence Labeling on EEG Data \sx{TBD}}

\author[1]{Xiaohang Ma}
\author[2]{Shiying Xiao}
\author[2]{Xiaohui Yin}

\affil[1]{Department of Mathematics, University of Connecticut}
\affil[2]{Department of Statistics, University of Connecticut}

\date{\today}

\maketitle

\section{Project Overview}

Sequence labeling tasks are pivotal in diverse domains, especially healthcare,
where precise and timely data classification can significantly impact patient
outcomes. Medical datasets, such as physiological signals, often exhibit
complexity and heterogeneity, posing challenges for traditional modeling
techniques. Therefore, developing models capable of extracting meaningful
patterns from these datasets is crucial for accurate diagnostic and prognostic
decisions.

In recent years, deep learning has emerged as a powerful tool for sequence
labeling. Neural network architectures like long short-term memory (LSTM) and
transformer models have revolutionized sequential data processing.
Additionally, conditional random fields (CRFs) have been employed for modeling
label dependencies, demonstrating promising results.
However, while these methods have shown individual strengths, a comprehensive
approach that leverages their complementary capabilities is essential for
tackling complex, multi-modal datasets.

This project introduces a novel framework that combines deep neural networks
and probabilistic graphical models (PGMs) to enhance sequence labeling
performance. By employing bidirectional encoder representations from
transformers (BERT) for feature extraction and integrating them with CRF models,
we aim to improve sequence tagging accuracy. Furthermore, mean-field
approximation techniques are utilized for efficient inference, ensuring a
computationally efficient yet effective approach.
This integration of modern neural network architectures with CRFs offers
a robust and promising solution for complex sequence labeling tasks.

\bibliographystyle{chicago}
\bibliography{../manuscript/refs}

\end{document}

### 2024-10-09 proposal/proposal.tex
edit sec2

@@
 \usepackage{setspace}
-\onehalfspacing
+\doublespacing
 
@@
 
+\section{Literature Review}
+
+Deep neural networks, particularly recurrent neural networks (RNNs) like LSTM,
+are widely used for sequence labeling due to their ability to capture
+long-range dependencies in sequences. Bidirectional LSTM (Bi-LSTM) further
+enhances this by processing sequences in both directions, improving context
+understanding.
+CRFs, on the other hand, are commonly used for structured predictions,
+modeling dependencies between neighboring labels to ensure valid output
+sequences.
+\citet{huang2015bidirectional} pioneered the use of Bi-LSTM integrated with
+CRFs for sequence labeling tasks, demonstrating their effectiveness in
+extracting features and ensuring label consistency.
+
+However, Bi-LSTMs can still struggle to fully account for structural
+dependencies between predicted labels, leading to potential inconsistencies in
+the output sequence. Additionally, they typically require large amounts of
+labeled data and are computationally expensive, making them challenging to
+deploy in resource-constrained environments, such as hospitals.
+
+Transformer-based models, such as BERT~\citep{devlin2019bert},
+have emerged as powerful alternatives to traditional RNN-based approaches.
+Trained on massive datasets in an unsupervised manner, BERT can be fine-tuned
+for various natural language processing tasks, including sequence labeling.
+\citet{devlin2019bert} directly compared BERT to traditional models, including
+Bi-LSTM, and demonstrated its superior performance on tasks like named entity
+recognition, highlighting its ability to capture more contextual information
+than LSTM models.
+
+While CRFs are effective for modeling label dependencies, they often lack
+efficient closed-form solutions, especially for complex models beyond
+linear-chain CRFs. To address this, \citet{krahenbuhl2011efficient} introduced
+the use of mean-field approximation to solve more general CRFs in the context
+of complex prediction tasks. By minimizing the Kullback-Leibler (KL)
+divergence, mean-field approximation can be reduced to a fixed-point iteration,
+providing a faster and more efficient approach for training and inference
+while maintaining suitable accuracy. \citet{zheng2015conditional} further
+extended this technique by reformulating mean-field approximation for fully
+connected CRFs, integrating an RNN structure to capture temporal dependencies.
+
+\citet{fan2023environment}
+
+\section{Potential Data Sets}
+
+This dataset consists of electroencephalography (EEG) signals collected from
+hospital patients and provided by Harvard Medical School. It was originally
+part of a Kaggle competition that concluded in April 2024. The data includes
+spectrograms and time-series representations of EEG signals, with the goal of
+classifying six patterns of harmful brain activity, such as seizure (SZ) and
+generalized periodic discharges (GPD). For our project, we will segment the EEG
+data into time steps and use deep neural networks for feature extraction,
+followed by a fully connected CRF layer to perform the sequence labeling of
+brain activity patterns.
+
 \bibliographystyle{chicago}

### 2024-10-11 proposal/proposal.tex
proposal

@@
 \usepackage{authblk}
+\usepackage[inline]{enumitem}
 \usepackage[colorlinks=true,citecolor=blue,urlcolor=blue]{hyperref}
@@
 \usepackage{setspace}
-\doublespacing
+\onehalfspacing
 
@@
 
-\title{Proposal:
-BERT-CRF Integration for Sequence Labeling on EEG Data \sx{TBD}}
+\title{Proposal: Transformer-CRF Integration for Sequence Labeling on EEG Data}
 
@@
 labeling. Neural network architectures like long short-term memory (LSTM) and
-transformer models have revolutionized sequential data processing.
+Transformer models have revolutionized sequential data processing.
 Additionally, conditional random fields (CRFs) have been employed for modeling
@@
 and probabilistic graphical models (PGMs) to enhance sequence labeling
-performance. By employing bidirectional encoder representations from
-transformers (BERT) for feature extraction and integrating them with CRF models,
+performance. By employing Transformer~\citep{vaswani2017attention} for feature
+extraction and integrating them with CRF models,
 we aim to improve sequence tagging accuracy. Furthermore, mean-field
@@
 have emerged as powerful alternatives to traditional RNN-based approaches.
-Trained on massive datasets in an unsupervised manner, BERT can be fine-tuned
-for various natural language processing tasks, including sequence labeling.
+Trained on massive datasets in an unsupervised manner, Transformer can be
+fine-tuned for various natural language processing tasks,
+including sequence labeling.
 \citet{devlin2019bert} directly compared BERT to traditional models, including
 Bi-LSTM, and demonstrated its superior performance on tasks like named entity
-recognition, highlighting its ability to capture more contextual information
+recognition, highlighting its ability to capture more feature information
 than LSTM models.
@@
 
-\citet{fan2023environment}
+\section{Data}
 
-\section{Potential Data Sets}
+The harmful brain activity classification dataset~\citep{jing2023development}
+offers a valuable resource for studying brain health. This dataset,
+originally used in a Kaggle competition that concluded on April 8th, 2024
+(available at
+\url{https://www.kaggle.com/competitions/hms-harmful-brain-activity-classification}),
+contains electroencephalography (EEG) signals from hospital patients, collected
+by Harvard Medical School. The data comes in two primary formats:
+\begin{enumerate*}[label = (\roman*)]
+\item EEG time series data, which records 50-second windows of brain activity
+measured by 19 electrodes placed on the scalp;
+\item Spectrogram data, which provides frequency-based representations of the
+EEG signals.
+\end{enumerate*}
 
-This dataset consists of electroencephalography (EEG) signals collected from
-hospital patients and provided by Harvard Medical School. It was originally
-part of a Kaggle competition that concluded in April 2024. The data includes
-spectrograms and time-series representations of EEG signals, with the goal of
-classifying six patterns of harmful brain activity, such as seizure (SZ) and
-generalized periodic discharges (GPD). For our project, we will segment the EEG
-data into time steps and use deep neural networks for feature extraction,
-followed by a fully connected CRF layer to perform the sequence labeling of
-brain activity patterns.
+\section{Activities Plan}
+
+\begin{enumerate}
+\item \textbf{Probability graphical model}
+(Shiying Xiao and Xiaohang Ma)
+\begin{itemize}
+\item Build the CRF layer for label predition
+\item Mean field approximation inference and training of CRF model
+\end{itemize}
+\item \textbf{Deep learning methods} (Shiying Xiao, Xiaohui Yin and Xiaohang Ma)
+\begin{itemize}
+\item CNN and RNN layers for feature extraction
+\item RNN-CRF layers for label discrimination
+\end{itemize}
+\item \textbf{Manuscript writing} (Shiying Xiao, Xiaohui Yin and Xiaohang Ma)
+\item \textbf{Project presentation} (Shiying Xiao, Xiaohui Yin and Xiaohang Ma)
+\end{enumerate}
 

### 2024-10-30 progress/progress.tex
readme; progress template

--- 2024-10-09 proposal/proposal.tex
@@
 
-\newcommand{\sx}[1]{\textcolor{red}{(SX: #1)}}
-
 \begin{document}
 
-\title{Proposal:
-BERT-CRF Integration for Sequence Labeling on EEG Data \sx{TBD}}
+\title{Progress Report: Transformer-CRF Integration for Sequence Labeling on
+EEG Data}
 
@@
 
-\section{Project Overview}
+\section{Progress}
+% Overview of progress made and key accomplishments.
 
-Sequence labeling tasks are pivotal in diverse domains, especially healthcare,
-where precise and timely data classification can significantly impact patient
-outcomes. Medical datasets, such as physiological signals, often exhibit
-complexity and heterogeneity, posing challenges for traditional modeling
-techniques. Therefore, developing models capable of extracting meaningful
-patterns from these datasets is crucial for accurate diagnostic and prognostic
-decisions.
+\section{Preliminary Results}
+% Summary of initial findings and analyses.
 
-In recent years, deep learning has emerged as a powerful tool for sequence
-labeling. Neural network architectures like long short-term memory (LSTM) and
-transformer models have revolutionized sequential data processing.
-Additionally, conditional random fields (CRFs) have been employed for modeling
-label dependencies, demonstrating promising results.
-However, while these methods have shown individual strengths, a comprehensive
-approach that leverages their complementary capabilities is essential for
-tackling complex, multi-modal datasets.
+\section{Challenges}
+% Discussion of any challenges faced during the project.
 
-This project introduces a novel framework that combines deep neural networks
-and probabilistic graphical models (PGMs) to enhance sequence labeling
-performance. By employing bidirectional encoder representations from
-transformers (BERT) for feature extraction and integrating them with CRF models,
-we aim to improve sequence tagging accuracy. Furthermore, mean-field
-approximation techniques are utilized for efficient inference, ensuring a
-computationally efficient yet effective approach.
-This integration of modern neural network architectures with CRFs offers
-a robust and promising solution for complex sequence labeling tasks.
+\section{Plan Changes}
+% Outline of any adjustments made to the original plan.
 
-\bibliographystyle{chicago}
-\bibliography{../manuscript/refs}
+\section{Next Steps}
+% Description of remaining tasks and strategy for completion.
+
+\section{Updates}
+% Additional relevant updates or future directions.
 

### 2024-11-06 progress/Formula.tex
Updata the formula of our model.

\documentclass[conference]{IEEEtran}

\usepackage{amsfonts,amsmath,amssymb}
\usepackage{graphicx}
\usepackage{lmodern}
\usepackage{xcolor}
\usepackage{natbib}
\usepackage{bbm}
%notations
\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{defn}[thm]{Definition}
\newtheorem{exm}[thm]{Example}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{rem}[thm]{Remark}

\newcommand{\corref}[1]{Corollary~{\rm \ref{#1}}}
\newcommand{\defref}[1]{Definition~{\rm \ref{#1}}}
\newcommand{\exmref}[1]{Example~{\rm \ref{#1}}}
\newcommand{\lemref}[1]{Lemma~{\rm \ref{#1}}}
\newcommand{\propref}[1]{Proposition~{\rm \ref{#1}}}
\newcommand{\prpropref}[1]{Proposition~{\bf \ref{#1}}}
\newcommand{\remref}[1]{Remark~{\rm \ref{#1}}}
\newcommand{\thmref}[1]{Theorem~{\rm \ref{#1}}}
\newcommand{\ththmref}[1]{Theorem~{\bf \ref{#1}}}

\renewcommand{\thesection}{\arabic{section}}
\renewcommand{\theequation}{\thesection.\arabic{equation}}
\newcommand{\beq}[1]{\begin{equation} \label{#1}}
	\newcommand{\eeq}{\end{equation}}
\newcommand{\bea}{\bed\begin{array}{rl}}
	\newcommand{\eea}{\end{array}\eed}
\newcommand{\bed}{\begin{displaymath}}
	\newcommand{\eed}{\end{displaymath}}
\newcommand{\barray}{\begin{array}{ll}}
	\newcommand{\earray}{\end{array}}

\newcommand{\aad}{&\disp}
\newcommand{\ad}{&\!\!\!\disp}
\newcommand{\al}{\alpha}
\newcommand{\be}{\beta}
\newcommand{\disp}{\displaystyle}
\newcommand{\dl}{\delta}
\newcommand{\E}{{\mathbb E}}
\newcommand{\e}{\varepsilon}
\newcommand{\ee}{\Delta}
\newcommand{\la}{\lambda}
\newcommand{\PP}{{\mathbb P}}
\newcommand{\nd}{\noindent}
\newcommand{\sg}{\sigma}

\newcommand{\NM}{{\rm NM}}
\newcommand{\AM}{{\rm AM}}

\newcommand{\A}{{\cal A}}
\newcommand{\F}{{\cal F}}
\newcommand{\I}{{\cal I}}
\newcommand{\M}{{\cal M}}
\newcommand{\U}{{\cal U}}
\newcommand{\X}{{\cal X}}
\newcommand{\II}{{\cal I}}
\newcommand{\HH}{{\cal H}}

\newcommand{\bdd}{\hspace*{-0.08in}{\bf.}\hspace*{0.05in}}

\newcommand{\lbar}{\overline}
\newcommand{\wdt}{\widetilde}
\newcommand{\wdh}{\widehat}
\newcommand{\diag}{{\rm diag}}
\def\th{\theta}
\def\para#1{\vskip 0.4\baselineskip\noindent{\bf #1}}
\def\paraa#1{\vskip 0.4\baselineskip\noindent{\sf #1}}

\def\rr{{\Bbb R}}
%\def\({\left(}
%\def\){\right)}
\def\one{{\hbox{1{\kern -0.35em}1}}}
\def\op{{\cal L}}

\def\gg{{\cal G}}
\def\sth{\sum^2_{\th=1}}

\begin{document}

\title{CSE 5830 Course Project Proposal}

\author{\IEEEauthorblockN{Xiaohang Ma}
\IEEEauthorblockA{\textit{Department of Mathematics} \\
\textit{University of Connecticut}\\
Storrs, CT, USA \\
xiaohang.ma@uconn.edu}
\and
\IEEEauthorblockN{Shiying Xiao}
\IEEEauthorblockA{\textit{Department of Statistics} \\
\textit{University of Connecticut}\\
Storrs, CT, USA \\
shiying.xiao@uconn.edu}
\and
\IEEEauthorblockN{Xiaohui Yin}
\IEEEauthorblockA{\textit{Department of Statistics} \\
\textit{University of Connecticut}\\
Storrs, CT, USA \\
xiaohui.yin@uconn.edu}
}

\maketitle

\begin{abstract}
In this project, we will integrate powerful deep neural networks and
statistical methods with a probabilistic graphical model (PGM) to effectively
tackle sequence labeling tasks on a heterogeneous medical care dataset.
\end{abstract}

\begin{IEEEkeywords}
probabilistic graphical model
\end{IEEEkeywords}

\section{Introduction}

The

\section{CNN-HCRF Model}

The classical HCRF model can be formulated as
\begin{align}
	\mathbb{P}(y| \mathbf{x}; \theta) & = \sum_{\mathbf{h} \in H^{N}} \mathbb{P}(y, \mathbf{h}| \mathbf{x}; \theta) \\
	& = \frac{\sum_{\mathbf{h}\in H^{N}} \exp(\Phi(y, \mathbf{h}, \mathbf{x}; \theta))}
	{\sum_{y^{'}} \sum_{\mathbf{h}\in H^{N}} \exp(\Phi(y^{'}, \mathbf{h}, \mathbf{x}; \theta))}.
\end{align}

Different to the design of most existing HCRF models, which they assign potentials based on the features predefined from the input data \(\mathbf{I}\).
% \( \mathbb{P}(y, \mathbf{h}| \mathbf{x};\theta) \).
%  Here, we consider a generative model that constructing
% \(\PP (\mathbf{h}|\mathbf{x}; \theta)\) through a CRF framework then using an unary potential for , 
Our HCRF replace \(I\) with features learned by CNN, \(i.e. \ \mathbf{F}(\mathbf{I})\). 
Thus the new potentials become:
\[
	\Phi(y, \mathbf{h}, \mathbf{x}; \theta) = \Phi(y, \mathbf{h}, \mathbf{F}(\mathbf{I}); \theta).
	% \PP (y, \mathbf{h}| \mathbf{x}; \theta) = \PP (y | \mathbf{h},  \mathbf{x}; \theta)
	% \cdot \PP ( \mathbf{h}| \mathbf{x}; \theta).
\]
Thus, we design the potential of our HCRF model as.
\begin{align}
	\Phi(y, \mathbf{h}, \mathbf{x}; \theta) & = \underbrace{\sum_{j \in \nu} \phi(h_{j}, x_{j}; \omega) +
	\sum_{i \neq j} \psi(h_{i}, h_{j}, x_{i}, x_{j}; \eta) }_{
		% \propto \log\PP ( \mathbf{h}| \mathbf{x}; \theta)
		\textrm{Measures  log-likelihood \(\log\PP ( \mathbf{h}| \mathbf{x}; \theta)\)}
		}\\
	& + \underbrace{ \sum_{j \in \nu}\varphi
	(y, h_{j}, x_{j}; \delta) + \vartheta(y, \mathbf{x};
	\varpi)}_{
		% \propto \log \PP (y | \mathbf{h},  \mathbf{x}; \theta)
		\textrm{Measures log-likelihood \(\log \PP (y | \mathbf{h},  \mathbf{x}; \theta)\)}
		}
\end{align}

\noindent
{\textbf{Unary Potential}} \(\phi(h_{j}, x_{j}; \omega)\): measures the likelihood of the local feature
\(x_{j}\) is assigned as the hidden attention state \(h_{j}\).
We generate the likelihood through a softmax operation on \(\mathbf{F}(\mathbf{x})\)
the feature maps of CNN with \((N, C_{\textrm{out}})\) dimensions,
by which the potential of each pixel \(x_j\) assigned as each
state of hidden state set \(\{0, 1, \dots, 5\}\)
is drawn and represented by a matrix with \((N, 2)\) dimension.

The parameter \(\omega\) is learned by our end-to-end CNN-Unet training structure.

\noindent
{\textbf{Unary Potential}} \(\varphi(y, h_{j}; \delta)\) measures the compatibility between global class label \(y\)
and the hidden local state \(h_{j}\).
This potential is parametrized as:
\[
	\varphi(y, h_{j};\delta) = \sum_{a \in Y} \sum_{b \in H} \delta_{a, b}
	\cdot \mathbbm{1}(y = a) \cdot \mathbbm{1}(h_{j} = b).
\]

\noindent
{\textbf{Binary Potential}} \(\psi(h_{i}, h_{j}, x_{i}, x_{j}; \eta)\) 
balances the hidden label compatibility of the neighboring pixels on the feature image \(\mathbf{F}(\mathbf{I})\).
We define this potential through a common contrast-sensitive two-kernal potentials see([Krähenbühl P, Koltun V. 2011], [Chen S, Gamechi Z S, Dubost F, et al. 2022]):
\begin{eqnarray}
	 \psi(h_{i}, h_{j}, x_{i}, x_{j};  \eta) = \mu(h_{i}, h_{j}) \Big[ \omega_{1} 
	\exp ( - \frac{\left\lvert p_{i} - p_{j} \right\rvert^{2} }{2 \eta_{\alpha}^{2}}\\
	 - \frac{\left\lvert \mathbf{F}_{i}(\mathbf{I}) - \mathbf{F}_{j}(\mathbf{I}) \right\rvert^{2} }{2 \eta_{\beta}^{2}}) 
	 + \omega_{2}\exp ( - \frac{\left\lvert p_{i} - p_{j} \right\rvert^{2}}{2 \eta_{\gamma}^{2}})\Big].
\end{eqnarray}

\section{Variational Inference of HCRF}
\subsection{\textbf{ELBO Under Mean-Filed Variational Family}}
The existence of the hidden variable \(\mathbf{h}\) makes our model more interpretable. However, since We will train our model through maximal likelihood method, 
 introducing the hidden variable will make precisely computation the summation \(\sum_{\mathbf{h} \in H^{N}} \mathbb{P}(y, \mathbf{h}| \mathbf{x}; \theta)\) intractable.
To address this issue, we make our computation more feasible through a mean-field variational inference framework, which enables our model efficiently estimating the log-likelihood \(\log\PP(y| \mathbf{x}; \theta).\)

Firstly, we introduce the mean-field varepsilon family approximation of \(\log\PP (\mathbf{h}|y, \mathbf{x};\theta)\) as:
\[
	\mathbb{Q}(\mathbf{h}) = \prod_{i=1}^{N} q_{i}(h_{i}).
\]
Then the evidence lower bound(ELBO) under the mean-field family is given by
\begin{align}
	\textrm{ELBO}(\mathbb{Q}) & = \E_{\mathbb{Q}(\mathbf{h})}\log \PP (y, \mathbf{h}| \mathbf{x}; \theta) - \E_{\mathbb{Q}(\mathbf{h})}\log q(\mathbf{h}) \\
	& = \sum_{i=1}^{N} \E_{q_{i}(h_{i})} [\phi(h_{j}, x_{j}; \omega) + \varphi
	(y, h_{j}, x_{j}; \delta) ]  \\ 
	& + \sum_{i=1}^{N} \sum_{j=1}^{N} \sum_{l,l^{'}} \mathbbm{1}_{\{i \neq j\}} \mathbb{Q}(h_{i} = l)\mathbb{Q}(h_{j} = l^{'})\mu(l, l^{'}) \\
	& 
	\Big[ \omega_{1} 
	\exp ( - \frac{\left\lvert p_{i} - p_{j} \right\rvert^{2} }{2 \eta_{\alpha}^{2}}
	 - \frac{\left\lvert \mathbf{F}_{i}(\mathbf{I}) - \mathbf{F}_{j}(\mathbf{I}) \right\rvert^{2} }{2 \eta_{\beta}^{2}}) \\
	 & + \omega_{2}\exp ( - \frac{\left\lvert p_{i} - p_{j} \right\rvert^{2}}{2 \eta_{\gamma}^{2}})\Big] \\ 
	&  - \sum_{i=1}^{N}\E_{q_{i}}\log q_{i}(h_{i}) + \vartheta(y, \mathbf{x}; \varpi).
\end{align}
\subsection{\textbf{Updating Variational Family}}
By the coordinate ascent variational inference(CAVI) algorithm, we can obtain that fix \(q_{j}(h_{j}), \forall j \neq i\), the optimal \(q_{i}(h_{i})\) that maximizes \(\textrm{ELBO}(q_{i}(h_{i}))\) is given by
\begin{align}
	\mathbb{Q}(h_{i}) \propto \exp\{ \E_{\mathbb{Q}({\mathbf{h}_{-i}})} \log \PP (y, h_{i}, \mathbf{h}_{-i}|\mathbf{x}; \theta)\}.
\end{align}

Thus, we can get that
\begin{align}
	q_{i}(l) &= \frac{1}{Z_{i}} \exp \Big\{\phi(h_{j}, x_{j}; \omega) + \varphi
	(y, h_{j}, x_{j}; \delta) \\
	& +  \sum_{j\neq i } \sum_{l^{'}}  q(l^{'})\mu(l, l^{'}) 
	\Big[ \omega_{1} 
	\exp ( - \frac{\left\lvert p_{i} - p_{j} \right\rvert^{2} }{2 \eta_{\alpha}^{2}}\\
	&
	 - \frac{\left\lvert \mathbf{F}_{i}(\mathbf{I}) - \mathbf{F}_{j}(\mathbf{I}) \right\rvert^{2} }{2 \eta_{\beta}^{2}})  + \omega_{2}\exp ( - \frac{\left\lvert p_{i} - p_{j} \right\rvert^{2}}{2 \eta_{\gamma}^{2}})\Big]
	 \Big\}
\end{align}

\subsection{\textbf{Computational Complexity Analysis}}
\noindent
\begin{itemize}
	\item {\textbf{Updating Variational Family}} \(\{q_{i}(h_{i})\}\): The computationally expensive part of the fixed point iteration of \(q_{i}(h_i)\) comes
	 from the message passing, \(i.e\). the convolution of \(\mathbb{Q}(\mathbf{h})\) and gaussian kernal in \(\psi(h_{i}, h_{j}, x_{i}, x_{j}; \eta)\). We need \(O(N^{2})\) runtime for N pixels when evaluating this term precisely.
	 To make the computation more feasible, a truncated gaussian kernal will be employed, which is only supported on spaces within certain proportion of the untruncated standard deviation. Then the approximate message passing is linear in the number of pixels \(N\).
	\item {\textbf{Computation of the ELBO}}: The summation over the product of \(q_{i}(h_{i})q_{j}(h_{j})\) and the gaussian kernal in \(\psi(h_{i}, h_{j}, x_{i}, x_{j}; \eta)\) also requires a \(O(N^{2})\) complexity in time. 
	Owes to a similar truncated gaussian kernal approxiamtion, the runtime can be reduced to \(O(N)\).
\end{itemize}

\citep{huang2015bidirectional}

\bibliographystyle{plain}
\bibliography{refs}

\end{document}

### 2024-11-08 progress/progress.tex
progress draft

@@
-\documentclass[12pt]{article}
+\documentclass[10pt]{article}
 
-\usepackage[margin = 1in]{geometry}
+\usepackage[margin = .8in]{geometry}
+\usepackage{amsfonts,amsmath,amssymb}
 \usepackage{authblk}
+\usepackage{bbm}
 \usepackage[inline]{enumitem}
+\usepackage{graphicx}
+\graphicspath{{./}{../image/}}
 \usepackage[colorlinks=true,citecolor=blue,urlcolor=blue]{hyperref}
@@
 \usepackage{setspace}
-\onehalfspacing
+\setstretch{1.2}
+
+\newcommand{\EE}{{\mathbb{E}}}
+\newcommand{\PP}{{\mathbb{P}}}
+\newcommand{\QQ}{{\mathbb{Q}}}
+\newcommand{\Fb}{\mathbf{F}}
+\newcommand{\Ib}{\mathbf{I}}
+\newcommand{\hb}{\mathbf{h}}
+\newcommand{\xb}{\mathbf{x}}
+\newcommand{\one}{\mathbbm{1}}
+\newcommand{\Ocal}{\mathcal{O}}
 
@@
 
+The project is progressing according to schedule. To date, we have made
+significant strides in several key areas:
+\begin{enumerate}
+\item Model Development:
+We have initiated development of a convolutional neural network (CNN) combined
+with a hidden conditional random field (HCRF). This hybrid approach aims to
+create a robust sequence labeling framework capable of capturing both local
+and global dependencies within the data sequences.
+\item Variational Inference Implementation:
+To efficiently approximate the posterior distributions in our complex model,
+we have integrated variational inference into the HCRF component.
+This technique is particularly effective in handling the computational
+challenges posed by the dataset's heterogeneity.
+\item Data Visualization:
+We have successfully visualized key features of the dataset.
+\end{enumerate}
+%This progress has provided a foundation for the next
+%steps of our research.
+
 \section{Preliminary Results}
 % Summary of initial findings and analyses.
+
+The potential of our HCRF model:
+\begin{equation*}
+\Phi(y, \hb, \xb; \theta) = \underbrace{
+\sum_{j \in \nu} \phi(h_j, x_j; \omega) +
+\sum_{i \neq j} \psi(h_i, h_j, x_i, x_j; \eta)}_{
+% \propto \log\PP ( \hb \vert \xb; \theta)
+\textrm{Measures log-likelihood $\log\PP ( \hb \vert \xb; \theta)$}}
++ \underbrace{
+\sum_{j \in \nu} \varphi(y, h_j, x_j; \delta) +
+\vartheta(y, \xb; \varpi)}_{
+% \propto \log \PP (y | \hb,  \xb; \theta)
+\textrm{Measures log-likelihood $\log \PP (y \vert \hb, \xb; \theta)$}}
+\end{equation*}
+
+\begin{itemize}
+\item \textbf{Unary Potential} $\phi(h_j, x_j; \omega)$ measures
+the likelihood of the local feature $x_j$ is assigned as
+the hidden attention state $h_j$.
+We generate the likelihood through a softmax operation on $\Fb(\xb)$
+the feature maps of CNN with $(N, C_{\textrm{out}})$ dimensions,
+by which the potential of each pixel $x_j$ assigned as each state of
+hidden state set $\{0, 1, \dots, 5\}$
+is drawn and represented by a matrix with $(N, 2)$ dimension.
+\item The parameter $\omega$ is learned by our end-to-end CNN-Unet training 
+structure.
+\item \textbf{Unary Potential} $\varphi(y, h_j; \delta)$ measures
+the compatibility between global class label $y$ and the hidden local
+state $h_j$. This potential is parametrized as:
+\begin{align*}
+\varphi(y, h_{j}; \delta) = \sum_{a \in Y} \sum_{b \in H} \delta_{a, b}
+\cdot \one(y = a) \cdot \one(h_j = b).
+\end{align*}
+\item \textbf{Binary Potential} $\psi(h_i, h_j, x_i, x_j; \eta)$
+balances the hidden label compatibility of the neighboring pixels on the 
+feature image $\Fb(\Ib)$.
+We define this potential through a common contrast-sensitive two-kernal 
+potentials (see~\citet{krahenbuhl2011efficient} and \citet{chen2022end}):
+\begin{equation*}
+\psi(h_i, h_j, x_i, x_j; \eta) = \mu(h_i, h_j) \Bigg[
+\omega_1 \exp \left(
+-\frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2}-
+-\frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib) \right\rvert^2}{2\eta_\beta^2} 
+\right) + \omega_2 \exp \left(
+- \frac{\left\lvert p_i - p_j \right\rvert^2}{2 \eta_\gamma^2}
+\right)
+\Bigg]
+\end{equation*}
+\end{itemize}
+
+While the models are still under development and optimization,
+initial results are promising. For instance, Figure~\ref{fig:SEIZURE_3}
+illustrates a sample prediction with ground truth 94.\% Seizure and 2.9\% LPD.
+Our model predicts 88.4\% Seizure and 1.6\% LPD.
+Yellow highlights indicate the regions where the model focused its attention.
+These preliminary results show trends that are consistent with our initial
+hypotheses. The visualizations have proven useful in pinpointing areas for
+further investigation, and the derived formulas appear to be robust and
+aligned with the expected model outcomes. Further analysis will refine
+these findings and explore additional dimensions of data.
+
+\begin{figure}[tbp]
+\centering
+\includegraphics[width=.8\textwidth]{grad_cam_SEIZURE_3}
+\caption{An example of data visualization. Areas highlighted in yellow
+indicate regions where the model focused its attention during this specific
+prediction.}
+\label{fig:SEIZURE_3}
+\end{figure}
 
@@
 
+We encountered minor challenges related to data handling and computational
+limitations. In particular, the complexity of certain calculations required
+additional processing time, which slightly delayed our progress.
+We are currently exploring ways to optimize these computations.
+
 \section{Plan Changes}
 % Outline of any adjustments made to the original plan.
+
+At this stage, we have not made significant changes to our original plan.
+However, due to the complexity of the dataset and model, we may consider
+slightly adjust the project timeline to allocate additional time for
+model fine-tuning.
 
@@
 
+The remaining tasks include:
+\begin{enumerate*}[label = (\roman*)]
+\item Expanding and refining the visualizations based on updated data.
+\item Finalizing the derivation and validation of all essential formulas.
+\item Conducting a comprehensive analysis with the refined model.
+\item Preparing the final report and presentation materials.
+\end{enumerate*}
+
 \section{Updates}
@@
 
+We will provide regular updates on model performance, challenges encountered,
+and any significant changes to the project plan.
+
+%\section{CNN-HCRF Model}
+%
+%The classical hidden conditional random field (HCRF) model can be formulated as
+%\begin{align}
+%\PP(y \vert \xb; \theta) &= \sum_{\hb \in H^N}
+%\mathbb{P}(y, \hb \vert \xb; \theta) \\
+%&= \frac{\sum_{\hb \in H^N} \exp\left(\Phi(y, \hb, \xb; \theta)\right)}
+%{\sum_{y^\prime} \sum_{\hb \in H^N}
+%	\exp\left(\Phi(y^\prime, \hb, \xb; \theta)\right)}.
+%\end{align}
+%
+%
+%Different to the design of most existing HCRF models, which they assign
+%potentials based on the features predefined from the input data $\Ib$.
+%% \( \PP(y, \hb \vert \xb; \theta) \).
+%%  Here, we consider a generative model that constructing
+%% \( \PP(\hb \vert \xb; \theta) \) through a CRF framework then using an 
+%%unary potential for , 
+%Our HCRF replace $I$ with features learned by CNN, \(i.e. \ 
+%\Fb(\Ib)\). 
+%Thus the new potentials become:
+%\begin{align*}
+%\Phi(y, \hb, \xb; \theta) = \Phi(y, \hb, \Fb(\Ib); \theta).
+%%\PP (y, \hb \vert \xb; \theta) = \PP (y \vert \hb, \xb; \theta)
+%%\cdot \PP ( \hb \vert \xb; \theta).
+%\end{align*}
+%Thus, we design the potential of our HCRF model as.
+%\begin{align}
+%\Phi(y, \hb, \xb; \theta) & = \underbrace{
+%\sum_{j \in \nu} \phi(h_j, x_j; \omega) +
+%\sum_{i \neq j} \psi(h_i, h_j, x_i, x_j; \eta)}_{
+%% \propto \log\PP ( \hb \vert \xb; \theta)
+%\textrm{Measures log-likelihood $\log\PP ( \hb \vert \xb; \theta)$}} \\
+%& + \underbrace{
+%\sum_{j \in \nu} \varphi(y, h_j, x_j; \delta) +
+%\vartheta(y, \xb; \varpi)}_{
+%% \propto \log \PP (y | \hb,  \xb; \theta)
+%\textrm{Measures log-likelihood $\log \PP (y \vert \hb, \xb; \theta)$}}
+%\end{align}
+%
+%
+%{\textbf{Unary Potential}} $\phi(h_j, x_j; \omega)$: measures
+%the likelihood of the local feature $x_j$ is assigned as
+%the hidden attention state $h_j$.
+%We generate the likelihood through a softmax operation on $\Fb(\xb)$
+%the feature maps of CNN with $(N, C_{\textrm{out}})$ dimensions,
+%by which the potential of each pixel $x_j$ assigned as each state of
+%hidden state set $\{0, 1, \dots, 5\}$
+%is drawn and represented by a matrix with $(N, 2)$ dimension.
+%
+%The parameter $\omega$ is learned by our end-to-end CNN-Unet training 
+%structure.
+%
+%
+%{\textbf{Unary Potential}} $\varphi(y, h_j; \delta)$ measures
+%the compatibility between global class label $y$ and the hidden local
+%state $h_j$. This potential is parametrized as:
+%\begin{align*}
+%\varphi(y, h_{j}; \delta) = \sum_{a \in Y} \sum_{b \in H} \delta_{a, b}
+%\cdot \one(y = a) \cdot \one(h_j = b).
+%\end{align*}
+%
+%
+%{\textbf{Binary Potential}} $\psi(h_i, h_j, x_i, x_j; \eta)$
+%balances the hidden label compatibility of the neighboring pixels on the 
+%feature image $\Fb(\Ib)$.
+%We define this potential through a common contrast-sensitive two-kernal 
+%potentials see~\citet{krahenbuhl2011efficient} and \citep{chen2022end}:
+%
+%
+%\begin{equation}
+%\begin{split}
+%\psi(h_i, h_j, x_i, x_j; \eta) = \mu(h_i, h_j) \Bigg[
+%& \omega_1 \exp \left(
+%-\frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2}-
+%-\frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib) \right\rvert^2}{2\eta_\beta^2} 
+%\right) \\
+%& + \omega_2 \exp \left(
+%- \frac{\left\lvert p_i - p_j \right\rvert^2}{2 \eta_\gamma^2}
+%\right)
+%\Bigg]
+%\end{split}
+%\end{equation}
+%
+%
+%
+%\section{Variational Inference of HCRF}
+%
+%
+%\subsection{ELBO Under Mean-Filed Variational Family}
+%
+%
+%The existence of the hidden variable $\hb$ makes our model more 
+%interpretable. However, since We will train our model through maximal 
+%likelihood method, 
+%introducing the hidden variable will make precisely computation the summation 
+%$\sum_{\hb \in H^N} \PP(y, \hb \vert \xb; \theta)$ intractable.
+%To address this issue, we make our computation more feasible through a 
+%mean-field variational inference framework, which enables our model 
+%efficiently 
+%estimating the log-likelihood $\log\PP(y \vert \xb; \theta)$.
+%
+%
+%Firstly, we introduce the mean-field varepsilon family approximation of 
+%$\log\PP(\hb \vert y, \xb; \theta)$ as:
+%\begin{align*}
+%\QQ(\hb) = \prod_{i=1}^N q_i(h_i).
+%\end{align*}
+%Then the evidence lower bound(ELBO) under the mean-field family is given by
+%
+%
+%\begin{equation}
+%\begin{split}
+%\textrm{ELBO}(\QQ) &= \EE_{\QQ(\hb)} \log \PP (y, \hb \vert \xb; \theta)
+%- \EE_{\QQ(\hb)} \log q(\hb) \\
+%& = \sum_{i=1}^N \EE_{q_i(h_i)}
+%\left[ \phi(h_j, x_j; \omega) + \varphi(y, h_j, x_j; \delta) \right] \\
+%&\quad + \sum_{i=1}^N \sum_{j=1}^N \sum_{l,l^\prime} \one_{\{i \neq j\}} 
+%\QQ(h_i = l) \QQ(h_j = l^\prime) \mu(l, l^\prime) \\
+%&\qquad \Bigg[ \omega_1 \exp \left(
+%- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2}
+%- \frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib) \right\rvert^2}{2\eta_\beta^2}
+%\right) \\
+%&\qquad + \omega_2 \exp \left(
+%- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\gamma^2}
+%\right) \Bigg] \\
+%&\quad - \sum_{i=1}^N \EE_{q_i} \log q_i(h_i) + \vartheta(y, \xb; \varpi).
+%\end{split}
+%\end{equation}
+%
+%
+%\subsection{Updating Variational Family}
+%
+%
+%By the coordinate ascent variational inference(CAVI) algorithm, we can obtain 
+%that fix \(q_{j}(h_{j}), \forall j \neq i\), the optimal \(q_{i}(h_{i})\) that 
+%maximizes \(\textrm{ELBO}(q_{i}(h_{i}))\) is given by
+%\begin{align}
+%	\mathbb{Q}(h_{i}) \propto \exp\{ \EE_{\mathbb{Q}({\hb_{-i}})} \log 
+%	\PP (y, h_{i}, \hb_{-i}|\xb; \theta)\}.
+%\end{align}
+%
+%Thus, we can get that
+%\begin{align}
+%	q_{i}(l) &= \frac{1}{Z_{i}} \exp \Big\{\phi(h_{j}, x_{j}; \omega) + \varphi
+%	(y, h_{j}, x_{j}; \delta) \\
+%	& +  \sum_{j\neq i } \sum_{l^{'}}  q(l^{'})\mu(l, l^{'}) 
+%	\Big[ \omega_{1} 
+%	\exp ( - \frac{\left\lvert p_{i} - p_{j} \right\rvert^{2} }{2 
+%	\eta_{\alpha}^{2}}\\
+%	&
+%	- \frac{\left\lvert \Fb_{i}(\Ib) - \Fb_{j}(\Ib) 
+%	\right\rvert^{2} }{2 \eta_{\beta}^{2}})  + \omega_{2}\exp ( - 
+%	\frac{\left\lvert p_{i} - p_{j} \right\rvert^{2}}{2 \eta_{\gamma}^{2}})\Big]
+%	\Big\}
+%\end{align}
+%
+%
+%\subsection{Computational Complexity Analysis}
+%
+%
+%\begin{itemize}
+%\item {\textbf{Updating Variational Family}} $\{q_i(h_i)\}$:
+%The computationally expensive part of the fixed point iteration of $q_i(h_i)$
+%comes from the message passing, \(i.e\). the convolution of $\QQ(\hb)$ and
+%gaussian kernal in $\psi(h_i, h_j, x_i, x_j; \eta)$. We need $\Ocal(N^2)$
+%runtime for $N$ pixels when evaluating this term precisely.
+%To make the computation more feasible, a truncated gaussian kernal will be
+%employed, which is only supported on spaces within certain proportion of
+%the untruncated standard deviation. Then the approximate message passing is
+%linear in the number of pixels $N$.
+%\item {\textbf{Computation of the ELBO}}: The summation over the product of 
+%$q_i(h_i)q_j(h_j)$ and the gaussian kernal in $\psi(h_i, h_j, x_i, x_j; \eta)$
+%also requires a $\Ocal(N^2)$ complexity in time. 
+%Owes to a similar truncated gaussian kernal approxiamtion, the runtime can
+%be reduced to $\Ocal(N)$.
+%\end{itemize}
+
+\bibliography{../manuscript/refs}
+\bibliographystyle{chicago}
+
 \end{document}

### 2024-11-09 manuscript/main.tex
add ELBO

@@
 
+%% keep it clean; only include those you need
 \usepackage{amsfonts,amsmath,amssymb}
+\usepackage{bbm}
 \usepackage{graphicx}
+\graphicspath{{./}{../image/}}
+\usepackage[colorlinks=true,citecolor=blue,urlcolor=blue]{hyperref}
 \usepackage{xcolor}
-\usepackage{natbib}
 
-%notations
-\newtheorem{thm}{Theorem}[section]
-\newtheorem{cor}[thm]{Corollary}
-\newtheorem{defn}[thm]{Definition}
-\newtheorem{exm}[thm]{Example}
-\newtheorem{lem}[thm]{Lemma}
-\newtheorem{prop}[thm]{Proposition}
-\newtheorem{rem}[thm]{Remark}
-
-\newcommand{\corref}[1]{Corollary~{\rm \ref{#1}}}
-\newcommand{\defref}[1]{Definition~{\rm \ref{#1}}}
-\newcommand{\exmref}[1]{Example~{\rm \ref{#1}}}
-\newcommand{\lemref}[1]{Lemma~{\rm \ref{#1}}}
-\newcommand{\propref}[1]{Proposition~{\rm \ref{#1}}}
-\newcommand{\prpropref}[1]{Proposition~{\bf \ref{#1}}}
-\newcommand{\remref}[1]{Remark~{\rm \ref{#1}}}
-\newcommand{\thmref}[1]{Theorem~{\rm \ref{#1}}}
-\newcommand{\ththmref}[1]{Theorem~{\bf \ref{#1}}}
-
-\renewcommand{\thesection}{\arabic{section}}
-\renewcommand{\theequation}{\thesection.\arabic{equation}}
-\newcommand{\beq}[1]{\begin{equation} \label{#1}}
-	\newcommand{\eeq}{\end{equation}}
-\newcommand{\bea}{\bed\begin{array}{rl}}
-	\newcommand{\eea}{\end{array}\eed}
-\newcommand{\bed}{\begin{displaymath}}
-	\newcommand{\eed}{\end{displaymath}}
-\newcommand{\barray}{\begin{array}{ll}}
-	\newcommand{\earray}{\end{array}}
-
-\newcommand{\aad}{&\disp}
-\newcommand{\ad}{&\!\!\!\disp}
-\newcommand{\al}{\alpha}
-\newcommand{\be}{\beta}
-\newcommand{\disp}{\displaystyle}
-\newcommand{\dl}{\delta}
-\newcommand{\E}{{\mathbb E}}
-\newcommand{\e}{\varepsilon}
-\newcommand{\ee}{\Delta}
-\newcommand{\la}{\lambda}
-\newcommand{\PP}{{\mathbb P}}
-\newcommand{\nd}{\noindent}
-\newcommand{\sg}{\sigma}
-
-\newcommand{\NM}{{\rm NM}}
-\newcommand{\AM}{{\rm AM}}
-
-\newcommand{\A}{{\cal A}}
-\newcommand{\F}{{\cal F}}
-\newcommand{\I}{{\cal I}}
-\newcommand{\M}{{\cal M}}
-\newcommand{\U}{{\cal U}}
-\newcommand{\X}{{\cal X}}
-\newcommand{\II}{{\cal I}}
-\newcommand{\HH}{{\cal H}}
-
-\newcommand{\bdd}{\hspace*{-0.08in}{\bf.}\hspace*{0.05in}}
-
-\newcommand{\lbar}{\overline}
-\newcommand{\wdt}{\widetilde}
-\newcommand{\wdh}{\widehat}
-\newcommand{\diag}{{\rm diag}}
-\def\th{\theta}
-\def\para#1{\vskip 0.4\baselineskip\noindent{\bf #1}}
-\def\paraa#1{\vskip 0.4\baselineskip\noindent{\sf #1}}
-
-\def\rr{{\Bbb R}}
-\def\({\left(}
-\def\){\right)}
-\def\one{{\hbox{1{\kern -0.35em}1}}}
-\def\op{{\cal L}}
-
-\def\gg{{\cal G}}
-\def\sth{\sum^2_{\th=1}}
+%% notations; only define those you use frequently
+\newcommand{\EE}{{\mathbb{E}}}
+\newcommand{\PP}{{\mathbb{P}}}
+\newcommand{\QQ}{{\mathbb{Q}}}
+\newcommand{\Fb}{\mathbf{F}}
+\newcommand{\Ib}{\mathbf{I}}
+\newcommand{\hb}{\mathbf{h}}
+\newcommand{\xb}{\mathbf{x}}
+\newcommand{\one}{\mathbbm{1}}
+\newcommand{\Ocal}{\mathcal{O}}
 
@@
 
-\title{CSE 5830 Course Project Proposal}
+\title{\LARGE Transformer-CRF Integration for Sequence Labeling on EEG Data}
 
@@
 \begin{abstract}
+
 In this project, we will integrate powerful deep neural networks and
@@
 tackle sequence labeling tasks on a heterogeneous medical care dataset.
+
 \end{abstract}
@@
 \begin{IEEEkeywords}
+
 probabilistic graphical model
+
 \end{IEEEkeywords}
 
-\section{Introduction}
+\section{CNN-HCRF Model}
 
-\citep{huang2015bidirectional}
+The classical hidden conditional random field (HCRF) model can be formulated as
+\begin{equation}
+\begin{split}
+\PP(y \vert \xb; \theta) &= \sum_{\hb \in H^N}
+\mathbb{P}(y, \hb \vert \xb; \theta) \\
+&= \frac{\sum_{\hb \in H^N} \exp\left(\Phi(y, \hb, \xb; \theta)\right)}
+{\sum_{y^\prime} \sum_{\hb \in H^N}
+\exp\left(\Phi(y^\prime, \hb, \xb; \theta)\right)}.
+\end{split}
+\end{equation}
 
-\bibliographystyle{plain}
+Different to the design of most existing HCRF models, which they assign
+potentials based on the features predefined from the input data $\Ib$.
+%$\PP(y, \hb \vert \xb; \theta)$.
+%Here, we consider a generative model that constructing
+%$\PP(\hb \vert \xb; \theta)$ through a CRF framework then using an 
+%unary potential for , 
+Our HCRF replace $I$ with features learned by CNN, i.e. $\Fb(\Ib)$. 
+Thus the new potentials become:
+\begin{equation*}
+\Phi(y, \hb, \xb; \theta) = \Phi(y, \hb, \Fb(\Ib); \theta).
+%\PP(y, \hb \vert \xb; \theta) = \PP(y \vert \hb, \xb; \theta)
+%\cdot \PP(\hb \vert \xb; \theta).
+\end{equation*}
+Thus, we design the potential of our HCRF model as.
+\begin{equation}
+\begin{split}
+\Phi(y, \hb, \xb; \theta) &= \underbrace{
+\sum_{j \in \nu} \phi(h_j, x_j; \omega)
++ \sum_{i \neq j} \psi(h_i, h_j, x_i, x_j; \eta)}_{
+% \propto \log\PP(\hb \vert \xb; \theta)
+\textrm{Measures log-likelihood $\log\PP(\hb \vert \xb; \theta)$}} \\
+&+ \underbrace{
+\sum_{j \in \nu} \varphi(y, h_j, x_j; \delta) + \vartheta(y, \xb; \varpi)}_{
+% \propto \log\PP(y | \hb,  \xb; \theta)
+\textrm{Measures log-likelihood $\log\PP(y \vert \hb, \xb; \theta)$}}
+\end{split}
+\end{equation}
+
+{\textbf{Unary Potential}}
+$\phi(h_j, x_j; \omega)$ measures the likelihood of the local feature
+$x_j$ is assigned as the hidden attention state $h_j$.
+We generate the likelihood through a softmax operation on $\Fb(\xb)$
+the feature maps of CNN with $(N, C_{\textrm{out}})$ dimensions,
+by which the potential of each pixel $x_j$ assigned as each state of
+hidden state set $\{0, 1, \dots, 5\}$ is drawn and represented by
+a matrix with $(N, 2)$ dimension.
+
+The parameter $\omega$ is learned by our end-to-end CNN-Unet training 
+structure.
+
+{\textbf{Unary Potential}}
+$\varphi(y, h_j; \delta)$ measures the compatibility between global
+class label $y$ and the hidden local state $h_j$. This potential is
+parametrized as:
+\begin{equation*}
+\varphi(y, h_{j}; \delta) = \sum_{a \in Y} \sum_{b \in H} \delta_{a, b}
+\cdot \one(y = a) \cdot \one(h_j = b).
+\end{equation*}
+
+{\textbf{Binary Potential}}
+$\psi(h_i, h_j, x_i, x_j; \eta)$
+balances the hidden label compatibility of the neighboring pixels on the 
+feature image $\Fb(\Ib)$.
+We define this potential through a common contrast-sensitive two-kernal 
+potentials~\cite{krahenbuhl2011efficient}\cite{chen2022end}:
+\begin{equation}
+\begin{split}
+& \psi(h_i, h_j, x_i, x_j; \eta) = \mu(h_i, h_j) \Bigg[
+\omega_1 \exp \bigg(
+-\frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2} \\
+&\quad
+- \frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib)\right\rvert^2}{2\eta_\beta^2}
+\bigg)
++ \omega_2 \exp \left(
+- \frac{\left\lvert p_i - p_j \right\rvert^2}{2 \eta_\gamma^2}
+\right)
+\Bigg]
+\end{split}
+\end{equation}
+
+\section{Variational Inference of HCRF}
+
+\subsection{ELBO Under Mean-Filed Variational Family}
+
+The existence of the hidden variable $\hb$ makes our model more interpretable.
+However, since We will train our model through maximal likelihood method,
+introducing the hidden variable will make precisely computation the summation
+$\sum_{\hb \in H^N} \PP(y, \hb \vert \xb; \theta)$ intractable.
+To address this issue, we make our computation more feasible through a
+mean-field variational inference framework, which enables our model
+efficiently estimating the log-likelihood $\log\PP(y \vert \xb; \theta)$.
+
+Firstly, we introduce the mean-field varepsilon family approximation of
+$\log\PP(\hb \vert y, \xb; \theta)$ as:
+\begin{equation*}
+\QQ(\hb) = \prod_{i=1}^N q_i(h_i).
+\end{equation*}
+Then the evidence lower bound (ELBO) under the mean-field family is given by
+\begin{equation}
+\begin{split}
+& \textrm{ELBO}(\QQ) = \EE_{\QQ(\hb)} \log\PP(y, \hb \vert \xb; \theta)
+- \EE_{\QQ(\hb)} \log q(\hb) \\
+&\quad = \sum_{i=1}^N \EE_{q_i(h_i)}
+\left[ \phi(h_j, x_j; \omega) + \varphi(y, h_j, x_j; \delta) \right] \\
+&\quad + \sum_{i=1}^N \sum_{j=1}^N \sum_{l,l^\prime} \one_{\{i \neq j\}} 
+\QQ(h_i = l) \QQ(h_j = l^\prime) \mu(l, l^\prime) \\
+&\quad \Bigg[ \omega_1 \exp \left(
+- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2}
+- \frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib) \right\rvert^2}{2\eta_\beta^2}
+\right) \\
+&\quad + \omega_2 \exp \left(
+- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\gamma^2}
+\right) \Bigg] \\
+&\quad - \sum_{i=1}^N \EE_{q_i} \log q_i(h_i) + \vartheta(y, \xb; \varpi).
+\end{split}
+\end{equation}
+
+\subsection{Updating Variational Family}
+
+By the coordinate ascent variational inference(CAVI) algorithm,
+we can obtain that fix $q_j(h_j), \forall j \neq i$, the optimal $q_i(h_i)$
+that maximizes $\textrm{ELBO}(q_i(h_i))$ is given by
+\begin{equation}
+\QQ(h_i) \propto \exp\{\EE_{\QQ(\hb_{-i})}
+\log\PP(y, h_i, \hb_{-i} \vert \xb; \theta)\}.
+\end{equation}
+Thus, we can get that
+\begin{equation}
+\begin{split}
+q_i(l) &= \frac{1}{Z_i} \exp \Bigg\{
+\phi(h_j, x_j; \omega) + \varphi(y, h_j, x_j; \delta) \\
+&+ \sum_{j \neq i} \sum_{l^\prime} q(l^\prime) \mu(l, l^\prime) 
+\Bigg[ \omega_1 \exp \big(
+- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2}\\
+&- \frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib) \right\rvert^2}{2\eta_\beta^2}
+\big)
++ \omega_2 \exp \big(
+- \frac{\left\lvert p_i - p_j 
+\right\rvert^2}{2\eta_\gamma^2}
+\big)
+\Bigg]
+\Bigg\}
+\end{split}
+\end{equation}
+
+\subsection{Computational Complexity Analysis}
+
+\begin{itemize}
+\item {\textbf{Updating Variational Family}} $\{q_i(h_i)\}$:
+The computationally expensive part of the fixed point iteration of $q_i(h_i)$
+comes from the message passing, i.e. the convolution of $\QQ(\hb)$ and
+Gaussian kernal in $\psi(h_i, h_j, x_i, x_j; \eta)$. We need $\Ocal(N^2)$
+runtime for $N$ pixels when evaluating this term precisely.
+To make the computation more feasible, a truncated gaussian kernal will be
+employed, which is only supported on spaces within certain proportion of
+the untruncated standard deviation. Then the approximate message passing is
+linear in the number of pixels $N$.
+\item {\textbf{Computation of the ELBO}}: The summation over the product of
+$q_i(h_i)q_j(h_j)$ and the gaussian kernal in $\psi(h_i, h_j, x_i, x_j; \eta)$
+also requires a $\Ocal(N^2)$ complexity in time.
+Owes to a similar truncated gaussian kernal approxiamtion, the runtime can
+be reduced to $\Ocal(N)$.
+\end{itemize}
+
+\bibliographystyle{IEEEtran}
 \bibliography{refs}

### 2024-11-09 progress/progress.tex
revise margins
add ELBO

--- 2024-10-30 progress/progress.tex
@@
-\documentclass[12pt]{article}
+\documentclass[10pt]{article}
 
-\usepackage[margin = 1in]{geometry}
+\usepackage[left=.5in, right=.5in, top=.3in, bottom=.5in]{geometry}
+\usepackage{amsfonts,amsmath,amssymb}
 \usepackage{authblk}
+%% put affiliations into one line
+\makeatletter
+\renewcommand\AB@affilsepx{\quad\protect\Affilfont}
+\usepackage{bbm}
 \usepackage[inline]{enumitem}
+\usepackage{graphicx}
+\graphicspath{{./}{../image/}}
 \usepackage[colorlinks=true,citecolor=blue,urlcolor=blue]{hyperref}
@@
 \usepackage{setspace}
-\onehalfspacing
+\setstretch{1.2}
+\usepackage{titlesec}
+\titlespacing*{\section}{0pt}{5pt}{5pt}
+
+\newcommand{\EE}{{\mathbb{E}}}
+\newcommand{\PP}{{\mathbb{P}}}
+\newcommand{\QQ}{{\mathbb{Q}}}
+\newcommand{\Fb}{\mathbf{F}}
+\newcommand{\Ib}{\mathbf{I}}
+\newcommand{\hb}{\mathbf{h}}
+\newcommand{\xb}{\mathbf{x}}
+\newcommand{\one}{\mathbbm{1}}
+\newcommand{\Ocal}{\mathcal{O}}
 
@@
 
-\title{Progress Report: Transformer-CRF Integration for Sequence Labeling on
-EEG Data}
+\title{\vspace{-1cm} \Large
+Progress Report: Transformer-CRF Integration for Sequence Labeling on EEG Data}
 
@@
 
-\affil[1]{Department of Mathematics, University of Connecticut}
-\affil[2]{Department of Statistics, University of Connecticut}
+\affil[1]{Department of Mathematics, UConn}
+\affil[2]{Department of Statistics, UConn}
 
-\date{\today}
+\date{\vspace{-1.3cm}}
 
@@
 
+The project is progressing according to schedule. To date, we have made
+significant strides in several key areas:
+\begin{enumerate}
+\item Model Development:
+We have initiated development of a convolutional neural network (CNN) combined
+with a hidden conditional random field (HCRF). This hybrid approach aims to
+create a robust sequence labeling framework capable of capturing both local
+and global dependencies within the data sequences.
+\item Variational Inference Implementation:
+To efficiently approximate the posterior distributions in our complex model,
+we have integrated variational inference into the HCRF component.
+This technique is particularly effective in handling the computational
+challenges posed by the dataset's heterogeneity.
+\item Data Visualization:
+We have successfully visualized key features of the dataset.
+\end{enumerate}
+%This progress has provided a foundation for the next
+%steps of our research.
+
 \section{Preliminary Results}
 % Summary of initial findings and analyses.
+
+The potential of our HCRF model:
+\begin{equation*}
+\Phi(y, \hb, \xb; \theta) = \underbrace{
+\sum_{j \in \nu} \phi(h_j, x_j; \omega) +
+\sum_{i \neq j} \psi(h_i, h_j, x_i, x_j; \eta)}_{
+% \propto \log\PP( \hb \vert \xb; \theta)
+\textrm{Measures log-likelihood $\log\PP(\hb \vert \xb; \theta)$}}
++ \underbrace{
+\sum_{j \in \nu} \varphi(y, h_j, x_j; \delta) + \vartheta(y, \xb; \varpi)
+}_{
+% \propto \log\PP(y | \hb, \xb; \theta)
+\textrm{Measures log-likelihood $\log\PP(y \vert \hb, \xb; \theta)$}
+}
+\end{equation*}
+
+\begin{itemize}
+\item \textbf{Unary Potential}
+$\phi(h_j, x_j; \omega)$ measures the likelihood of the local feature
+$x_j$ is assigned as the hidden attention state $h_j$.
+We generate the likelihood through a softmax operation on $\Fb(\xb)$
+the feature maps of CNN with $(N, C_{\textrm{out}})$ dimensions,
+by which the potential of each pixel $x_j$ assigned as each state of
+hidden state set $\{0, 1, \dots, 5\}$
+is drawn and represented by a matrix with $(N, 2)$ dimension.
+\item The parameter $\omega$ is learned by our end-to-end CNN-Unet
+training structure.
+\item \textbf{Unary Potential} $\varphi(y, h_j; \delta)$ measures
+the compatibility between global class label $y$ and the hidden local
+state $h_j$. This potential is parametrized as:
+\begin{align*}
+\varphi(y, h_{j}; \delta) = \sum_{a \in Y} \sum_{b \in H} \delta_{a, b}
+\cdot \one(y = a) \cdot \one(h_j = b).
+\end{align*}
+\item \textbf{Binary Potential} $\psi(h_i, h_j, x_i, x_j; \eta)$
+balances the hidden label compatibility of the neighboring pixels on the 
+feature image $\Fb(\Ib)$.
+We define this potential through a common contrast-sensitive two-kernal 
+potentials~\citep{krahenbuhl2011efficient,chen2022end}:
+\begin{equation*}
+\psi(h_i, h_j, x_i, x_j; \eta) = \mu(h_i, h_j) \Bigg[
+\omega_1 \exp \left(
+-\frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2}
+-\frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib) \right\rvert^2}{2\eta_\beta^2} 
+\right) + \omega_2 \exp \left(
+- \frac{\left\lvert p_i - p_j \right\rvert^2}{2 \eta_\gamma^2}
+\right)
+\Bigg]
+\end{equation*}
+\end{itemize}
+
+The mean-field varepsilon family approximation of
+$\log\PP(\hb \vert y, \xb; \theta)$ as $\QQ(\hb) = \prod_{i=1}^N q_i(h_i)$.
+Then the evidence lower bound under the mean-field family is given by
+\begin{equation*}
+\begin{split}
+& \textrm{ELBO}(\QQ) = \EE_{\QQ(\hb)} \log\PP(y, \hb \vert \xb; \theta)
+- \EE_{\QQ(\hb)} \log q(\hb) = \sum_{i=1}^N \EE_{q_i(h_i)}
+\left[ \phi(h_j, x_j; \omega) + \varphi(y, h_j, x_j; \delta) \right]
++ \sum_{i=1}^N \sum_{j=1}^N \sum_{l,l^\prime} \one_{\{i \neq j\}}\QQ(h_i = l) \\
+&\quad \QQ(h_j = l^\prime) \mu(l, l^\prime)
+\Bigg[ \omega_1 \exp \left(
+- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2}
+- \frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib) \right\rvert^2}{2\eta_\beta^2}
+\right)
++ \omega_2 \exp \left(
+- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\gamma^2}
+\right) \Bigg]
+- \sum_{i=1}^N \EE_{q_i} \log q_i(h_i) + \vartheta(y, \xb; \varpi).
+\end{split}
+\end{equation*}
+
+By the coordinate ascent variational inference(CAVI) algorithm,
+we can obtain that fix $q_j(h_j), \forall j \neq i$, the optimal $q_i(h_i)$
+that maximizes $\textrm{ELBO}(q_i(h_i))$ is given by
+$\QQ(h_i) \propto \exp\{\EE_{\QQ(\hb_{-i})}
+\log\PP(y, h_i, \hb_{-i} \vert \xb; \theta)\}$.
+Thus, we can get that
+
+\begin{equation*}
+q_i(l) = \frac{ \exp \left\{
+\phi(h_j, x_j; \omega) + \varphi(y, h_j, x_j; \delta)
++ \sum_{j \neq i} \sum_{l^\prime} q(l^\prime) \mu(l, l^\prime) 
+\left[ \omega_1 \exp \left(
+- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2}
+- \frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib) \right\rvert^2}{2\eta_\beta^2}
+\right)
++ \omega_2 \exp \left(
+- \frac{\left\lvert p_i - p_j 
+\right\rvert^2}{2\eta_\gamma^2}
+\right)
+\right]
+\right\}
+}{Z_i}
+\end{equation*}
+
+While the models are still under development and optimization,
+initial results are promising. For instance, Figure~\ref{fig:SEIZURE_3}
+illustrates a sample prediction with ground truth 94.\% Seizure and 2.9\% LPD.
+Our model predicts 88.4\% Seizure and 1.6\% LPD.
+Yellow highlights indicate the regions where the model focused its attention.
+These preliminary results show trends that are consistent with our initial
+hypotheses. The visualizations have proven useful in pinpointing areas for
+further investigation, and the derived formulas appear to be robust and
+aligned with the expected model outcomes. Further analysis will refine
+these findings and explore additional dimensions of data.
+
+\begin{figure}[tbp]
+\centering
+\includegraphics[width=.7\textwidth]{grad_cam_SEIZURE_3}
+\caption{An example of data visualization. Areas highlighted in yellow
+indicate regions where the model focused its attention during this specific
+prediction.}
+\label{fig:SEIZURE_3}
+\end{figure}
 
@@
 
+We will provide regular updates on model performance, challenges encountered,
+and any significant changes to the project plan.
+
+\bibliography{../manuscript/refs}
+\bibliographystyle{chicago}
+
 \end{document}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2024-11-11 progress/progress (1).tex
add modification of the progress report

--- 2024-11-09 progress/progress.tex
@@
 \usepackage{authblk}
+% \usepackage{hyperref}
 %% put affiliations into one line
@@
 with a hidden conditional random field (HCRF). This hybrid approach aims to
-create a robust sequence labeling framework capable of capturing both local
-and global dependencies within the data sequences.
+create a robust image segmentation framework capable of capturing both local
+and global dependencies within the data structure.
 \item Variational Inference Implementation:
@@
 challenges posed by the dataset's heterogeneity.
-\item Data Visualization:
-We have successfully visualized key features of the dataset.
+\item Data Preprocessing:
+% We have finished data preprocessing of the eeg figures and spectrogram image. We transformer the 
+% eeg signals to spectrograms through build-in functions from open source libraries like OpenCV and PyWavelets.
+% We then concate these spectrograms as a \(512 \times 512\) image, which is ready to be feeded in our model.
+We completed the data preprocessing for EEG signals and their corresponding spectrograms.
+The EEG signals were transformed into spectrogram images using built-in functions from open-source libraries such as OpenCV and PyWavelets.
+We then concatenated these spectrograms into a \(512 \times 512\) image, which is now ready for input into our model.
 \end{enumerate}
@@
 
-The potential of our HCRF model:
+In our model, the variable \(y \) represents the corresponding brain signal label of the inpute image \(\mathbf{x}\). 
+The hidden variable \(\mathbf{h} = \{h_{i}\}\) represents the label at each pixels of \(\mathbf{x}\). The parameter \(\theta\)
+contain parameters of neural networks as well as the HCRF.
+We introduce the potential functions of our HCRF model as:
 \begin{equation*}
@@
 is drawn and represented by a matrix with $(N, 2)$ dimension.
-\item The parameter $\omega$ is learned by our end-to-end CNN-Unet
+\item The parameter $\varpi$ will be learned by end-to-end CNN-Unet
 training structure.
@@
 Thus, we can get that
-
 \begin{equation*}
-q_i(l) = \frac{ \exp \left\{
+\begin{aligned}
+& q_i(l) = \\
+&\frac{1}{Z_i} \exp \left\{
 \phi(h_j, x_j; \omega) + \varphi(y, h_j, x_j; \delta)
@@
 \right\}
-}{Z_i}
+\end{aligned}
 \end{equation*}
 
-While the models are still under development and optimization,
-initial results are promising. For instance, Figure~\ref{fig:SEIZURE_3}
-illustrates a sample prediction with ground truth 94.\% Seizure and 2.9\% LPD.
-Our model predicts 88.4\% Seizure and 1.6\% LPD.
-Yellow highlights indicate the regions where the model focused its attention.
-These preliminary results show trends that are consistent with our initial
-hypotheses. The visualizations have proven useful in pinpointing areas for
-further investigation, and the derived formulas appear to be robust and
-aligned with the expected model outcomes. Further analysis will refine
-these findings and explore additional dimensions of data.
+% While the models are still under development and optimization,
+% initial results are promising. 
+% For instance, Figure~\ref{fig:SEIZURE_3}
+% illustrates a sample prediction with ground truth 94.\% Seizure and 2.9\% LPD.
+% Our model predicts 88.4\% Seizure and 1.6\% LPD.
+% Yellow highlights indicate the regions where the model focused its attention.
+% These preliminary results show trends that are consistent with our initial
+% hypotheses. The visualizations have proven useful in pinpointing areas for
+% further investigation, and 
+% The derived formulas appear to be robust and
+% aligned with the expected model outcomes. Further analysis will refine
+% these findings and explore additional dimensions of data.
+After building the theoretical model formualtion of our project, we also finished 
+ preliminary analyses of the data.
+Our preprocessed spectrograms dataset that are ready for training model.
+The dataset are available:{https://github.com/Carol-seven/Transformer-CRF\_EEG/tree/main}.
 
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=.7\textwidth]{grad_cam_SEIZURE_3}
-\caption{An example of data visualization. Areas highlighted in yellow
-indicate regions where the model focused its attention during this specific
-prediction.}
-\label{fig:SEIZURE_3}
-\end{figure}
+% \begin{figure}[tbp]
+% \centering
+% % \includegraphics[width=.7\textwidth]{grad_cam_SEIZURE_3}
+% \caption{An example of data visualization. Areas highlighted in yellow
+% indicate regions where the model focused its attention during this specific
+% prediction.}
+% \label{fig:SEIZURE_3}
+% \end{figure}
 
@@
 % Outline of any adjustments made to the original plan.
-
-At this stage, we have not made significant changes to our original plan.
-However, due to the complexity of the dataset and model, we may consider
-slightly adjust the project timeline to allocate additional time for
-model fine-tuning.
+% At this stage, 
+% \noindent
+% \textcolor{blue}{
+	% Based on extra literature reviews and our preliminary results, we have made two major adjustments to our project.
+	\begin{itemize}
+		\item \textbf{CRF Layer Modification:} Our motivation for introducing the CRF layer was to allow the model to identify regions of input 
+		images that correspond to abnormal brain activity signals. However, the original dataset, including spectrograms 
+		and EEG-transformed spectrograms, lacked segmentation information, so no labels were available to train the CRF. 
+		Our initial solution was to generate pixel-wise CRF labels based on the global ground truth distribution. 
+		However, further review revealed that this approach could lead to isolated label regions, counteracting 
+		our intent for CRF integration. Consequently, we modified the CRF layer to a hidden CRF (HCRF) 
+		by setting pixel labels as hidden variables and adopted a variational inference framework to manage
+		 these hidden variables in the CRF model.
+		\item \textbf{Model Structure Adjustment:}We replaced the Transformer-based architecture with a U-Net CNN structure. This choice reflects 
+		findings in recent research that fine-tuning large models like Transformers can be time-consuming. Instead,
+		 we opted for the well-established U-Net CNN, a classical model for image segmentation, 
+		 which integrates effectively with the CRF layer.
+	\end{itemize}
+% }
+% we have not made significant changes to our original plan.
+% However, due to the complexity of the dataset and model, we may consider
+% slightly adjust the project timeline to allocate additional time for
+% model fine-tuning.
 
@@
 \item Expanding and refining the visualizations based on updated data.
-\item Finalizing the derivation and validation of all essential formulas.
+\item Finalizing the conduction of the neural network model including the Unet CNN layer, HCRF layer based on the formulas we have derived.
+\item Training and tuning the model on our preprocessed data. 
 \item Conducting a comprehensive analysis with the refined model.

### 2024-11-11 progress/progress.tex
reformat report

@@
 
-\usepackage[left=.5in, right=.5in, top=.3in, bottom=.5in]{geometry}
+\usepackage[left=.5in, right=.5in, top=.3in, bottom=.3in]{geometry}
 \usepackage{amsfonts,amsmath,amssymb}
@@
 \usepackage{titlesec}
-\titlespacing*{\section}{0pt}{5pt}{5pt}
+\titlespacing*{\section}{0pt}{8pt}{8pt}
 
@@
 challenges posed by the dataset's heterogeneity.
-\item Data Visualization:
-We have successfully visualized key features of the dataset.
+\item Data Preprocessing:
+We completed the data preprocessing for EEG signals and their corresponding
+spectrograms. The EEG signals were transformed into spectrogram images using
+built-in functions from open-source libraries such as OpenCV and PyWavelets.
+We then concatenated these spectrograms into a 512 $\times$ 512 image,
+which is now ready to be used as input for our model.
 \end{enumerate}
@@
 
-The potential of our HCRF model:
+In our model, the variable $y$ represents the corresponding brain signal
+label of the inpute image $\xb$. The hidden variable $\hb= \{h_i\}$
+represents the label at each pixels of $\xb$. The parameter $\theta$
+contain parameters of neural networks as well as the HCRF.
+We introduce the potential functions of our HCRF model as:
 \begin{equation*}
@@
 by which the potential of each pixel $x_j$ assigned as each state of
-hidden state set $\{0, 1, \dots, 5\}$
-is drawn and represented by a matrix with $(N, 2)$ dimension.
+hidden state set $\{0, 1, \dots, 5\}$ is drawn and represented by
+a matrix with $(N, 2)$ dimension.
 \item The parameter $\omega$ is learned by our end-to-end CNN-Unet
 training structure.
-\item \textbf{Unary Potential} $\varphi(y, h_j; \delta)$ measures
-the compatibility between global class label $y$ and the hidden local
-state $h_j$. This potential is parametrized as:
+\item \textbf{Unary Potential}
+$\varphi(y, h_j; \delta)$ measures the compatibility between global class
+label $y$ and the hidden local state $h_j$. This potential is parametrized as:
 \begin{align*}
@@
 \end{align*}
-\item \textbf{Binary Potential} $\psi(h_i, h_j, x_i, x_j; \eta)$
-balances the hidden label compatibility of the neighboring pixels on the 
-feature image $\Fb(\Ib)$.
+\item \textbf{Binary Potential}
+$\psi(h_i, h_j, x_i, x_j; \eta)$ balances the hidden label compatibility
+of the neighboring pixels on the feature image $\Fb(\Ib)$.
 We define this potential through a common contrast-sensitive two-kernal 
@@
 Then the evidence lower bound under the mean-field family is given by
+\vspace{-1cm}
 \begin{equation*}
@@
 Thus, we can get that
-
 \begin{equation*}
-q_i(l) = \frac{ \exp \left\{
+\begin{split}
+q_i(l) &= \frac{1}{Z_i} \exp \Bigg\{
 \phi(h_j, x_j; \omega) + \varphi(y, h_j, x_j; \delta)
-+ \sum_{j \neq i} \sum_{l^\prime} q(l^\prime) \mu(l, l^\prime) 
-\left[ \omega_1 \exp \left(
++ \sum_{j \neq i} \sum_{l^\prime} q(l^\prime) \mu(l, l^\prime) \\
+&\quad \left[ \omega_1 \exp \left(
 - \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2}
@@
 \right]
-\right\}
-}{Z_i}
+\Bigg\}
+\end{split}
 \end{equation*}
 
-While the models are still under development and optimization,
-initial results are promising. For instance, Figure~\ref{fig:SEIZURE_3}
-illustrates a sample prediction with ground truth 94.\% Seizure and 2.9\% LPD.
-Our model predicts 88.4\% Seizure and 1.6\% LPD.
-Yellow highlights indicate the regions where the model focused its attention.
-These preliminary results show trends that are consistent with our initial
-hypotheses. The visualizations have proven useful in pinpointing areas for
-further investigation, and the derived formulas appear to be robust and
-aligned with the expected model outcomes. Further analysis will refine
-these findings and explore additional dimensions of data.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=.7\textwidth]{grad_cam_SEIZURE_3}
-\caption{An example of data visualization. Areas highlighted in yellow
-indicate regions where the model focused its attention during this specific
-prediction.}
-\label{fig:SEIZURE_3}
-\end{figure}
+After building the theoretical model formualtion of our project, we also
+finished preliminary analyses of the data. Our preprocessed spectrograms
+dataset that are ready for training model. The dataset will be available on
+\href{https://github.com/Carol-seven/Transformer-CRF_EEG}{GitHub} and
+made public after the project is completed.
 
@@
 
-We encountered minor challenges related to data handling and computational
-limitations. In particular, the complexity of certain calculations required
-additional processing time, which slightly delayed our progress.
-We are currently exploring ways to optimize these computations.
+We encountered minor challenges with data handling and computational
+limitations. The complexity of certain calculations required
+additional processing time, slightly delayed our progress.
+We are exploring ways to optimize these computations.
 
@@
 
-At this stage, we have not made significant changes to our original plan.
-However, due to the complexity of the dataset and model, we may consider
-slightly adjust the project timeline to allocate additional time for
-model fine-tuning.
+\begin{itemize}
+\item \textbf{CRF Layer Modification:}
+Our motivation for introducing the CRF layer was to allow the model to
+identify regions of input images that correspond to abnormal brain activity
+signals. However, the original dataset, including spectrograms and
+EEG-transformed spectrograms, lacked segmentation information, meaning no
+labels were available to train the CRF.
+Our initial solution was to generate pixel-wise CRF labels based on
+the global ground truth distribution. However, further review revealed that
+this approach could lead to isolated label regions, which would counteract
+our intent for CRF integration. Consequently, we modified the CRF layer into
+a HCRF by setting pixel labels as hidden variables and adopted a variational
+inference framework to manage these hidden variables in the CRF model.
+\item \textbf{Model Structure Adjustment:}
+We replaced the Transformer-based architecture with a U-Net CNN structure.
+This choice reflects findings in recent research that fine-tuning large models
+like Transformers can be time-consuming. Instead, we opted for the
+well-established U-Net CNN, a classical model for image segmentation,
+which integrates effectively with the CRF layer.
+\end{itemize}
 
@@
 \item Expanding and refining the visualizations based on updated data.
-\item Finalizing the derivation and validation of all essential formulas.
+\item Finalizing the construction of the neural network model, including
+the U-Net CNN layer and HCRF layer, based on the formulas we have derived.
+\item Training and tuning the model on our preprocessed data.
 \item Conducting a comprehensive analysis with the refined model.
@@
 
-We will provide regular updates on model performance, challenges encountered,
+We will provide regular updates on model performance, challenges,
 and any significant changes to the project plan.
 
+\footnotesize
 \bibliography{../manuscript/refs}

### 2024-11-11 progress/progress.tex
minor
rename the project

@@
 \usepackage{titlesec}
-\titlespacing*{\section}{0pt}{8pt}{8pt}
+\titlespacing*{\section}{0pt}{5pt}{5pt}
 
@@
 
-\title{\vspace{-1cm} \Large
-Progress Report: Transformer-CRF Integration for Sequence Labeling on EEG Data}
+\title{\vspace{-1cm}
+Progress Report: UNet-CRF Integration for Sequence Labeling on EEG Data}
 
@@
 dataset that are ready for training model. The dataset will be available on
-\href{https://github.com/Carol-seven/Transformer-CRF_EEG}{GitHub} and
+\href{https://github.com/Carol-seven/UNet-CRF_EEG}{GitHub} and
 made public after the project is completed.
@@
 
-\footnotesize
 \bibliography{../manuscript/refs}

### 2024-11-15 manuscript/main.tex
combine the current results; IEEEtranN

@@
 \usepackage[colorlinks=true,citecolor=blue,urlcolor=blue]{hyperref}
+\usepackage[numbers]{natbib}
 \usepackage{xcolor}
@@
 
+\newcommand{\sx}[1]{\textcolor{red}{(SX: #1)}}
+
 \begin{document}
 
-\title{\LARGE Transformer-CRF Integration for Sequence Labeling on EEG Data}
+\title{\huge UNet-CRF Integration for Sequence Labeling on EEG Data}
 
@@
 
-In this project, we will integrate powerful deep neural networks and
-statistical methods with a probabilistic graphical model (PGM) to effectively
-tackle sequence labeling tasks on a heterogeneous medical care dataset.
+%In this project, we will integrate powerful deep neural networks and
+%statistical methods with a probabilistic graphical model (PGM) to effectively
+%tackle sequence labeling tasks on a heterogeneous medical care dataset.
 
@@
 
-probabilistic graphical model
+probabilistic graphical model, UNet, fully connected CRF,
 
@@
 
-\section{CNN-HCRF Model}
+\section{Introduction}
+%Introduction: problem definition and motivation.
+
+Sequence labeling tasks are pivotal in diverse domains, especially healthcare,
+where precise and timely data classification can significantly impact patient
+outcomes. Medical datasets, such as physiological signals, often exhibit
+complexity and heterogeneity, posing challenges for traditional modeling
+techniques. Therefore, developing models capable of extracting meaningful
+patterns from these datasets is crucial for accurate diagnostic and prognostic
+decisions.
+
+In recent years, deep learning has emerged as a powerful tool for sequence
+labeling. Neural network architectures like long short-term memory (LSTM) and
+Transformer models have revolutionized sequential data processing.
+Additionally, conditional random fields (CRFs) have been employed for modeling
+label dependencies, demonstrating promising results.
+However, while these methods have shown individual strengths, a comprehensive
+approach that leverages their complementary capabilities is essential for
+tackling complex, multi-modal datasets.
+
+This project introduces a novel framework that combines deep neural networks
+and probabilistic graphical models (PGMs) to enhance sequence labeling
+performance. By employing Transformer~\citep{vaswani2017attention} for feature
+extraction and integrating them with CRF models,
+we aim to improve sequence tagging accuracy. Furthermore, mean-field
+approximation techniques are utilized for efficient inference, ensuring a
+computationally efficient yet effective approach.
+This integration of modern neural network architectures with CRFs offers
+a robust and promising solution for complex sequence labeling tasks.
+
+\section{Background}
+%Background & Related Work: background info and literature survey.
+
+\sx{Add UNet}
+
+Deep neural networks, particularly recurrent neural networks (RNNs) like LSTM,
+are widely used for sequence labeling due to their ability to capture
+long-range dependencies in sequences. Bidirectional LSTM (Bi-LSTM) further
+enhances this by processing sequences in both directions, improving context
+understanding.
+CRFs, on the other hand, are commonly used for structured predictions,
+modeling dependencies between neighboring labels to ensure valid output
+sequences.
+\citet{huang2015bidirectional} pioneered the use of Bi-LSTM integrated with
+CRFs for sequence labeling tasks, demonstrating their effectiveness in
+extracting features and ensuring label consistency.
+
+However, Bi-LSTMs can still struggle to fully account for structural
+dependencies between predicted labels, leading to potential inconsistencies in
+the output sequence. Additionally, they typically require large amounts of
+labeled data and are computationally expensive, making them challenging to
+deploy in resource-constrained environments, such as hospitals.
+
+Transformer-based models, such as BERT~\citep{devlin2019bert},
+have emerged as powerful alternatives to traditional RNN-based approaches.
+Trained on massive datasets in an unsupervised manner, Transformer can be
+fine-tuned for various natural language processing tasks,
+including sequence labeling.
+\citet{devlin2019bert} directly compared BERT to traditional models, including
+Bi-LSTM, and demonstrated its superior performance on tasks like named entity
+recognition, highlighting its ability to capture more feature information
+than LSTM models.
+
+While CRFs are effective for modeling label dependencies, they often lack
+efficient closed-form solutions, especially for complex models beyond
+linear-chain CRFs. To address this, \citet{krahenbuhl2011efficient} introduced
+the use of mean-field approximation to solve more general CRFs in the context
+of complex prediction tasks. By minimizing the Kullback-Leibler (KL)
+divergence, mean-field approximation can be reduced to a fixed-point iteration,
+providing a faster and more efficient approach for training and inference
+while maintaining suitable accuracy. \citet{zheng2015conditional} further
+extended this technique by reformulating mean-field approximation for fully
+connected CRFs, integrating an RNN structure to capture temporal dependencies.
+
+\section{Methods}
+%Methods: overview of your proposed method, Intuition on why should it be
+%better than the state of the art, and details of models and algorithms that
+%you developed.
+
+\subsection{CNN-HCRF Model}
 
@@
 We define this potential through a common contrast-sensitive two-kernal 
-potentials~\cite{krahenbuhl2011efficient}\cite{chen2022end}:
+potentials~\citep{krahenbuhl2011efficient, chen2022end}:
 \begin{equation}
@@
 
-\section{Variational Inference of HCRF}
+\subsection{Variational Inference of HCRF}
 
-\subsection{ELBO Under Mean-Filed Variational Family}
+\subsubsection{ELBO Under Mean-Filed Variational Family}
 
@@
 
-\subsection{Updating Variational Family}
+\subsubsection{Updating Variational Family}
 
@@
 
-\subsection{Computational Complexity Analysis}
+\subsubsection{Computational Complexity Analysis}
 
@@
 
-\bibliographystyle{IEEEtran}
+\section{Results}
+%Results: Description of your experiments and results and a list of questions
+%your experiments are designed to answer.
+
+\section{Conclusion}
+%Conclusion: discussion and future work.
+
+\bibliographystyle{IEEEtranN}
 \bibliography{refs}

### 2024-12-06 Code_for_Xiaohui/UNet_HCRF_model.py
Update code for pretrain

@@ -1,0 +1,49 @@
+import tensorflow as tf
+from tensorflow.keras.models import Model
+from tensorflow.keras.layers import Input, concatenate, Conv2D, AveragePooling2D, UpSampling2D, \
+  BatchNormalization, Activation, LeakyReLU, Dropout, GlobalAveragePooling2D
+from tensorflow.keras.layers import Layer, InputSpec
+from tensorflow.keras import initializers, regularizers, constraints
+from tensorflow.keras import backend as K
+
+sys.path.append(os.path.dirname(os.path.abspath(__file__)))
+from utility import InstanceNormalization, UNet_layer
+from tensorflow.keras.layers import GroupNormalization
+from tensorflow.keras.optimizers import Adam, Adadelta, SGD
+import Modules.LossFunction as lf
+import Modules.Common_modules as cm
+import numpy as np
+
+
+def unet_global(image_size = (512, 512), num_classes=6, **kwargs):
+    '''
+    You can adjust the Unet structure by setting:
+    - kwargs.num_classes
+    - kwargs.num_features
+    '''
+    inputs = Input(image_size + (3,))
+    extract_feature = UNet_layer(**kwargs)(inputs)
+    x = Conv2D(filters=64, kernel_size=3, strides=1, padding="same")(x)
+    x = LeakyReLU(alpha=1e-1)(x)
+    x = AveragePooling2D(2, strides=2, padding="same")(x)
+    x = BatchNormalization()(x)
+    x = Conv2D(filters=128, kernel_size=3, strides=1, padding="same")(x)
+    x = LeakyReLU(alpha=1e-1)(x)
+    x = AveragePooling2D(2, strides=2, padding="same")(x)
+    x = BatchNormalization()(x)
+
+    x = GlobalAveragePooling2D()(x)
+    x = Dense(units=265, activation="relu")(x)
+    
+    x = Dropout(0.3)(x)
+    outputs = Dense(units=num_classes, activation="softmax")(x)
+    model = Model(inputs=inputs, outputs=outputs)
+    opt = tf.keras.optimizers.Adam(learning_rate = 1e-4)
+    loss = tf.keras.losses.KLDivergence()
+
+    model.compile(loss=loss, optimizer = opt) 
+    return model
+
+
+
+

### 2024-12-06 Code_for_Xiaohui/utility.py

@@ -1,0 +1,54 @@
+from tensorflow.keras.models import Model
+from tensorflow.keras.layers import Input, concatenate, Conv2D, AveragePooling2D, UpSampling2D, \
+  BatchNormalization, Activation, LeakyReLU, Dropout
+from tensorflow.keras.layers import Layer, InputSpec
+from tensorflow.keras import initializers, regularizers, constraints
+from tensorflow.keras import Sequential
+from tensorflow.keras.layers import GroupNormalization
+
+"""
+Normalization layer that is used in:
+    Chen S, Gamechi Z S, Dubost F, et al. An end-to-end approach to segmentation in medical images with CNN and posterior-CRF.
+"""
+
+class UNet_layer(Layer):
+    def __init__(self, input_shape, num_classes=6, num_features=32, blocks=[1, 2, 4, 8, 16], trainable=False, **kwargs):
+        self.input_shape_ = kwargs.get('input_shape' ,input_shape)
+        self.num_classes = kwargs.get('num_classes' ,num_classes)
+        self.num_features = kwargs.get('num_features' ,num_features) # default 32
+        self.blocks = blocks # default [1, 2, 4, 8, 16]
+        self.down_block_names = [f'down_block_relu_{i}' for i in range(len(blocks)-1)] + [f'down_block_relu_{len(blocks)-1}']
+        inputs = Input(shape=self.input_shape_)
+        x = inputs
+        skips = []
+        for i, block in enumerate(self.blocks):
+            x = Conv2D(self.num_features * block, 3, strides=1, padding="same")(x)
+            x = GroupNormalization(groups=-1)(x)
+            x = LeakyReLU(alpha=1e-1)(x)
+            x = Conv2D(self.num_features * block, 3, strides=1, padding="same")(x)
+            x = GroupNormalization(groups=-1)(x)
+            x = LeakyReLU(alpha=1e-1, name=f'down_block_relu_{i}')(x)
+            if i < len(self.blocks) - 1:
+                x = AveragePooling2D(2, strides=2, padding="same", name=f'down_block_pooling_{i}')(x)
+        base_model = Model(inputs=inputs, outputs=x)
+        self.down_stack = Model(inputs, outputs = [
+            base_model.get_layer(name).output for name in self.down_block_names
+        ])  # Outputs: bottleneck + skips
+        self.down_stack.trainable = trainable
+        super(UNet_layer, self).__init__(**kwargs)
+
+
[…]

### 2024-12-07 Code_for_Xiaohui/module_and_example.py
an example with toy data

@@ -1,0 +1,209 @@
+from tensorflow.keras.layers import (
+    Layer,
+    Conv2D,
+    UpSampling2D,
+    AveragePooling2D,
+    LeakyReLU,
+    Input,
+    concatenate,
+    BatchNormalization,
+    Dense,
+    Dropout,
+    GlobalAveragePooling2D,
+)
+from tensorflow.keras.models import Model
+
+class UNet_layer(Layer):
+    def __init__(self, input_shape, num_classes=6, num_features=32, blocks=[1, 2, 4, 8, 16], trainable=False, **kwargs):
+        super(UNet_layer, self).__init__(**kwargs)
+
+        # Save parameters
+        self.input_shape_ = input_shape
+        self.num_classes = num_classes
+        self.num_features = num_features
+        self.blocks = blocks
+        self.trainable = trainable
+
+        # Define down-sampling layers
+        self.down_stack = []
+        for i, block in enumerate(self.blocks):
+            self.down_stack.append(
+                [
+                    Conv2D(self.num_features * block, 3, strides=1, padding="same"),
+                    BatchNormalization(),
+                    LeakyReLU(alpha=0.1),
+                    Conv2D(self.num_features * block, 3, strides=1, padding="same"),
+                    BatchNormalization(),
+                    LeakyReLU(alpha=0.1, name=f"down_block_relu_{i}"),
+                    AveragePooling2D(pool_size=(2, 2), strides=2, padding="same") if i < len(self.blocks) - 1 else None,
+                ]
+            )
+
+        # Define up-sampling layers
+        self.up_stack = []
+        for i, block in enumerate(self.blocks[:-1][::-1]):
+            self.up_stack.append(
+                [
+                    UpSampling2D(size=(2, 2)),
+                    Conv2D(self.num_features * block, 3, strides=1, padding="same"),
+                    BatchNormalization(),
+                    LeakyReLU(alpha=0.1),
+                    Conv2D(self.num_features * block, 3, strides=1, padding="same"),
+                    BatchNormalization(),
+                    LeakyReLU(alpha=0.1),
+                ]
+            )
+
+    def call(self, inputs, training=None):
+        """
+        Forward pass through the U-Net.
+        """
+        # Down-sampling path
+        x = inputs
+        skip_connections = []
[…]

### 2024-12-08 manuscript/slides.tex
slides draft

\documentclass[leqno]{beamer}
\usetheme{Madrid}
\usecolortheme{seahorse}

\usepackage[linesnumbered,algoruled,boxed,lined]{algorithm2e}
\usepackage{amsfonts,amsmath,amssymb}
\usepackage{bbm}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{color}
\usepackage{graphicx}
\graphicspath{{./}{../image/}} % graphic path
\usepackage{hyperref}
\hypersetup{colorlinks,citecolor=blue,linkcolor=[RGB]{50,50,172}}
\usepackage{multicol}
\usepackage{multirow}
\usepackage[authoryear]{natbib}
\usepackage{setspace}
\usepackage{soul}
\usepackage{textpos}
\usepackage{tikz}

%% notations; only define those you use frequently
\newcommand{\EE}{{\mathbb{E}}}
\newcommand{\PP}{{\mathbb{P}}}
\newcommand{\QQ}{{\mathbb{Q}}}
\newcommand{\Fb}{\mathbf{F}}
\newcommand{\Ib}{\mathbf{I}}
\newcommand{\hb}{\mathbf{h}}
\newcommand{\xb}{\mathbf{x}}
\newcommand{\one}{\mathbbm{1}}
\newcommand{\Ocal}{\mathcal{O}}

% consistent with R manual
\newcommand\pkg[1]{\texttt{#1}}
\let\proglang=\textsf \let\code=\texttt

\setbeamercovered{transparent}
\setbeamertemplate{caption}[numbered]
\setbeamertemplate{enumerate item}[default]
\setbeamertemplate{itemize item}[circle]
\setbeamertemplate{section in toc}[default]
\setbeamertemplate{subsection in toc}[default]

\AtBeginSection[]{
\begin{frame}<beamer>{Overview}
\tableofcontents[currentsection]
\end{frame}}

\title[\textcolor{black}{UNet-CRF\_EEG}]{\large
UNet-CRF Integration for Sequence Labeling on EEG Data}

\author[\scalebox{.85}{Xiaohang Ma, Shiying Xiao, Xiaohui Yin}]
{Xiaohang Ma$^1$, Shiying Xiao$^2$, Xiaohui Yin$^2$}

\institute[\scalebox{.85}{UConn}]
{$^1$Department of Mathematics, University of Connecticut \\
$^2$Department of Statistics, University of Connecticut}

\date[December 9, 2024]{
{\small Storrs, CT} \\
{\small December 9, 2024}}

\begin{document}

\begin{frame}[plain]
\titlepage
\end{frame}

\begin{frame}
\frametitle{Overview}
\tableofcontents
\end{frame}

\section[Introduction]{Introduction}

\section[Data Preprocessing]{Data Preprocessing}

\begin{frame}{Data Preparation}
\begin{itemize}
\item Extract the 10 min long Kaggle Spectrogram centered at the middle time
where the experts rating (Global EEGs)
\item Transform the 50s long EEG waveforms into Spectrograms
\end{itemize}
\end{frame}

\begin{frame}{EEG signal denosing using Wavelet transform}
\begin{footnotesize}
\begin{align*}
& \text{LL Spec} = \frac{1}{4} \left[
spec(Fp1-F7) + spec(F7-T3) + spec(T3-T5) + spec(T5-O1) \right], \\
& \text{LP Spec} = \frac{1}{4} \left[
spec(Fp1-F3) + spec(F3-C3) + spec(C3-P3) + spec(P3-O1) \right], \\
& \text{RP Spec} = \frac{1}{4} \left[
spec(Fp2-F4) + spec(F4-C4) + spec(C4-P4) + spec(P4-O2) \right], \\
& \text{RR Spec} = \frac{1}{4} \left[
spec(Fp2-F8) + spec(F8-T4) + spec(T4-T6) + spec(T6-O2) \right]
\end{align*}
\end{footnotesize}
\begin{figure}[tbp]
\centering
\includegraphics[width=.55\textwidth]{inion}
\end{figure}
\end{frame}

\begin{frame}{EEG Signal Denosing using Wavelet Transform}
\begin{enumerate}
\item Calculates the mean absolute deviation of a signal $d$:
\begin{equation*}
\mathrm{maddest}(d, axis) = \frac{1}{N} \sum_{i=1}^N
\left\vert d_i - \frac{1}{N} \sum_{j=1}^N d_j \right\vert,
\end{equation*}
where $d_i$ is the value of the signal at index $i$, and
$N$ is the length of $d$.
\item Denoises a signal using wavelet decomposition:
\begin{enumerate}
\item Perform wavelet decomposition
\item Calculate the noise standard deviation using the last level of
coefficients
\item Compute the universal threshold
\item Apply a hard thresholding operation to the coefficients
\item Reconstruct the signal using the inverse wavelet transform
\end{enumerate}
\end{enumerate}
\end{frame}

\begin{frame}{Create Spectrograms from the Middle 50 Seconds}
\begin{figure}[tbp]
\centering
\includegraphics[width=.4\textwidth]{EEG_Signal}
\end{figure}
\begin{figure}[tbp]
\centering
\includegraphics[width=.5\textwidth]{EEG_Spectrogram}
\end{figure}
\end{frame}

\section[Model Structure]{Model Structure}

\begin{frame}{Model Structure}
\begin{small}
\begin{itemize}
\item An Unet CNN was designed to generated a feature map, and a global
potential for the label.
\item A hidden CRF layer will address the pixel wise label information of the
feature map, then assign a latent label $h_i$ at each pixel.
\item The latent label $h_i$ will contribute the global feature through the
following potetial.
\begin{equation*}
\varphi(y, h_{j}; \delta) = \sum_{a \in Y} \sum_{b \in H} \delta_{a, b}
\cdot \one(y = a) \cdot \one(h_j = b)
\end{equation*}
\item The jointly probability distribution of the model
\begin{equation*}
\begin{split}
\Phi(y, \hb, \xb; \theta) &= \underbrace{
\sum_{j \in \nu} \phi(h_j, x_j; \omega)
+ \sum_{i \neq j} \psi(h_i, h_j, x_i, x_j; \eta)}_{
% \propto \log\PP(\hb \vert \xb; \theta)
\textrm{Measures log-likelihood $\log\PP(\hb \vert \xb; \theta)$}} \\
&+ \underbrace{
\sum_{j \in \nu} \varphi(y, h_j, x_j; \delta) + \vartheta(y, \xb; \varpi)}_{
% \propto \log\PP(y | \hb,  \xb; \theta)
\textrm{Measures log-likelihood $\log\PP(y \vert \hb, \xb; \theta)$}}
\end{split}
\end{equation*}
\end{itemize}
\end{small}
\end{frame}

\begin{frame}{Model Inference}
\begin{small}
\begin{itemize}
\item The inference for the latent variable follows a standard mean-field
Variational Inference scheme.
\begin{equation*}
\QQ(h_i) \propto \exp\{\EE_{\QQ(\hb_{-i})}
\log\PP(y, h_i, \hb_{-i} \vert \xb; \theta)\}.
\end{equation*}
\item The resulting ELBO will be used as predicted log-likelihood of each image:
\begin{equation*}
\begin{split}
& \textrm{ELBO}(\QQ) = \EE_{\QQ(\hb)} \log\PP(y, \hb \vert \xb; \theta)
- \EE_{\QQ(\hb)} \log q(\hb) \\
&\quad = \sum_{i=1}^N \EE_{q_i(h_i)}
\left[ \phi(h_j, x_j; \omega) + \varphi(y, h_j, x_j; \delta) \right] \\
&\quad + \sum_{i=1}^N \sum_{j=1}^N \sum_{l,l^\prime} \one_{\{i \neq j\}} 
\QQ(h_i = l) \QQ(h_j = l^\prime) \mu(l, l^\prime) \\
&\quad \left[ \omega_1 \exp \left(
- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2}
- \frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib) \right\rvert^2}{2\eta_\beta^2}
\right) + \omega_2 \exp \left(
- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\gamma^2}
\right) \right] \\
&\quad - \sum_{i=1}^N \EE_{q_i} \log q_i(h_i) + \vartheta(y, \xb; 
\varpi).
\end{split}
\end{equation*}
\end{itemize}
\end{small}
\end{frame}

%\begin{frame}[allowframebreaks]
%\frametitle{References}
%\bibliographystyle{plain}
%\bibliography{bib}
%\end{frame}

\end{document}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2024-12-09 manuscript/slides.tex
edit introduction and model structure

@@
 
-\title[\textcolor{black}{UNet-CRF\_EEG}]{\large
-UNet-CRF Integration for Sequence Labeling on EEG Data}
+\title[\textcolor{black}{UNet-HCRF\_EEG}]{\large
+UNet-HCRF Integration for Sequence Labeling on EEG Data}
 
@@
 \section[Introduction]{Introduction}
+
+\begin{frame}{Introduction}
+\begin{itemize}
+\setlength{\itemsep}{1.5em}
+\item Sequence labeling tasks are crucial for healthcare, where accurate and
+timely data classification can significantly impact patient outcomes.
+\item Deep learning has recently emerged as a powerful tool for sequence
+labeling.
+\item Conditional random fields (CRFs) have been employed for modeling label
+dependencies, demonstrating promising results.
+\end{itemize}
+\end{frame}
+
+\section[Methods]{Methods}
+
+\begin{frame}{CNN-HCRF Model Structure}
+\begin{itemize}
+\item An UNet CNN was designed to generate a feature map, and a global
+potential for the label.
+\item A hidden conditional random field (HCRF) layer processes the pixel-wise
+label information from the feature map, assigning a latent label $h_i$ to each
+pixel.
+\item The latent label $h_i$ contributes to the global feature through the
+following potetial:
+\begin{equation*}
+\varphi(y, h_{j}; \delta) = \sum_{a \in Y} \sum_{b \in H} \delta_{a, b}
+\cdot \one(y = a) \cdot \one(h_j = b),
+\end{equation*}
+where $\delta_{a,b}$ denotes the likelihood of class label $y = a$
+containing a joint with hidden state $h_j = b$, learned during training.
+\end{itemize}
+\end{frame}
+
+\begin{frame}{CNN-HCRF Model Structure (cont.)}
+\begin{small}
+\begin{itemize}
+\item Unary potential $\phi(h_j, x_j; \omega)$: Measures the likelihood of the
+local feature $x_j$ is assigned as the hidden part state $h_j$, whose parameter
+$\omega$ is learned by end-to-end CNN-UNet structure.
+\item Binary potential $\psi(h_i, h_j, x_i, x_j; \eta)$: Balances the hidden
+label compatibility of the neighboring pixels $p$ on the feature image 
+$\Fb(\Ib)$:
+\begin{align*}
+\psi(h_i, h_j, x_i, x_j; \eta) &= \mu(h_i, h_j) \Bigg[
+\omega_1 \exp \left(
+-\frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2}
+-\frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib) \right\rvert^2}{2\eta_\beta^2}
+\right) \\
+&+ \omega_2 \exp \left(
+- \frac{\left\lvert p_i - p_j \right\rvert^2}{2 \eta_\gamma^2}
+\right)
+\Bigg],
+\end{align*}
+where the parameters $\eta_\alpha$, $\eta_\beta$ and $\eta_\gamma$ control the
+influence of the corresponding feature spaces.
+\item Global potential $\vartheta(y, \xb; \varpi)$: Measures the compatibility
+of class label $y$ and the global feature vector $x_0$ of the whole action,
+whose parameter $\varpi$ will be learned during training process.
+\end{itemize}
+\end{small}
+\end{frame}
+
+\begin{frame}{CNN-HCRF Model Structure (cont..)}
+Let $y$ represent the corresponding label of the input image $\xb$.
+The jointly probability distribution of the HCRF model is:
+\begin{align*}
+\Phi(y, \hb, \xb; \theta) &= \underbrace{
+\sum_{j \in \nu} \phi(h_j, x_j; \omega) +
+\sum_{i \neq j} \psi(h_i, h_j, x_i, x_j; \eta)}_{
+% \propto \log\PP( \hb \vert \xb; \theta)
+\textrm{Measures log-likelihood $\log\PP(\hb \vert \xb; \theta)$}} \\
+& + \underbrace{
+\sum_{j \in \nu} \varphi(y, h_j, x_j; \delta) + \vartheta(y, \xb; \varpi)}_{
+% \propto \log\PP(y | \hb, \xb; \theta)
+\textrm{Measures log-likelihood $\log\PP(y \vert \hb, \xb; \theta)$}}
+\end{align*}
+\end{frame}
+
+%\begin{frame}{Model Structure}
+%\begin{small}
+%\begin{itemize}
+%\item An Unet CNN was designed to generated a feature map, and a global
+%potential for the label.
+%\item A hidden CRF layer will address the pixel wise label information of the
+%feature map, then assign a latent label $h_i$ at each pixel.
+%\item The latent label $h_i$ will contribute the global feature through the
+%following potetial.
+%\begin{equation*}
+%\varphi(y, h_{j}; \delta) = \sum_{a \in Y} \sum_{b \in H} \delta_{a, b}
+%\cdot \one(y = a) \cdot \one(h_j = b)
+%\end{equation*}
+%\item The jointly probability distribution of the model
+%\begin{equation*}
+%\begin{split}
+%\Phi(y, \hb, \xb; \theta) &= \underbrace{
+%\sum_{j \in \nu} \phi(h_j, x_j; \omega)
+%+ \sum_{i \neq j} \psi(h_i, h_j, x_i, x_j; \eta)}_{
+%% \propto \log\PP(\hb \vert \xb; \theta)
+%\textrm{Measures log-likelihood $\log\PP(\hb \vert \xb; \theta)$}} \\
+%&+ \underbrace{
+%\sum_{j \in \nu} \varphi(y, h_j, x_j; \delta) + \vartheta(y, \xb; \varpi)}_{
+%% \propto \log\PP(y | \hb,  \xb; \theta)
+%\textrm{Measures log-likelihood $\log\PP(y \vert \hb, \xb; \theta)$}}
+%\end{split}
+%\end{equation*}
+%\end{itemize}
+%\end{small}
+%\end{frame}
+
+\begin{frame}{Model Inference}
+\begin{small}
+\begin{itemize}
+\item The inference for the latent variable follows a standard mean-field
+Variational Inference scheme.
+\begin{equation*}
+\QQ(h_i) \propto \exp\{\EE_{\QQ(\hb_{-i})}
+\log\PP(y, h_i, \hb_{-i} \vert \xb; \theta)\}.
+\end{equation*}
+\item The resulting ELBO will be used as predicted log-likelihood of each image:
+\begin{equation*}
+\begin{split}
+& \textrm{ELBO}(\QQ) = \EE_{\QQ(\hb)} \log\PP(y, \hb \vert \xb; \theta)
+- \EE_{\QQ(\hb)} \log q(\hb) \\
+&\quad = \sum_{i=1}^N \EE_{q_i(h_i)}
+\left[ \phi(h_j, x_j; \omega) + \varphi(y, h_j, x_j; \delta) \right] \\
+&\quad + \sum_{i=1}^N \sum_{j=1}^N \sum_{l,l^\prime} \one_{\{i \neq j\}} 
+\QQ(h_i = l) \QQ(h_j = l^\prime) \mu(l, l^\prime) \\
+&\quad \left[ \omega_1 \exp \left(
+- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2}
+- \frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib) \right\rvert^2}{2\eta_\beta^2}
+\right) + \omega_2 \exp \left(
+- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\gamma^2}
+\right) \right] \\
+&\quad - \sum_{i=1}^N \EE_{q_i} \log q_i(h_i) + \vartheta(y, \xb; 
+\varpi).
+\end{split}
+\end{equation*}
+\end{itemize}
+\end{small}
+\end{frame}
 
@@
 
-\section[Model Structure]{Model Structure}
-
-\begin{frame}{Model Structure}
-\begin{small}
-\begin{itemize}
-\item An Unet CNN was designed to generated a feature map, and a global
-potential for the label.
-\item A hidden CRF layer will address the pixel wise label information of the
-feature map, then assign a latent label $h_i$ at each pixel.
-\item The latent label $h_i$ will contribute the global feature through the
-following potetial.
-\begin{equation*}
-\varphi(y, h_{j}; \delta) = \sum_{a \in Y} \sum_{b \in H} \delta_{a, b}
-\cdot \one(y = a) \cdot \one(h_j = b)
-\end{equation*}
-\item The jointly probability distribution of the model
-\begin{equation*}
-\begin{split}
-\Phi(y, \hb, \xb; \theta) &= \underbrace{
-\sum_{j \in \nu} \phi(h_j, x_j; \omega)
-+ \sum_{i \neq j} \psi(h_i, h_j, x_i, x_j; \eta)}_{
-% \propto \log\PP(\hb \vert \xb; \theta)
-\textrm{Measures log-likelihood $\log\PP(\hb \vert \xb; \theta)$}} \\
-&+ \underbrace{
-\sum_{j \in \nu} \varphi(y, h_j, x_j; \delta) + \vartheta(y, \xb; \varpi)}_{
-% \propto \log\PP(y | \hb,  \xb; \theta)
-\textrm{Measures log-likelihood $\log\PP(y \vert \hb, \xb; \theta)$}}
-\end{split}
-\end{equation*}
-\end{itemize}
-\end{small}
-\end{frame}
-
-\begin{frame}{Model Inference}
-\begin{small}
-\begin{itemize}
-\item The inference for the latent variable follows a standard mean-field
-Variational Inference scheme.
-\begin{equation*}
-\QQ(h_i) \propto \exp\{\EE_{\QQ(\hb_{-i})}
-\log\PP(y, h_i, \hb_{-i} \vert \xb; \theta)\}.
-\end{equation*}
-\item The resulting ELBO will be used as predicted log-likelihood of each image:
-\begin{equation*}
-\begin{split}
-& \textrm{ELBO}(\QQ) = \EE_{\QQ(\hb)} \log\PP(y, \hb \vert \xb; \theta)
-- \EE_{\QQ(\hb)} \log q(\hb) \\
-&\quad = \sum_{i=1}^N \EE_{q_i(h_i)}
-\left[ \phi(h_j, x_j; \omega) + \varphi(y, h_j, x_j; \delta) \right] \\
-&\quad + \sum_{i=1}^N \sum_{j=1}^N \sum_{l,l^\prime} \one_{\{i \neq j\}} 
-\QQ(h_i = l) \QQ(h_j = l^\prime) \mu(l, l^\prime) \\
-&\quad \left[ \omega_1 \exp \left(
-- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2}
-- \frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib) \right\rvert^2}{2\eta_\beta^2}
-\right) + \omega_2 \exp \left(
-- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\gamma^2}
-\right) \right] \\
-&\quad - \sum_{i=1}^N \EE_{q_i} \log q_i(h_i) + \vartheta(y, \xb; 
-\varpi).
-\end{split}
-\end{equation*}
-\end{itemize}
-\end{small}
-\end{frame}
-
 %\begin{frame}[allowframebreaks]

### 2024-12-09 manuscript/slides.tex
edit methods

@@
 \begin{itemize}
-\item An UNet CNN was designed to generate a feature map, and a global
-potential for the label.
-\item A hidden conditional random field (HCRF) layer processes the pixel-wise
-label information from the feature map, assigning a latent label $h_i$ to each
-pixel.
-\item The latent label $h_i$ contributes to the global feature through the
-following potetial:
+\item Given the feature of the input data $\Ib$ learned by CNN, i.e.
+$\xb = \Fb(\Ib)$, its corresponding hidden part states $\hb$, and the class
+label $y$, a hidden conditional random field (HCRF) has the exponential form:
 \begin{equation*}
-\varphi(y, h_{j}; \delta) = \sum_{a \in Y} \sum_{b \in H} \delta_{a, b}
-\cdot \one(y = a) \cdot \one(h_j = b),
+\PP(y, \hb \vert \xb; \theta)
+= \frac{\exp\left(\Phi(y, \hb, \xb; \theta)\right)}
+{\sum_{y^\prime \in Y} \sum_{\hb \in H^N}
+\exp\left(\Phi(y^\prime, \hb, \xb; \theta)\right)},
 \end{equation*}
-where $\delta_{a,b}$ denotes the likelihood of class label $y = a$
-containing a joint with hidden state $h_j = b$, learned during training.
+where $\theta$ is the model parameter, $H^N$ denotes the set of all possible
+hidden part states of $N$ hidden parts, and $\Phi(y, \hb, \xb; \theta)$
+refers to potential function depending on the feature $\xb$.
+\item The probability of class label $y$ for the given feature $\xb$ is:
+\begin{equation*}
+\PP(y \vert \xb; \theta) = \sum_{\hb \in H^N} \PP(y, \hb \mid \xb; \theta)
+= \frac{\sum_{\hb \in H^N} \exp\left(\Phi(y, \hb, \xb; \theta)\right)}
+{\sum_{y^\prime \in Y} \sum_{\hb \in H^N}
+\exp\left(\Phi(y^\prime, \hb, \xb; \theta)\right)}.
+\end{equation*}
 \end{itemize}
@@
 \begin{frame}{CNN-HCRF Model Structure (cont.)}
-\begin{small}
-\begin{itemize}
-\item Unary potential $\phi(h_j, x_j; \omega)$: Measures the likelihood of the
-local feature $x_j$ is assigned as the hidden part state $h_j$, whose parameter
-$\omega$ is learned by end-to-end CNN-UNet structure.
-\item Binary potential $\psi(h_i, h_j, x_i, x_j; \eta)$: Balances the hidden
-label compatibility of the neighboring pixels $p$ on the feature image 
-$\Fb(\Ib)$:
-\begin{align*}
-\psi(h_i, h_j, x_i, x_j; \eta) &= \mu(h_i, h_j) \Bigg[
-\omega_1 \exp \left(
--\frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2}
--\frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib) \right\rvert^2}{2\eta_\beta^2}
-\right) \\
-&+ \omega_2 \exp \left(
-- \frac{\left\lvert p_i - p_j \right\rvert^2}{2 \eta_\gamma^2}
-\right)
-\Bigg],
-\end{align*}
-where the parameters $\eta_\alpha$, $\eta_\beta$ and $\eta_\gamma$ control the
-influence of the corresponding feature spaces.
-\item Global potential $\vartheta(y, \xb; \varpi)$: Measures the compatibility
-of class label $y$ and the global feature vector $x_0$ of the whole action,
-whose parameter $\varpi$ will be learned during training process.
-\end{itemize}
-\end{small}
-\end{frame}
-
-\begin{frame}{CNN-HCRF Model Structure (cont..)}
-Let $y$ represent the corresponding label of the input image $\xb$.
 The jointly probability distribution of the HCRF model is:
@@
 & + \underbrace{
-\sum_{j \in \nu} \varphi(y, h_j, x_j; \delta) + \vartheta(y, \xb; \varpi)}_{
+\sum_{j \in \nu} \varphi(y, h_j, x_j; \delta) + \vartheta(y, x_0; \varpi)}_{
 % \propto \log\PP(y | \hb, \xb; \theta)
-\textrm{Measures log-likelihood $\log\PP(y \vert \hb, \xb; \theta)$}}
+\textrm{Measures log-likelihood $\log\PP(y \vert \hb, \xb; \theta)$}},
 \end{align*}
+where:
+\begin{itemize}
+\item Unary potential $\phi(h_j, x_j; \omega)$: Measures the likelihood of the
+local feature $x_j$ is assigned as the hidden part state $h_j$, with parameter
+$\omega$ is learned by end-to-end CNN-UNet structure.
+\end{itemize}
+\end{frame}
+
+\begin{frame}{CNN-HCRF Model Structure (cont..)}
+\begin{itemize}
+\item Binary potential $\psi(h_i, h_j, x_i, x_j; \eta)$: Balances the hidden
+label compatibility of the neighboring pixels $p$ on the feature $\xb$:
+\begin{align*}
+\psi(h_i, h_j, x_i, x_j; \eta) &= \mu(h_i, h_j) \Bigg[
+\omega_1 \exp \left(
+-\frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2}
+-\frac{\left\lvert x_i - x_j \right\rvert^2}{2\eta_\beta^2}
+\right) \\
+&+ \omega_2 \exp \left(
+- \frac{\left\lvert p_i - p_j \right\rvert^2}{2 \eta_\gamma^2}
+\right)
+\Bigg],
+\end{align*}
+where $\mu(h_i, h_j)$ is a label compatibility function,
+%that describes the interactive influences between different pairs of classes.
+$\omega_1$ and $\omega_2$ are linear combination weights of different kernels,
+and the parameters $\eta_\alpha$, $\eta_\beta$ and $\eta_\gamma$ control the
+influence of the corresponding feature spaces.
+\end{itemize}
+\end{frame}
+
+\begin{frame}{CNN-HCRF Model Structure (cont...)}
+\begin{itemize}
+\item Unary potential $\varphi(y, h_{j}; \delta)$: Measures the compatibility
+between global class label $y$ and the hidden local state $h_j$:
+\begin{equation*}
+\varphi(y, h_{j}; \delta) = \sum_{a \in Y} \sum_{b \in H} \delta_{a, b}
+\cdot \one(y = a) \cdot \one(h_j = b),
+\end{equation*}
+where $\one(\cdot)$ is a indicator function, and $\delta_{a,b}$ denotes the
+likelihood of class label $y = a$ containing a joint with hidden state
+$h_j = b$, learned during training.
+\item Global potential $\vartheta(y, x_0; \varpi)$: Measures the likelihood of
+the global feature $x_0$ is assigned as the action label $y$, with parameter
+$\varpi$ learned during training process.
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Variational Inference of HCRF}
+\begin{small}
+\begin{itemize}
+\item The mean-field varepsilon family approximation of
+$\log P(\hb \mid y,\xb,\theta)$ is expressed as
+$\QQ(h) = \prod_{i=1}^N q_i(h_i)$.
+\item The evidence lower bound (ELBO) under the mean-field family is:
+\begin{equation*}
+\begin{split}
+& \textrm{ELBO}(\QQ) = \EE_{\QQ(\hb)} \log\PP(y, \hb \vert \xb; \theta)
+- \EE_{\QQ(\hb)} \log q(\hb) \\
+&\quad = \sum_{i=1}^N \EE_{q_i(h_i)}
+\left[ \phi(h_j, x_j; \omega) + \varphi(y, h_j, x_j; \delta) \right] \\
+&\quad + \sum_{i=1}^N \sum_{j=1}^N \sum_{l,l^\prime} \one_{\{i \neq j\}} 
+\QQ(h_i = l) \QQ(h_j = l^\prime) \mu(l, l^\prime) \\
+&\quad \left[ \omega_1 \exp \left(
+- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2}
+- \frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib) \right\rvert^2}{2\eta_\beta^2}
+\right) + \omega_2 \exp \left(
+- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\gamma^2}
+\right) \right] \\
+&\quad - \sum_{i=1}^N \EE_{q_i} \log q_i(h_i) + \vartheta(y, \xb; 
+\varpi).
+\end{split}
+\end{equation*}
+\end{itemize}
+\end{small}
+\end{frame}
+
+\begin{frame}{Variational Inference of HCRF (cont.)}
+\begin{itemize}
+\item By the coordinate ascent variational inference(CAVI) algorithm, fixing
+$q_j(h_j), \forall j \neq i$, the optimal $q_i(h_i)$ that maximizes
+$\mathrm{ELBO}(q_i(h_i))$ is:
+\begin{equation*}
+\QQ(h_i) \propto \exp\{\EE_{\QQ(\hb_{-i})}
+\log\PP(y, h_i, \hb_{-i} \mid \xb; \theta)\}.
+\end{equation*}
+\item The explicit form of $q_i(l)$ is derived as follows:
+\begin{equation*}
+\begin{split}
+& q_i(l) = \frac{1}{Z_i} \exp \Bigg\{
+\phi(h_j, x_j; \omega) + \varphi(y, h_j, x_j; \delta)
++ \sum_{j \neq i} \sum_{l^\prime} q(l^\prime) \mu(l, l^\prime) \\
+& \left[ \omega_1 \exp \left(
+- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2}
+- \frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib) \right\rvert^2}{2\eta_\beta^2}
+\right)
++ \omega_2 \exp \left(
+- \frac{\left\lvert p_i - p_j 
+\right\rvert^2}{2\eta_\gamma^2}
+\right)
+\right]
+\Bigg\},
+\end{split}
+\end{equation*}
+where $Z_i$ is the normalization constant.
+\end{itemize}
 \end{frame}
@@
 
-\begin{frame}{Model Inference}
-\begin{small}
-\begin{itemize}
-\item The inference for the latent variable follows a standard mean-field
-Variational Inference scheme.
-\begin{equation*}
-\QQ(h_i) \propto \exp\{\EE_{\QQ(\hb_{-i})}
-\log\PP(y, h_i, \hb_{-i} \vert \xb; \theta)\}.
-\end{equation*}
-\item The resulting ELBO will be used as predicted log-likelihood of each image:
-\begin{equation*}
-\begin{split}
-& \textrm{ELBO}(\QQ) = \EE_{\QQ(\hb)} \log\PP(y, \hb \vert \xb; \theta)
-- \EE_{\QQ(\hb)} \log q(\hb) \\
-&\quad = \sum_{i=1}^N \EE_{q_i(h_i)}
-\left[ \phi(h_j, x_j; \omega) + \varphi(y, h_j, x_j; \delta) \right] \\
-&\quad + \sum_{i=1}^N \sum_{j=1}^N \sum_{l,l^\prime} \one_{\{i \neq j\}} 
-\QQ(h_i = l) \QQ(h_j = l^\prime) \mu(l, l^\prime) \\
-&\quad \left[ \omega_1 \exp \left(
-- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2}
-- \frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib) \right\rvert^2}{2\eta_\beta^2}
-\right) + \omega_2 \exp \left(
-- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\gamma^2}
-\right) \right] \\
-&\quad - \sum_{i=1}^N \EE_{q_i} \log q_i(h_i) + \vartheta(y, \xb; 
-\varpi).
-\end{split}
-\end{equation*}
-\end{itemize}
-\end{small}
-\end{frame}
+%\begin{frame}{Model Inference}
+%\begin{small}
+%\begin{itemize}
+%\item The inference for the latent variable follows a standard mean-field
+%Variational Inference scheme.
+%\begin{equation*}
+%\QQ(h_i) \propto \exp\{\EE_{\QQ(\hb_{-i})}
+%\log\PP(y, h_i, \hb_{-i} \vert \xb; \theta)\}.
+%\end{equation*}
+%\item The resulting ELBO will be used as predicted log-likelihood of each 
+%image:
+%\begin{equation*}
+%\begin{split}
+%& \textrm{ELBO}(\QQ) = \EE_{\QQ(\hb)} \log\PP(y, \hb \vert \xb; \theta)
+%- \EE_{\QQ(\hb)} \log q(\hb) \\
+%&\quad = \sum_{i=1}^N \EE_{q_i(h_i)}
+%\left[ \phi(h_j, x_j; \omega) + \varphi(y, h_j, x_j; \delta) \right] \\
+%&\quad + \sum_{i=1}^N \sum_{j=1}^N \sum_{l,l^\prime} \one_{\{i \neq j\}} 
+%\QQ(h_i = l) \QQ(h_j = l^\prime) \mu(l, l^\prime) \\
+%&\quad \left[ \omega_1 \exp \left(
+%- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2}
+%- \frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib) \right\rvert^2}{2\eta_\beta^2}
+%\right) + \omega_2 \exp \left(
+%- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\gamma^2}
+%\right) \right] \\
+%&\quad - \sum_{i=1}^N \EE_{q_i} \log q_i(h_i) + \vartheta(y, \xb; 
+%\varpi).
+%\end{split}
+%\end{equation*}
+%\end{itemize}
+%\end{small}
+%\end{frame}
 

### 2024-12-09 manuscript/slides.tex
edit experiment

@@
 
-\section[Data Preprocessing]{Data Preprocessing}
+\section[Experiment]{Experiment}
+
+\begin{frame}{Harmful Brain Activity Classification Dataset}
+\begin{itemize}
+\item The dataset contains 50-second long electroencephalography (EEG) samples
+covering a 10-minute window as well as spectrograms matching each sample.
+\bigskip
+\item The EEG samples were labeled by expert annotators to determine which of
+the six features each spectrogram belonged to.
+\bigskip
+\item The longer 10-minute window allows for the model to pick up on further
+insights regarding the nature of the brain activity.
+\end{itemize}
+\end{frame}
 
@@
 \begin{itemize}
-\item Extract the 10 min long Kaggle Spectrogram centered at the middle time
-where the experts rating (Global EEGs)
-\item Transform the 50s long EEG waveforms into Spectrograms
+\item Extract 10-minute long spectrograms centered at the midpoint time,
+where the expert ratings (Global EEGs) were provided.
+\bigskip
+\item Convert the 50-second long EEG waveforms into spectrograms.
 \end{itemize}
@@
 
-\begin{frame}{EEG signal denosing using Wavelet transform}
+\begin{frame}{Make Spectrogram from EEG}
+\begin{itemize}
+\item Bipolar Montage combines multiple EEG electrode signals by calculating
+the difference between adjacent electrodes, yielding signals representing
+specific brain regions.
+\end{itemize}
+\begin{figure}[tbp]
+\centering
+\includegraphics[width=.7\textwidth]{inion}
+\end{figure}
+\end{frame}
+
+\begin{frame}{Make Spectrogram from EEG (cont.)}
+\begin{itemize} 
+\item A single time series signal is required to generate a spectrogram.
+To create four spectrograms from the dataset's 19 EEG time series signals,
+we employ Bipolar Montage to combine them into four representative signals.
+\end{itemize}
 \begin{footnotesize}
@@
 & \text{RR Spec} = \frac{1}{4} \left[
-spec(Fp2-F8) + spec(F8-T4) + spec(T4-T6) + spec(T6-O2) \right]
+spec(Fp2-F8) + spec(F8-T4) + spec(T4-T6) + spec(T6-O2) \right],
 \end{align*}
 \end{footnotesize}
+where $spec(\cdot)$ is a custom function implemented using the \pkg{pywt} and
+\pkg{librosa} packages for signal transformation.
+\end{frame}
+
+%\begin{frame}{EEG Signal Denosing using Wavelet Transform}
+%\begin{enumerate}
+%\item Calculates the mean absolute deviation of a signal $d$:
+%\begin{equation*}
+%\mathrm{maddest}(d, axis) = \frac{1}{N} \sum_{i=1}^N
+%\left\vert d_i - \frac{1}{N} \sum_{j=1}^N d_j \right\vert,
+%\end{equation*}
+%where $d_i$ is the value of the signal at index $i$, and
+%$N$ is the length of $d$.
+%\item Denoises a signal using wavelet decomposition:
+%\begin{enumerate}
+%\item Perform wavelet decomposition
+%\item Calculate the noise standard deviation using the last level of
+%coefficients
+%\item Compute the universal threshold
+%\item Apply a hard thresholding operation to the coefficients
+%\item Reconstruct the signal using the inverse wavelet transform
+%\end{enumerate}
+%\end{enumerate}
+%\end{frame}
+
+\begin{frame}{Make Spectrogram from EEG (cont..)}
 \begin{figure}[tbp]
 \centering
-\includegraphics[width=.55\textwidth]{inion}
+\vspace{-.5em}
+\includegraphics[width=.5\textwidth]{EEG_Signal}
 \end{figure}
-\end{frame}
-
-\begin{frame}{EEG Signal Denosing using Wavelet Transform}
-\begin{enumerate}
-\item Calculates the mean absolute deviation of a signal $d$:
-\begin{equation*}
-\mathrm{maddest}(d, axis) = \frac{1}{N} \sum_{i=1}^N
-\left\vert d_i - \frac{1}{N} \sum_{j=1}^N d_j \right\vert,
-\end{equation*}
-where $d_i$ is the value of the signal at index $i$, and
-$N$ is the length of $d$.
-\item Denoises a signal using wavelet decomposition:
-\begin{enumerate}
-\item Perform wavelet decomposition
-\item Calculate the noise standard deviation using the last level of
-coefficients
-\item Compute the universal threshold
-\item Apply a hard thresholding operation to the coefficients
-\item Reconstruct the signal using the inverse wavelet transform
-\end{enumerate}
-\end{enumerate}
-\end{frame}
-
-\begin{frame}{Create Spectrograms from the Middle 50 Seconds}
+\vspace{-1em}
 \begin{figure}[tbp]
 \centering
-\includegraphics[width=.4\textwidth]{EEG_Signal}
-\end{figure}
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=.5\textwidth]{EEG_Spectrogram}
+\includegraphics[width=.55\textwidth]{EEG_Spectrogram}
 \end{figure}

### 2024-12-09 manuscript/slides.tex
edit intro; add framework

@@
 \setlength{\itemsep}{1.5em}
-\item Sequence labeling tasks are crucial for healthcare, where accurate and
-timely data classification can significantly impact patient outcomes.
+\item Electroencephalography (EEG) pattern recognition tasks are crucial for
+healthcare, where accurate and timely data classification can significantly
+impact patient outcomes.
 \item Deep learning has recently emerged as a powerful tool for sequence
@@
 \end{itemize}
+\end{frame}
+
+\begin{frame}{Framework}
+\begin{figure}[tbp]
+\centering
+\includegraphics[width=\textwidth]{model}
+\end{figure}
 \end{frame}
@@
 \begin{split}
-& \textrm{ELBO}(\QQ) = \EE_{\QQ(\hb)} \log\PP(y, \hb \vert \xb; \theta)
+& \mathrm{ELBO}(\QQ) = \EE_{\QQ(\hb)} \log\PP(y, \hb \vert \xb; \theta)
 - \EE_{\QQ(\hb)} \log q(\hb) \\
@@
 \begin{itemize}
-\item By the coordinate ascent variational inference(CAVI) algorithm, fixing
+\item By the coordinate ascent variational inference (CAVI) algorithm, fixing
 $q_j(h_j), \forall j \neq i$, the optimal $q_i(h_i)$ that maximizes
@@
 \begin{itemize}
-\item The dataset contains 50-second long electroencephalography (EEG) samples
-covering a 10-minute window as well as spectrograms matching each sample.
+\item The dataset contains 50-second long EEG samples covering a 10-minute
+window as well as spectrograms matching each sample.
 \bigskip

### 2024-12-09 manuscript/slides_results.tex
Add files via upload

--- 2024-12-09 manuscript/slides.tex
@@
 \usepackage{tikz}
-
+% \usepackage{algorithm}
+% \usepackage{algpseudocode}
 %% notations; only define those you use frequently
@@
 
-\title[\textcolor{black}{UNet-HCRF\_EEG}]{\large
-UNet-HCRF Integration for Sequence Labeling on EEG Data}
+\title[\textcolor{black}{UNet-CRF\_EEG}]{\large
+UNet-CRF Integration for Sequence Labeling on EEG Data}
 
@@
 \centering
-\includegraphics[width=.7\textwidth]{inion}
+% \includegraphics[width=.7\textwidth]{inion}
 \end{figure}
@@
 \vspace{-.5em}
-\includegraphics[width=.5\textwidth]{EEG_Signal}
+% \includegraphics[width=.5\textwidth]{EEG_Signal}
 \end{figure}
@@
 \centering
-\includegraphics[width=.55\textwidth]{EEG_Spectrogram}
+% \includegraphics[width=.55\textwidth]{EEG_Spectrogram}
 \end{figure}
+\end{frame}
+\begin{frame}
+	\frametitle{Mean Field Inference In Hidden CRF }
+	\begin{algorithm}[H]
+		\caption{Mean-field inference in fully-connected CRF}
+		\KwIn{$Q_i(h_i) \gets U_i(h_i), \; i = 1, 2, \dots, N$ \tcp*[f]{Initialize $Q(h)$}}
+		\While{not reach max iteration number}{
+		    $\hat{Q}_i^{(m)}(h_i) \gets \sum_{j \neq i} k^{(m)}(f_i, f_j) Q_j(h_i)$ for all $m$ \tcp*[f]{Message Passing}\;
+		    $\tilde{Q}_i^{(m)}(h_i) \gets \sum_{l \in \mathcal{L}} \mu^{(m)}(h_i, l) \sum_m \omega^{(m)} \hat{Q}_i^{(m)}(l)$ \tcp*[f]{Compatibility Transform}\;
+		    $Q_i(h_i) \gets \exp \{\phi(h_j, x_j; \omega) + \varphi(y, h_j, x_j; \delta) - \tilde{Q}_i(x_i)\}$ \tcp*[f]{Local Update}\;
+		    Normalize $Q_i(h_i)$\;
+		}
+		\end{algorithm}
+\end{frame}
+
+\begin{frame}
+	\frametitle{Training Schedule}
+	The loss function was designed as Kullback-leibler (kl) divergence.	
+	Pre-train Unet learning rate schedule
+	% \vspace{-1em}
+	\begin{figure}[tbp]
+	\centering
+	\includegraphics[width=.55\textwidth]{pretrain}
+	\end{figure}
+	Fine tuning HCRF  learning rate schedule
+	% \vspace{-1em}
+	\begin{figure}[tbp]
+	\centering
+	\includegraphics[width=.55\textwidth]{finetune}
+	\end{figure}
+\end{frame}
+
+\section{Results and Conclusion}
+
+\begin{frame}
+	\frametitle{Results}
+	We evaluate our model by 5 fold cross-validation.
+	\begin{itemize}
+		\item Our model achieves a \textbf{0.4992} KL-divergence.
+		\item The baseline model from Kaggle using EfficientNet obtains a 0.59 cross-validation KL-divergence.
+	\end{itemize}
+	However, there still are many teams on Kaggle achieve even lower KL-divergence(0.2-0.3) by using large model like Vision Transformers(ViT).
+\end{frame}
+\begin{frame}
+	\frametitle{Conclusion}
+	\begin{itemize}
+		\item We build a Hidden Conditional Random Fields for EEG based Spectrograms image classification.
+		\item The latent CRF layer is desgined to capture the latent segmentation information of the output of Neural Networks.
+		\item A variational mean-field inference scheme is used to address the posterior distribution.
+		\item We integrate the Unet and HiddenCRF into an end-to-end training structure. The Neural Networks feature learning and 
+		variational inference can be accelerated jointly by GPUs.
+	\end{itemize}
 \end{frame}

### 2024-12-09 manuscript/slides.tex
add results

@@
 
+\begin{frame}{Mean Field Inference in HCRF}
+\centering
+\resizebox{.95\textwidth}{!}{
+\begin{algorithm}[H]
+\caption{Mean-field inference in fully-connected CRF}
+\KwIn{$Q_i(h_i) \gets U_i(h_i), \; i = 1, 2, \dots, N$
+\tcp*[f]{Initialize $Q(h)$}}
+\While{not reach max iteration number}{
+$\hat{Q}_i^{(m)}(h_i) \gets \sum_{j \neq i} k^{(m)}(f_i, f_j) Q_j(h_i)$ for all
+$m$ \tcp*[f]{Message Passing}\;
+$\tilde{Q}_i^{(m)}(h_i) \gets \sum_{l \in \mathcal{L}} \mu^{(m)}(h_i, l)
+\sum_m \omega^{(m)} \hat{Q}_i^{(m)}(l)$ \tcp*[f]{Compatibility Transform}\;
+$Q_i(h_i) \gets \exp \{\phi(h_j, x_j; \omega) + \varphi(y, h_j, x_j; \delta)
+- \tilde{Q}_i(x_i)\}$ \tcp*[f]{Local Update}\;
+Normalize $Q_i(h_i)$\;
+}
+\end{algorithm}}
+\end{frame}
+
 %\begin{frame}{Model Structure}
@@
 
+\begin{frame}
+\frametitle{Training Schedule}
+The loss function was designed as Kullback-Leibler (KL) divergence.
+\begin{figure}[tbp]
+\centering
+\includegraphics[width=.55\textwidth]{pretrain}
+\vspace{-.3em}
+\caption{Pre-train UNet learning rate schedule.}
+\end{figure}
+\vspace{-1.2em}
+\begin{figure}[tbp]
+\centering
+\includegraphics[width=.55\textwidth]{finetune}
+\vspace{-.3em}
+\caption{Fine tuning HCRF learning rate schedule.}
+\end{figure}
+\end{frame}
+
+\section{Results and Conclusion}
+
+\begin{frame}{Results}
+We evaluate our model by 5 fold cross-validation.
+\bigskip
+\begin{itemize}
+\item Our model achieves a \textbf{0.4992} KL-divergence.
+\bigskip
+\item The baseline model from Kaggle using EfficientNet obtains a 0.59
+cross-validation KL-divergence.
+\end{itemize}
+\bigskip
+However, there still are many teams on Kaggle have achieved even lower
+KL-divergence scores (0.2-0.3) by leveraging the power of large models
+like Vision Transformers (ViT).
+\end{frame}
+
+\begin{frame}{Conclusion}
+\begin{itemize}
+\item We build a Hidden Conditional Random Fields for EEG based Spectrograms
+image classification.
+\bigskip
+\item The latent CRF layer is desgined to capture the latent segmentation
+information of the output of Neural Networks.
+\bigskip
+\item A variational mean-field inference scheme is used to address the
+posterior distribution.
+\bigskip
+\item We integrate the Unet and HiddenCRF into an end-to-end training structure.
+The Neural Networks feature learning and variational inference can be
+accelerated jointly by GPUs.
+\end{itemize}
+\end{frame}
+
 %\begin{frame}[allowframebreaks]

### 2024-12-09 manuscript/slides.tex
revise title
typo

@@
 
-\title[\textcolor{black}{UNet-HCRF\_EEG}]{\large
-UNet-HCRF Integration for Sequence Labeling on EEG Data}
+\title[\textcolor{black}{UNet-HCRF\_EEG}]{
+UNet-HCRF Integration for EEG Pattern Recognition}
 
@@
 \bigskip
-However, there still are many teams on Kaggle have achieved even lower
+However, many teams on Kaggle have achieved even lower
 KL-divergence scores (0.2-0.3) by leveraging the power of large models

### 2024-12-11 manuscript/main.tex
edit methods

@@
 \newcommand{\EE}{{\mathbb{E}}}
+\newcommand{\Fb}{\mathbf{F}}
+\newcommand{\hb}{\mathbf{h}}
+\newcommand{\Hc}{\mathcal{H}}
+\newcommand{\Ib}{\mathbf{I}}
+\newcommand{\Lc}{\mathcal{L}}
 \newcommand{\PP}{{\mathbb{P}}}
 \newcommand{\QQ}{{\mathbb{Q}}}
-\newcommand{\Fb}{\mathbf{F}}
-\newcommand{\Ib}{\mathbf{I}}
-\newcommand{\hb}{\mathbf{h}}
 \newcommand{\xb}{\mathbf{x}}
+\newcommand{\Oc}{\mathcal{O}}
 \newcommand{\one}{\mathbbm{1}}
 \newcommand{\Ocal}{\mathcal{O}}
+\newcommand{\RR}{{\mathbb{R}}}
+
+\let\oldsubsubsection\subsubsection
+\renewcommand{\subsubsection}[1]{\oldsubsubsection{\textbf{#1}}}
 
@@
 
-\title{\huge UNet-CRF Integration for Sequence Labeling on EEG Data}
+\title{\huge UNet-HCRF Integration for EEG Pattern Recognition}
 
@@
 
-probabilistic graphical model, UNet, fully connected CRF,
+classification, probabilistic graphical model, spectrogram,
+variational inference, wavelet transformation
 
@@
 %Introduction: problem definition and motivation.
+
+\sx{EEG tasks}
 
@@
 \section{Methods}
-%Methods: overview of your proposed method, Intuition on why should it be
-%better than the state of the art, and details of models and algorithms that
-%you developed.
+
+We begin by introducing the necessary notations.
+Let $\xb = (x_0, x_1, \dots, x_N)^\top$ represent the feature vector of the
+input image, where $x_0$ denotes the global feature capturing overall image
+characteristics, and $x_i$ ($1 \leq i \leq N$) corresponds to a local feature
+extracted from the $i$-th pixel. Here, $N$ denotes the total number of pixels
+in the image. Let $y \in Y$ represent the class label associated with the image,
+where $Y$ is a finite set containing $\vert Y \vert$ possible class labels.
+For any feature $\xb$, a corresponding vector of unobserved variables, referred
+to as hidden part states and denoted by $\hb = (h_1, \dots, h_N)^\top$,
+is introduced. Each $h_i \in H$ represents the hidden part state associated
+with $x_i$ for $i = 1, \dots, N$. Here, $H$ is a finite set of hidden part
+states, and $\vert H \vert$ denotes the cardinality of this set.
 
@@
 
-The classical hidden conditional random field (HCRF) model can be formulated as
+Following the theory of random fields as outlined
+by~\citet{quattoni2004conditional} and~\citet{wang2006hidden},
+given the feature $\xb$, its corresponding hidden part states $\hb$,
+and the class label $y$, a hidden conditional random field (HCRF) can be
+expressed in exponential form as follows:
 \begin{equation}
+\label{eq:hcrf}
+\PP(y, \hb \mid \xb; \theta)
+= \frac{\exp\left(\Phi(y, \hb, \xb; \theta)\right)}
+{\sum_{y^\prime \in Y} \sum_{\hb \in H^N}
+\exp\left(\Phi(y^\prime, \hb, \xb; \theta)\right)},
+\end{equation}
+where $\theta$ is the parameter of the model,
+$H^N$ denotes the set of all possible hidden part states of $N$ hidden parts,
+and $\Phi(y, \hb, \xb; \theta) \in \RR$ refers to a potential function
+depending on the feature $\xb$ and parameterized by $\theta$.
+
+From Equation~\ref{eq:hcrf}, the probability of class label $y$ for the given
+feature $\xb$ is:
+\begin{equation*}
 \begin{split}
-\PP(y \vert \xb; \theta) &= \sum_{\hb \in H^N}
-\mathbb{P}(y, \hb \vert \xb; \theta) \\
+\PP(y \vert \xb; \theta) &= \sum_{\hb \in H^N} \PP(y, \hb \mid \xb; \theta) \\
 &= \frac{\sum_{\hb \in H^N} \exp\left(\Phi(y, \hb, \xb; \theta)\right)}
-{\sum_{y^\prime} \sum_{\hb \in H^N}
+{\sum_{y^\prime \in Y} \sum_{\hb \in H^N}
 \exp\left(\Phi(y^\prime, \hb, \xb; \theta)\right)}.
 \end{split}
-\end{equation}
-
-Different to the design of most existing HCRF models, which they assign
-potentials based on the features predefined from the input data $\Ib$.
-%$\PP(y, \hb \vert \xb; \theta)$.
-%Here, we consider a generative model that constructing
-%$\PP(\hb \vert \xb; \theta)$ through a CRF framework then using an 
-%unary potential for , 
-Our HCRF replace $I$ with features learned by CNN, i.e. $\Fb(\Ib)$. 
-Thus the new potentials become:
-\begin{equation*}
-\Phi(y, \hb, \xb; \theta) = \Phi(y, \hb, \Fb(\Ib); \theta).
-%\PP(y, \hb \vert \xb; \theta) = \PP(y \vert \hb, \xb; \theta)
-%\cdot \PP(\hb \vert \xb; \theta).
-\end{equation*}
-Thus, we design the potential of our HCRF model as.
-\begin{equation}
-\begin{split}
-\Phi(y, \hb, \xb; \theta) &= \underbrace{
-\sum_{j \in \nu} \phi(h_j, x_j; \omega)
-+ \sum_{i \neq j} \psi(h_i, h_j, x_i, x_j; \eta)}_{
-% \propto \log\PP(\hb \vert \xb; \theta)
-\textrm{Measures log-likelihood $\log\PP(\hb \vert \xb; \theta)$}} \\
-&+ \underbrace{
-\sum_{j \in \nu} \varphi(y, h_j, x_j; \delta) + \vartheta(y, \xb; \varpi)}_{
-% \propto \log\PP(y | \hb,  \xb; \theta)
-\textrm{Measures log-likelihood $\log\PP(y \vert \hb, \xb; \theta)$}}
-\end{split}
-\end{equation}
-
-{\textbf{Unary Potential}}
-$\phi(h_j, x_j; \omega)$ measures the likelihood of the local feature
-$x_j$ is assigned as the hidden attention state $h_j$.
-We generate the likelihood through a softmax operation on $\Fb(\xb)$
-the feature maps of CNN with $(N, C_{\textrm{out}})$ dimensions,
-by which the potential of each pixel $x_j$ assigned as each state of
-hidden state set $\{0, 1, \dots, 5\}$ is drawn and represented by
-a matrix with $(N, 2)$ dimension.
-
-The parameter $\omega$ is learned by our end-to-end CNN-Unet training 
-structure.
-
-{\textbf{Unary Potential}}
-$\varphi(y, h_j; \delta)$ measures the compatibility between global
-class label $y$ and the hidden local state $h_j$. This potential is
-parametrized as:
-\begin{equation*}
-\varphi(y, h_{j}; \delta) = \sum_{a \in Y} \sum_{b \in H} \delta_{a, b}
-\cdot \one(y = a) \cdot \one(h_j = b).
 \end{equation*}
 
-{\textbf{Binary Potential}}
-$\psi(h_i, h_j, x_i, x_j; \eta)$
-balances the hidden label compatibility of the neighboring pixels on the 
-feature image $\Fb(\Ib)$.
-We define this potential through a common contrast-sensitive two-kernal 
+The potential of the HCRF model, $\Phi(y, \hb, \xb; \theta)$, is defined in the
+following form:
+\begin{align*}
+\Phi(y, \hb, \xb; \theta) &= \underbrace{
+\sum_{j \in \nu} \phi(h_j, x_j; \omega) +
+\sum_{i \neq j} \psi(h_i, h_j, x_i, x_j; \eta)}_{
+% \propto \log \PP( \hb \vert \xb; \theta)
+\textrm{Measures log-likelihood $\log \PP(\hb \vert \xb; \theta)$}} \\
+& + \underbrace{
+\sum_{j \in \nu} \varphi(y, h_j, x_j; \delta) + \vartheta(y, x_0; \varpi)}_{
+% \propto \log \PP(y | \hb, \xb; \theta)
+\textrm{Measures log-likelihood $\log \PP(y \vert \hb, \xb; \theta)$}},
+\end{align*}
+where the detailed components are described below.
+
+\textbf{Unary potential} $\phi(h_j, x_j; \omega)$ measures the likelihood of
+the local feature $x_j$ is assigned as the hidden attention state $h_j$.
+The parameter $\omega$ is learned through the end-to-end CNN-UNet training
+structure. This likelihood is computed by applying a softmax operation to the
+feature maps produced by the CNN.
+
+\textbf{Binary potential} $\psi(h_i, h_j, x_i, x_j; \eta)$ balances the
+hidden label compatibility of the neighboring pixels on the feature $\xb$.
+The potential is defined through a common contrast-sensitive two-kernal
 potentials~\citep{krahenbuhl2011efficient, chen2022end}:
-\begin{equation}
+\begin{equation*}
 \begin{split}
@@
 &\quad
-- \frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib)\right\rvert^2}{2\eta_\beta^2}
-\bigg)
-+ \omega_2 \exp \left(
+- \frac{\left\lvert x_i - x_j\right\rvert^2}{2\eta_\beta^2}
+\bigg) + \omega_2 \exp \left(
 - \frac{\left\lvert p_i - p_j \right\rvert^2}{2 \eta_\gamma^2}
 \right)
-\Bigg]
+\Bigg],
 \end{split}
-\end{equation}
+\end{equation*}
+where $\mu(h_i, h_j)$ is a label compatibility function that describes
+the interactive influences between different pairs of classes,
+and $p$ refers to the pixel position.
+The parameters $\omega_1$ and $\omega_2$ serve as linear combination weights,
+controlling the contributions of the first and second kernels, respectively.
+And the parameters $\eta_\alpha$, $\eta_\beta$ and $\eta_\gamma$ control the
+influence of the corresponding feature spaces.
+
+\textbf{Unary potential} $\varphi(y, h_j; \delta)$ measures the compatibility
+between global class label $y$ and the hidden local state $h_j$.
+This potential is parametrized as:
+\begin{equation*}
+\varphi(y, h_{j}; \delta) = \sum_{a \in Y} \sum_{b \in H} \delta_{a, b}
+\cdot \one(y = a) \cdot \one(h_j = b).
+\end{equation*}
+where $\one(\cdot)$ is an indicator function, and $\delta_{a,b}$ denotes the
+likelihood of class label $y = a$ containing a joint with hidden state
+$h_j = b$, which is learned during training.
+
+\textbf{Global potential} $\vartheta(y, x_0; \varpi)$ measures the likelihood of
+the global feature $x_0$ is assigned as the class label $y$, where the parameter
+$\varpi$ is learned during training process.
 
@@
 
+The inclusion of the hidden variable $\hb$ enhances the interpretability of the
+model. However, training the model using maximum likelihood estimation
+necessitates the calculation of the summation
+$\sum_{\hb \in H^N} \PP(y, \hb \mid \xb; \theta)$, and the presence of the
+hidden variable renders the exact computation of this summation intractable.
+To circumvent this challenge, a mean-field variational inference framework is
+employed, enabling efficient estimation of the log-likelihood
+$\log \PP(y \mid \xb; \theta)$.
+
 \subsubsection{ELBO Under Mean-Filed Variational Family}
 
-The existence of the hidden variable $\hb$ makes our model more interpretable.
-However, since We will train our model through maximal likelihood method,
-introducing the hidden variable will make precisely computation the summation
-$\sum_{\hb \in H^N} \PP(y, \hb \vert \xb; \theta)$ intractable.
-To address this issue, we make our computation more feasible through a
-mean-field variational inference framework, which enables our model
-efficiently estimating the log-likelihood $\log\PP(y \vert \xb; \theta)$.
+The mean-field varepsilon family approximation of
+$\log \PP(\hb \mid y, \xb; \theta)$ is given by:
+\begin{equation*}
+\QQ(\hb) = \prod_{i=1}^N q_i(h_i),
+\end{equation*}
+where $q_i(h_i)$ represents the approximate distribution of the individual
+hidden variable.
 
-Firstly, we introduce the mean-field varepsilon family approximation of
-$\log\PP(\hb \vert y, \xb; \theta)$ as:
-\begin{equation*}
-\QQ(\hb) = \prod_{i=1}^N q_i(h_i).
-\end{equation*}
-Then the evidence lower bound (ELBO) under the mean-field family is given by
+Under the mean-field family, the evidence lower bound (ELBO) is given by
 \begin{equation}
 \begin{split}
-& \textrm{ELBO}(\QQ) = \EE_{\QQ(\hb)} \log\PP(y, \hb \vert \xb; \theta)
+& \Lc(\QQ) = \EE_{\QQ(\hb)} \log\PP(y, \hb \mid \xb; \theta)
 - \EE_{\QQ(\hb)} \log q(\hb) \\
@@
 - \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2}
-- \frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib) \right\rvert^2}{2\eta_\beta^2}
-\right) \\
+- \frac{\left\lvert x_i - x_j \right\rvert^2}{2\eta_\beta^2} \right) \\
 &\quad + \omega_2 \exp \left(
@@
 \right) \Bigg] \\
-&\quad - \sum_{i=1}^N \EE_{q_i} \log q_i(h_i) + \vartheta(y, \xb; \varpi).
+&\quad - \sum_{i=1}^N \EE_{q_i} \log q_i(h_i) + \vartheta(y, x_0; \varpi).
 \end{split}
@@
 
+\begin{figure*}[tbp]
+\centering
+\includegraphics[width=\textwidth]{workframe}
+\caption{\sx{caption}.}
+\label{fig:workframe}
+\end{figure*}
+
 \subsubsection{Updating Variational Family}
 
-By the coordinate ascent variational inference(CAVI) algorithm,
-we can obtain that fix $q_j(h_j), \forall j \neq i$, the optimal $q_i(h_i)$
-that maximizes $\textrm{ELBO}(q_i(h_i))$ is given by
-\begin{equation}
+By employing the coordinate ascent variational inference (CAVI) algorithm,
+an efficient method for approximating posterior distributions in probabilistic
+models, the optimization process involves iteratively updating each variational
+factor while holding others fixed. Specifically, fixing $q_j(h_j)$,
+$\forall j \neq i$, the optimal $q_i(h_i)$ that maximizes $\Lc(q_i(h_i))$
+is computed as:
+\begin{equation*}
 \QQ(h_i) \propto \exp\{\EE_{\QQ(\hb_{-i})}
-\log\PP(y, h_i, \hb_{-i} \vert \xb; \theta)\}.
-\end{equation}
+\log\PP(y, h_i, \hb_{-i} \mid \xb; \theta)\}.
+\end{equation*}
 Thus, we can get that
-\begin{equation}
+\begin{equation*}
 \begin{split}
@@
 &+ \sum_{j \neq i} \sum_{l^\prime} q(l^\prime) \mu(l, l^\prime) 
-\Bigg[ \omega_1 \exp \big(
+\Bigg[ \omega_1 \exp \bigg(
 - \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2}\\
-&- \frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib) \right\rvert^2}{2\eta_\beta^2}
-\big)
-+ \omega_2 \exp \big(
+&- \frac{\left\lvert x_i - x_j \right\rvert^2}{2\eta_\beta^2}
+\bigg)
++ \omega_2 \exp \left(
 - \frac{\left\lvert p_i - p_j 
 \right\rvert^2}{2\eta_\gamma^2}
-\big)
+\right)
 \Bigg]
-\Bigg\}
+\Bigg\},
 \end{split}
-\end{equation}
+\end{equation*}
+where $Z_i$ is the normalizing constant.
 
@@
 
-\begin{itemize}
-\item {\textbf{Updating Variational Family}} $\{q_i(h_i)\}$:
-The computationally expensive part of the fixed point iteration of $q_i(h_i)$
-comes from the message passing, i.e. the convolution of $\QQ(\hb)$ and
-Gaussian kernal in $\psi(h_i, h_j, x_i, x_j; \eta)$. We need $\Ocal(N^2)$
-runtime for $N$ pixels when evaluating this term precisely.
-To make the computation more feasible, a truncated gaussian kernal will be
-employed, which is only supported on spaces within certain proportion of
-the untruncated standard deviation. Then the approximate message passing is
-linear in the number of pixels $N$.
-\item {\textbf{Computation of the ELBO}}: The summation over the product of
-$q_i(h_i)q_j(h_j)$ and the gaussian kernal in $\psi(h_i, h_j, x_i, x_j; \eta)$
-also requires a $\Ocal(N^2)$ complexity in time.
-Owes to a similar truncated gaussian kernal approxiamtion, the runtime can
-be reduced to $\Ocal(N)$.
-\end{itemize}
+\paragraph{Updating variational family}
+
+The computationally intensive step in the fixed-point iteration of $q_i(h_i)$
+is the message passing process, which involves convolving $\QQ(\hb)$ with a
+Gaussian kernel within the binary potential $\psi(h_i, h_j, x_i, x_j; \eta)$.
+This operation requires $\Oc(N^2)$ runtime for $N$ pixels when evaluated
+exactly. To improve computational efficiency, a truncated Gaussian kernel is
+utilized, which is supported only within a specific proportion of the
+untruncated standard deviation. This approximation reduces the complexity of 
+message passing to linear time, $\Oc(N)$.
+
+\paragraph{Computation of the ELBO}
+
+The summation over the product of $q_i(h_i)q_j(h_j)$ and the Gaussian kernel
+in $\psi(h_i, h_j, x_i, x_j; \eta)$ also incurs a time complexity of $\Oc(N^2)$.
+By applying a similar truncated Gaussian kernel approximation, the runtime can
+be reduced to $\Oc(N)$.
 
 \section{Results}
-%Results: Description of your experiments and results and a list of questions
-%your experiments are designed to answer.
 
 \section{Conclusion}
-%Conclusion: discussion and future work.
 

### 2024-12-11 hidden_crfnn_layer.py
add final code for this project

@@ -1,0 +1,403 @@
+"""
+MIT License
+
+Copyright (c) 2024 Xiaohang Ma
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import os
+import numpy as np
+import tensorflow as tf
+# from keras import ops
+from keras.layers import Layer
+# from keras import layers
+
+custom_module = tf.load_op_library(os.path.join(os.getcwd(), 'crfasrnn_keras', 'src','cpp', 'high_dim_filter.so'))
+# custom_module = tf.load_op_library(os.path.join( 'cpp', 'high_dim_filter.so'))
+from tensorflow.python.framework import ops
+@ops.RegisterGradient('HighDimFilter')
+def _high_dim_filter_grad(op, grad):
+    """ Gradients for the HighDimFilter op. We only need to calculate the gradients
+    w.r.t. the first input (unaries) as we never need to backprop errors to the
+    second input (RGB values of the image).
+
+    Args:
+    op: The `high_dim_filter` operation that we are differentiating.
+    grad: Gradients with respect to the output of the `high_dim_filter` op.
+
+    Returns:
+    Gradients with respect to the input of `high_dim_filter`.
+    """
+
+    rgb = op.inputs[1]
+    grad_vals = custom_module.high_dim_filter(grad, rgb,
[…]

### 2024-12-12 manuscript/main.tex
edit experiment

@@
 \usepackage{bbm}
+\usepackage[inline]{enumitem}
 \usepackage{graphicx}
@@
 \usepackage[numbers]{natbib}
+\usepackage{subcaption}
 \usepackage{xcolor}
@@
 \newcommand{\RR}{{\mathbb{R}}}
+\newcommand{\XX}{{\mathbb{X}}}
+\newcommand{\YY}{{\mathbb{Y}}}
 
@@
 \renewcommand{\subsubsection}[1]{\oldsubsubsection{\textbf{#1}}}
+
+\let\proglang=\textsf
+\let\pkg=\texttt
 
@@
 
+\section{Experiment Details}
+
+\subsection{Dataset Description}
+
+The dataset used in this research is the Harmful Brain Activity Classification
+(HMS) dataset~\citep{jing2023development}, provided by Harvard Medical School
+for research purposes. This dataset comprises EEG recordings and their
+corresponding spectrograms, segmented into 50-second-long samples derived from
+10-minute windows of brain activity.
+The longer 10-minute window allows for the model to pick up on further insights
+regarding the nature of the brain activity. Each EEG sample, representing 50
+seconds of brain activity measured by 19 scalp electrodes, is paired with a
+matching spectrogram.
+The EEG samples were labeled by expert annotators from Critical Care EEG
+Monitoring Research Consortium (CCEMRC) into one of six types of brain
+activity: seizures, generalized periodic discharges (GPD), lateralized
+periodic discharges (LPD), lateralized rhythmic delta activity (LRDA),
+generalized rhythmic delta activity (GRDA), and other.
+
+Expert agreement was high for some segments; nevertheless, instances of
+disagreement were observed. The segments were categorized as follows:
+\begin{enumerate*}[label = (\roman*)]
+\item ``Idealized'': Segments with high levels of expert agreement.
+\item ``Proto'': Segments where approximately half of the experts labeled them
+as ``other'', while the other half assigned one of the remaining five labels.
+\item ``edge'': Segments where experts were roughly split between two of the
+five named patterns.
+\end{enumerate*}
+Figure~\ref{fig:ex} shows an example of EEG patterns with different level of
+agreement.
+
+\begin{figure}[tbp]
+\centering
+\includegraphics[width=.48\textwidth]{example}
+\caption{Examples of EEG patterns with different levels of expert agreement,
+sourced from the HMS dataset.}
+\label{fig:ex}
+\end{figure}
+
+\subsection{Data Preparation}
+
+The data preparation process involves multiple steps to extract spectrogram
+features from EEG data, enhance signal quality, and ensure the features are
+appropriately scaled and normalized for analysis.
+
+\textbf{Wavelet denoising}: EEG signals were denoised using wavelet
+transformation to improve signal quality by decomposing the signal into wavelet
+coefficients, applying thresholding to remove noise, and reconstructing the
+denoised signal.
+%via \pkg{pywt} \proglang{Python} package
+
+\textbf{Feature Extraction}: A bipolar montage technique is applied to the
+denoised EEG signals to calculate the difference between adjacent electrode
+signals, generating four representative montage features from the original 19
+features. Figure~\ref{fig:inion} illustrates the bipolar electrode chains used
+in this study.
+
+\textbf{Spectrogram transformation}: The denoised EEG signals are converted
+into Mel-spectrograms, which are log-transformed to enhance spectral
+information and compress the dynamic range. Figure~\ref{fig:EEG} shows an
+example of the denoised EEG signal and its corresponding transformated
+spectrogram.
+%using the \pkg{librosa} \proglang{Python} package
+
+\textbf{Normalizatio}n: The Mel-spectrogram values are standardized to a range
+of $[-1,1]$ to ensure consistent feature scaling and facilitate model training.
+
+\textbf{Imputation}: Missing values in the data are imputed with the mean of
+non-missing values within each spectrogram. If all values are missing,
+the spectrogram is set to zero to maintain data consistency.
+
+\begin{figure}[tbp]
+\centering
+\includegraphics[width=.45\textwidth]{inion}
+\caption{Chains employed to determine representative montage features via
+a bipolar montage, sourced from Kaggle.}
+\label{fig:inion}
+\end{figure}
+
+\begin{figure}
+\centering
+\begin{subfigure}[b]{0.45\textwidth}
+\includegraphics[width=\textwidth]{EEG_Signal}
+\end{subfigure}
+\hfill
+\begin{subfigure}[b]{0.45\textwidth}
+\includegraphics[width=\textwidth]{EEG_Spectrogram}
+\end{subfigure}
+\caption{An example of the denoised EEG signal (top) and its corresponding
+	spectrogram representation (bottom).}
+\label{fig:EEG}
+\end{figure}
+
+\subsection{Loss Function}
+
+The model outputs an array of predicted probabilities for each label class.
+To evaluate its performance, the Kullback-Leibler (KL) divergence metric is
+employed. This metric is suitable for comparing two probability distributions: 
+the ground truth probability distribution $\XX$ and the predicted probability
+distribution $\YY$. The KL divergence between these two distributions is
+defined as:
+\begin{equation*}
+D_\mathrm{KL}(\XX \Vert \YY) = \sum \XX \log\left(\frac{\XX}{\YY}\right)
+\end{equation*}
+
+The KL divergence offers the advantage of direct applicability as a loss
+function during model training, eliminating the need for an alternative loss
+function. Its differentiability enables efficient gradient-based optimization
+techniques.
+
+\subsection{Learning Rate Scheduler}
+
+To optimize the model's training process, a cosine annealing learning rate
+scheduler was employed. This technique ensures a smooth and gradual decrease in
+the learning rate, preventing abrupt changes that can destabilize the training
+process. By allowing the model to explore a wider parameter space initially and
+then refining its solution towards the end, cosine annealing contributes to
+better convergence and reduced oscillations in the loss function.
+
+As depicted in Figure~\ref{fig:cos}, the learning rate for both the pre-trained
+UNet and the fine-tuned HCRF exhibit a similar pattern. Both start with a
+relatively high learning rate, which gradually decreases over time.
+This smooth decay helps ensure stable training and optimal convergence.
+
+\begin{figure}
+\centering
+\begin{subfigure}[b]{0.45\textwidth}
+\includegraphics[width=\textwidth]{pretrain}
+\end{subfigure}
+\hfill
+\begin{subfigure}[b]{0.45\textwidth}
+\includegraphics[width=\textwidth]{finetune}
+\end{subfigure}
+\caption{The learning rate schedulers for the pre-trained UNet (top) and
+the fine-tuned HCRF (bottom).}
+\label{fig:cos}
+\end{figure}
+
 \section{Results}

### 2024-12-12 manuscript/main.tex
edit intro

@@
 \section{Introduction}
-%Introduction: problem definition and motivation.
 
-\sx{EEG tasks}
+Electroencephalography (EEG) pattern recognition plays a pivotal role in
+various fields, including neuroscience, clinical medicine, and human-computer
+interaction. By analyzing brain electrical activity, we can gain deeper
+insights into brain states and functions, providing a scientific basis for
+diagnosing and treating neurological disorders, as well as developing
+brain-computer interfaces. EEG pattern recognition not only helps to unveil the
+mechanisms of the brain but also enables the development of smarter and more
+human-centric technologies. However, traditional manual analysis of EEG signals
+is a time-consuming and subjective process, often leading to inconsistencies in
+diagnosis~\citep{liu2013recognizing}.
 
-Sequence labeling tasks are pivotal in diverse domains, especially healthcare,
-where precise and timely data classification can significantly impact patient
-outcomes. Medical datasets, such as physiological signals, often exhibit
-complexity and heterogeneity, posing challenges for traditional modeling
-techniques. Therefore, developing models capable of extracting meaningful
-patterns from these datasets is crucial for accurate diagnostic and prognostic
-decisions.
-
-In recent years, deep learning has emerged as a powerful tool for sequence
-labeling. Neural network architectures like long short-term memory (LSTM) and
-Transformer models have revolutionized sequential data processing.
-Additionally, conditional random fields (CRFs) have been employed for modeling
-label dependencies, demonstrating promising results.
-However, while these methods have shown individual strengths, a comprehensive
+In recent years, deep learning has emerged as a powerful tool to automate and
+improve EEG pattern recognition~\citep{craik2019deep}. Neural network
+architectures like convolutional neural network (CNN), long short-term memory
+(LSTM) and Transformer models have revolutionized EEG data processing.
+Additionally, hidden conditional random fields (HCRFs) have been employed for
+modeling label dependencies, demonstrating promising results.
+While these methods have shown individual strengths, a comprehensive
 approach that leverages their complementary capabilities is essential for
@@
 This project introduces a novel framework that combines deep neural networks
-and probabilistic graphical models (PGMs) to enhance sequence labeling
-performance. By employing Transformer~\citep{vaswani2017attention} for feature
-extraction and integrating them with CRF models,
-we aim to improve sequence tagging accuracy. Furthermore, mean-field
+and probabilistic graphical models (PGMs) to enhance the accuracy of EEG
+pattern recognition . By employing UNet~\citep{ronneberger2015unet} for feature
+extraction and integrating them with HCRF models, we aim to improve the
+precision of EEG pattern recognition tagging. Furthermore, mean-field
 approximation techniques are utilized for efficient inference, ensuring a
 computationally efficient yet effective approach.
-This integration of modern neural network architectures with CRFs offers
-a robust and promising solution for complex sequence labeling tasks.
+This integration of modern neural network architectures with HCRFs offers
+a robust and promising solution for EEG pattern recognition tasks.
 

### 2024-12-12 manuscript/main.tex
update background; fig position

@@
 \section{Background}
-%Background & Related Work: background info and literature survey.
 
-\sx{Add UNet}
+Deep neural networks, particularly convolutional architectures like U-net, have 
+proven highly effective for image segmentation tasks due to their ability to
+capture rich spatial features and contextual information at different scales.
+U-net, in particular, has demonstrated remarkable performance in medical image
+segmentation by leveraging its encoder-decoder architecture with skip
+connections to preserve spatial resolution and improve
+localization~\citep{ronneberger2015unet,zhou2018unet}.
 
-Deep neural networks, particularly recurrent neural networks (RNNs) like LSTM,
-are widely used for sequence labeling due to their ability to capture
-long-range dependencies in sequences. Bidirectional LSTM (Bi-LSTM) further
-enhances this by processing sequences in both directions, improving context
-understanding.
-CRFs, on the other hand, are commonly used for structured predictions,
-modeling dependencies between neighboring labels to ensure valid output
-sequences.
-\citet{huang2015bidirectional} pioneered the use of Bi-LSTM integrated with
-CRFs for sequence labeling tasks, demonstrating their effectiveness in
-extracting features and ensuring label consistency.
+Traditional sequence-labeling approaches, such as recurrent neural networks
+(RNNs) like LSTM and bidirectional LSTMs (Bi-LSTM), have been effective in
+capturing long-range dependencies and processing temporal information. However,
+these models are less suited for spatially rich tasks like image segmentation.
+Moreover, LSTMs and Bi-LSTMs often require large labeled datasets, are
+computationally expensive, and struggle to account for the structural
+dependencies present in spatial data, which limits their applicability in
+resource-constrained environments like hospitals.
 
-However, Bi-LSTMs can still struggle to fully account for structural
-dependencies between predicted labels, leading to potential inconsistencies in
-the output sequence. Additionally, they typically require large amounts of
-labeled data and are computationally expensive, making them challenging to
-deploy in resource-constrained environments, such as hospitals.
+Transformer-based models, such as BERT, have emerged as powerful tools for
+sequence-labeling tasks, demonstrating superior performance over traditional
+RNN-based models. However, their advantages are primarily realized in natural
+language processing rather than spatially dependent tasks like image
+segmentation.
 
-Transformer-based models, such as BERT~\citep{devlin2019bert},
-have emerged as powerful alternatives to traditional RNN-based approaches.
-Trained on massive datasets in an unsupervised manner, Transformer can be
-fine-tuned for various natural language processing tasks,
-including sequence labeling.
-\citet{devlin2019bert} directly compared BERT to traditional models, including
-Bi-LSTM, and demonstrated its superior performance on tasks like named entity
-recognition, highlighting its ability to capture more feature information
-than LSTM models.
+U-net addresses these challenges by focusing on capturing structural
+dependencies in spatial domains. For EEG data, where the objective shifts from
+sequence-labeling to segmentation, U-net is a more appropriate choice. Its
+ability to efficiently learn and propagate contextual and spatial information
+makes it a robust alternative to sequence-based methods like Bi-LSTM.
 
@@
 
+\begin{figure*}[ht]
+\centering
+\includegraphics[width=\textwidth]{workframe}
+\caption{The proposed feature-learning-based HCRF using UNet feature maps.}
+\label{fig:workframe}
+\end{figure*}
+
+Figure~\ref{fig:workframe} illustrates the descriptive workframe layout
+employed in this study, highlighting the key components and processes involved.
+
 \subsection{Variational Inference of HCRF}
@@
 \end{equation}
-
-\begin{figure*}[tbp]
-\centering
-\includegraphics[width=\textwidth]{workframe}
-\caption{\sx{caption}.}
-\label{fig:workframe}
-\end{figure*}
 
@@
 
+\begin{figure}[ht]
+\centering
+\includegraphics[width=.48\textwidth]{example}
+\caption{Examples of EEG patterns with different levels of expert agreement,
+sourced from the HMS dataset.}
+\label{fig:ex}
+\end{figure}
+
 Expert agreement was high for some segments; nevertheless, instances of
@@
 agreement.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=.48\textwidth]{example}
-\caption{Examples of EEG patterns with different levels of expert agreement,
-sourced from the HMS dataset.}
-\label{fig:ex}
-\end{figure}
 
@@
 
+\begin{figure}[bp]
+\centering
+\includegraphics[width=.45\textwidth]{inion}
+\caption{Chains employed to determine representative montage features via
+a bipolar montage, sourced from Kaggle.}
+\label{fig:inion}
+\end{figure}
+
 \textbf{Feature Extraction}: A bipolar montage technique is applied to the
@@
 in this study.
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}[b]{0.45\textwidth}
+\includegraphics[width=\textwidth]{EEG_Signal}
+\end{subfigure}
+\hfill
+\begin{subfigure}[b]{0.45\textwidth}
+\includegraphics[width=\textwidth]{EEG_Spectrogram}
+\end{subfigure}
+\caption{An example of the denoised EEG signal (top) and its corresponding
+spectrogram representation (bottom).}
+\label{fig:EEG}
+\end{figure}
 
@@
 the spectrogram is set to zero to maintain data consistency.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=.45\textwidth]{inion}
-\caption{Chains employed to determine representative montage features via
-a bipolar montage, sourced from Kaggle.}
-\label{fig:inion}
-\end{figure}
-
-\begin{figure}
-\centering
-\begin{subfigure}[b]{0.45\textwidth}
-\includegraphics[width=\textwidth]{EEG_Signal}
-\end{subfigure}
-\hfill
-\begin{subfigure}[b]{0.45\textwidth}
-\includegraphics[width=\textwidth]{EEG_Spectrogram}
-\end{subfigure}
-\caption{An example of the denoised EEG signal (top) and its corresponding
-spectrogram representation (bottom).}
-\label{fig:EEG}
-\end{figure}
 
@@
 
-As depicted in Figure~\ref{fig:cos}, the learning rate for both the pre-trained
-UNet and the fine-tuned HCRF exhibit a similar pattern. Both start with a
-relatively high learning rate, which gradually decreases over time.
-This smooth decay helps ensure stable training and optimal convergence.
-
-\begin{figure}
+\begin{figure}[tbp]
 \centering
@@
 
+As depicted in Figure~\ref{fig:cos}, the learning rate for both the pre-trained
+UNet and the fine-tuned HCRF exhibit a similar pattern. Both start with a
+relatively high learning rate, which gradually decreases over time.
+This smooth decay helps ensure stable training and optimal convergence.
+
 \section{Results}

### 2024-12-12 manuscript/main.tex
edit abs, res

@@
 
-%In this project, we will integrate powerful deep neural networks and
-%statistical methods with a probabilistic graphical model (PGM) to effectively
-%tackle sequence labeling tasks on a heterogeneous medical care dataset.
+Electroencephalography (EEG) pattern recognition presents critical challenges
+in neurological diagnostics, where precise classification of brain activity can
+significantly impact patient outcomes. The complexity of EEG signals pose
+substantial challenges for traditional modeling techniques. This research
+introduces a novel framework integrating deep neural networks and probabilistic
+graphical models to enhance EEG pattern recognition performance. The proposed
+approach combines UNet-based feature extraction with hidden conditional random
+fields (HCRFs) to improve brain activity classification accuracy.
+The methodology employs advanced preprocessing techniques, including wavelet
+denoising and spectrogram transformation of EEG signals from the Harmful Brain
+Activity Classification dataset, which contains 50-second brain activity
+samples labeled into six distinct categories. The research addresses challenges
+in neurological signal processing by developing a sophisticated model capable
+of capturing intricate temporal dependencies and handling varying levels of
+expert annotation agreement. Experimental results demonstrated significant
+improvements in classifying complex EEG patterns across different levels of
+expert consensus. The complete implementation code is available on GitHub at
+\url{https://github.com/Carol-seven/UNet-HCRF_EEG}.
 
@@
 (LSTM) and Transformer models have revolutionized EEG data processing.
-Additionally, hidden conditional random fields (HCRFs) have been employed for
+Additionally, conditional random fields (CRFs) have been employed for
 modeling label dependencies, demonstrating promising results.
@@
 
-\begin{figure*}[ht]
+\begin{figure*}[tp]
 \centering
@@
 
-\begin{figure}[ht]
+\begin{figure}[tp]
 \centering
@@
 
-\begin{figure}[bp]
+\begin{figure}[tp]
 \centering
@@
 
-\section{Results}
+\section{Results and Discussion}
 
-\section{Conclusion}
+In this study, we developed a novel UNet-HCRF model for EEG pattern recognition.
+Our model achieved a competitive KL divergence of 0.4992 on a five-fold
+cross-validation, demonstrating competitive performance compared to the
+baseline EfficientNet model from Kaggle, which had a KL divergence of 0.59.
+However, our results indicated room for further improvement, as state-of-the-art
+models such as Vision Transformers (ViT) have achieved significantly lower KL
+divergence scores, ranging between 0.2 and 0.3.
+
+The proposed UNet-HCRF framework integrates a latent CRF layer with a neural
+network architecture to capture intricate segmentation information from EEG
+dataset. By employing a variational mean-field inference scheme and an
+end-to-end training strategy, we efficiently learned both neural network
+features and the CRF parameters. This approach offers a flexible and powerful
+framework for EEG analysis, leveraging the strengths of both deep learning and
+probabilistic graphical models.
+
+Our findings demonstrate the potential of HCRF models for EEG pattern
+classification. However, several avenues for future research remain. For
+instance, exploring alternative variational inference techniques or
+incorporating attention mechanisms within the CRF layer could potentially
+enhance model performance. Additionally, investigating the impact of different
+neural network architectures and hyperparameter tuning on the overall model
+performance is also a worthwhile direction for further research.
 

### 2024-12-12 manuscript/main.tex
minor
typo

@@
 \newcommand{\one}{\mathbbm{1}}
-\newcommand{\Ocal}{\mathcal{O}}
 \newcommand{\RR}{{\mathbb{R}}}
@@
 and probabilistic graphical models (PGMs) to enhance the accuracy of EEG
-pattern recognition . By employing UNet~\citep{ronneberger2015unet} for feature
+pattern recognition. By employing UNet~\citep{ronneberger2015unet} for feature
 extraction and integrating them with HCRF models, we aim to improve the
